### 2017-02-06 sedona-ieice16/revision/reviews/review_response.tex
SEDONA updates

\documentclass[10pt,letterpaper]{article}

\usepackage{graphicx}
\usepackage{longtable}
\usepackage{algorithmic}
\usepackage{alltt}
\usepackage{array}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
%\usepackage[caption=false]{caption}
\usepackage[dvipsnames]{color}
%\usepackage{colortbl}
%\usepackage{enumitem}
%\usepackage{eqparbox}
%\usepackage{fixltx2e}
%\usepackage{float}
%\usepackage{floatflt}
%\usepackage{mdwmath}
%\usepackage{mdwtab}
%\usepackage{multirow}
\usepackage{stfloats}
%\usepackage[font=footnotesize]{subfig}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\usepackage{subfigure} 
\usepackage{times}
\usepackage{url}
\usepackage{verbatim} 
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{color}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilite{{\bf TODO:} {\em #1}}}

\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}

% new commands: from file -- to permit re-use
%\input{newcommands.tex}

%Hyphenation problems
\hyphenation{name-space}
\hyphenation{schema-location}
\hyphenation{xml-lint}
\hyphenation{time-stamp}
\hyphenation{time-stamps}
\hyphenation{time-stamped}

\urlstyle{rm}
\graphicspath{{./figures/}}

\definecolor{gray}{rgb}{0.5,0.5,0.5}
%\newcommand{\rev}[1]{\textcolor{blue}{#1}}
%\newcommand{\quo}[1]{\textcolor{gray}{#1}}
%\newcommand{\rev}[1]{\vspace{3mm} \noindent {\em {\color{blue}{#1}}}}

\newenvironment{myindentpar}[1]%
{\begin{list}{}
         {\vspace{10pt}
					\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\rev}[1]{\begin{myindentpar}{.25in} {\em {\color{blue}{#1}}}\end{myindentpar}}

\newenvironment{myindentparQ}[1]%
{\begin{list}{}
         {\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\quo}[1]{\begin{myindentparQ}{.25in} {{\color{gray}{#1}}}\end{myindentparQ}}

\begin{document}

\title{Author's Reply}
\author{}
\maketitle

\section*{Overview}\label{sec:overview}
I appreciate these quite thorough reviews. I appreciate the feedback and suggestions.

\newcounter{RP}

\clearpage
\section*{Reviewer}\label{sec:rev1}

% 1
\rev{
$<<$ Reviewer's comments to the author(s) $>>$

The SEDONA is based on the probability that any daemons run during the
measurement. And cutoff time is based on the actual measurement time.
The proposed approach is interesting, but I cannot understand why the
author employs such statistics and heuristic way, and how efficient the
proposed approach will work. Please refer to my comments, and consider
to revise the manuscript and re-submit .
}

Your concerns are well understood. I respond to individually below. 
%Your comments are incorporated into the revision.

\rev{
Comments: 
(1) In the last of the manuscript, the author describes ``Our plan is to
integrate SEDONA into the query timing protocol [8]." If the research
goal is the combination of the SEDONA and the query timing protocol,
the proposed approach may be suitable. However, the goal is not
clearly described in the manuscript. The readers will recognize the
SEDONA is a generic protocol. If the research goal is the combination
of this work and the prior work, the author should mention so.
}

The described plan (in the previous paper) is one possibility in the future, 
but that is not stated as this paper's research goal. 
Our goal of this paper is to propose a {\em generic, better} timing protocol 
against the conventional approach of utilizing elapsed time. 
The goal is successfully achieved by significantly reducing variability of 
the execution time of a program of interest in our experiments. 
I now clearly state this goal, marked in red, at the last sentence in Section~1.

%Note also that the mention of my plan in the future work 
%is just one aspect of the extension of the SEDONA protocol. 
%I claim that the combination is not the ultimate research goal of this paper. 
%To avoid confusion, I removed that sentence.

\rev{
(2) If the proposed approach is for a generic purpose, there is a strong
doubt about usability. The improvement of the SEDONA is little. And the
extent of the each benchmark’s improvement described in Table 3
varies widely. Thus, the experimental results does not confirm whether
the SEDONA is widely practical or is useful. The description to clarify
the SEDONA’s characteristics is needed.
}

Your concern is understood, but it is hard to agree with the comment that the improvement is little, for the following reasons.
First, for a certain workload such as {\tt 434}, 
our protocol reduced the variability was reduced up to by about 10x (from 75 ms to 7 ms), 
which is *not* small. (Please refer to the last sentence 
in the first paragraph in the second column on page~4.)
There were also some other workloads such as {\tt 410} and {\tt 445} 
in which the improvement reached about 3x, respectively. 
Furthermore, every benchmark's standard deviation by our protocol 
was less than that of the conventional method. 
The relative error of the benchmarks was on average improved 
by 1.5x as well.

Second, our protocol scaled well over increasing execution time. 
(Please refer to the second paragraph in the right column on page~4.)
The relative error of our protocol was still smaller than that of the ORG method 
for a long-running benchmark like {\tt 436} and {\tt 454} 
as well as for a medium-length or a short-length one. 
Such scalability was still observed in our new experiment, 
which I describe shortly, as shown in Fig.~5.

Note also that the ET results by the ORG method were based 
on the settings taking care of several identified timing factors, such as
(i)~deactivating non-critical daemons, 
(ii)~switching on the network timing protocol daemon, and 
(iii)~kernel upgrade. 
(The description of these settings is marked in red in the paragraph below Fig.~1.) 
In the well-controlled measurement environment, our protocol further 
reduced the variability of execution time against the ORG method
in such various benchmarks, which actually demonstrated its general purpose and wide usability.

Lastly, the results from a new experiment provide 
another strongly evidence that our protocol is widely practical and useful. 
In the new experiment we evaluated the performance of our protocol 
on some real-world programs like insertion sort and matrix multiplication (in column major). 
The experimental results showed that the performance gap between the original and our methods 
was increasingly widened, e.g., up to by 6x, as workload level increased.

Therefore, I strongly claim that these various experiments confirm 
the wide acceptance and usability of SEDONA as an alternative to the ET-based measurement technique.

\rev{ 
(3) The improvement of the SEDONA is little. And the extent of the each
benchmark’s improvement described in Table 3 varies widely. Thus, the
experimental results does not confirm whether the SEDONA is widely
practical or is useful.
}

This review is basically the same as (2). Please refer to our response of (2) for this comment. 
One note. You may feel that the extent of the improvement varies widely, 
because the SPEC CPU2006 includes a set of very different compute-bound benchmarks. 
But even if so, we found that all the benchmarks in the suite benefited from our timing protocol, and in addition, 
for some workloads, our protocol significantly reduced their variability up to by 10x.

\rev{ 
(4) Since the explanation of ``dual-PUT'' is not enough, I can not well
understand the algorithm shown in Fig. 3. 
}

More descriptions of the term ``dual-PUT'' is provided in Section~2. 
\rev{
Similarly, what is “1st
Half’s ET (ms)” attached at Fig. 4(b), 4(c), 4(d) ? I understood only that
the label is related with “dual-PUT”. These words are important in the
manuscript. So, please keep in mind to write understandably.
}

\end{document}