### 2017-02-07 sedona-ieice16/revision/cover_letter.tex
Revision done

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Young-Kyoon Suh}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{
%Editors-in-Chief of ACM Transactions on Computer Systems (TOCS): \\
%\begin{tabular}{ll}
%\vspace{.1in}
%
%Prof. Todd C. Mowry\\
%Computer Science Department\\
%Carnegie Mellon University\\
%5000 Forbes Avenue\\
%Pittsburgh, PA 15213 \\
%USA\\ 
}

\begin{flushleft}
{\large\bf Young-Kyoon Suh}
\end{flushleft}
\medskip\hrule height 1pt
%\begin{flushright}
%\hfill Department of Computer Science, The University of Arizona, Tucson, AZ 85721 USA\\
%\hfill Tel: +1-520-370-1865
%(Please understand we cannot reveal our affiliation, as the manuscript should be double-blind.)
%\end{flushright} 
\begin{flushright}
%\hfill National Institute of Supercomputing and Networking, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Korea Institute of Science and Technology Information (KISTI), 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Republic of Korea, 34141
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Tel: +82-42-869-0725
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Email: youngkyoon.suh@gmail.com
\end{flushright} 
%\vfill % forces letterhead to top of page

\vspace{0.5in}

\opening{Dear Prof. Rajkumar Buyya:}

\noindent

Attached is a minor revision that responds to the first reviewer's comments. 
We marked in red some text in response to his/her comments. 
The revision also includes some texts marked in blue that are added or updated. 
Note also that we replaced the term of {\em PUT} with another term, called {\em INC}, 
in order to prevent readers from getting confused with a general concept of the program under test. 
Many thanks to you for taking sincere care of our manuscript during a series of reviews.

\closing{Sincerely yours,}

\end{letter}

\end{document}