### 2017-02-07 sedona-ieice16/revision/cover_letter.tex
SEDONA final updates

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Young-Kyoon Suh}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{
%Editors-in-Chief of ACM Transactions on Computer Systems (TOCS): \\
%\begin{tabular}{ll}
%\vspace{.1in}
%
%Prof. Todd C. Mowry\\
%Computer Science Department\\
%Carnegie Mellon University\\
%5000 Forbes Avenue\\
%Pittsburgh, PA 15213 \\
%USA\\ 
}

\begin{flushleft}
{\large\bf Young-Kyoon Suh}
\end{flushleft}
\medskip\hrule height 1pt
%\begin{flushright}
%\hfill Department of Computer Science, The University of Arizona, Tucson, AZ 85721 USA\\
%\hfill Tel: +1-520-370-1865
%(Please understand we cannot reveal our affiliation, as the manuscript should be double-blind.)
%\end{flushright} 
\begin{flushright}
%\hfill National Institute of Supercomputing and Networking, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Korea Institute of Science and Technology Information (KISTI), 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Republic of Korea, 34141
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Tel: +82-42-869-0725
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Email: youngkyoon.suh@gmail.com
\end{flushright} 
%\vfill % forces letterhead to top of page

\vspace{0.5in}

\opening{Dear Editor-in-Chief:}

\noindent

I am re-submitting a revised manuscript that responds to a previous reviewer's comments. 
I mark in red some text in response to the reviewer's comments. 
This revision also includes some text marked in blue that are added or updated, 
compared to its previous version.

Here are highlights of differences between the revision and the previous version.

\begin{itemize}

\item Research goal: the goal of this manuscript is clearly stated upfront in Section~1.

\item Characteristics: the distinctions of our protocol are discussed 
in the paragraph of ``Characteristics" in Section~1.

\item Timing environment: the description of our timing environment 
is moved upfront to Section~1 (from Section~2), 
to emphasize the fact that even though a timing environment was well configured, 
there still existed significant variability in the traditional measurement method.

\item Terminology: the definition of ``dual-PUT" is provided with its example in Section~2.

\item Experiments: the results from a new experiment are provided in Fig.~5 on page~4.

\end{itemize}

Many thanks to you for handling our manuscript.

\closing{Sincerely yours,}

\end{letter}

\end{document}