### 2017-02-09 sedona-ieice16/revision/resources/cover_letter.tex
SEDONA updates fitting 4 pages

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Young-Kyoon Suh}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{
%Editors-in-Chief of ACM Transactions on Computer Systems (TOCS): \\
%\begin{tabular}{ll}
%\vspace{.1in}
%
%Prof. Todd C. Mowry\\
%Computer Science Department\\
%Carnegie Mellon University\\
%5000 Forbes Avenue\\
%Pittsburgh, PA 15213 \\
%USA\\ 
}

\begin{flushleft}
{\large\bf Young-Kyoon Suh}
\end{flushleft}
\medskip\hrule height 1pt
%\begin{flushright}
%\hfill Department of Computer Science, The University of Arizona, Tucson, AZ 85721 USA\\
%\hfill Tel: +1-520-370-1865
%(Please understand we cannot reveal our affiliation, as the manuscript should be double-blind.)
%\end{flushright} 
\begin{flushright}
%\hfill National Institute of Supercomputing and Networking, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Korea Institute of Science and Technology Information (KISTI), 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Republic of Korea, 34141
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Tel: +82-42-869-0725
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Email: yksuh@kisti.re.kr
\end{flushright} 
%\vfill % forces letterhead to top of page

\vspace{0.5in}

\opening{Dear Dr. Koichi Mouri (Associated Editor):}

\noindent

I am submitting a revised manuscript that responds to a previous reviewer's comments. 
Here are highlights of differences between this revision and its previous version (paper ID number: 2016EDL8232).

\begin{itemize}

\item Research goal: the goal of this manuscript is clearly stated upfront in Section~1.

\item Characteristics: the distinctions of our protocol are discussed 
in the paragraph of ``Characteristics" in Section~1.

\item Timing environment: the description of our timing environment 
is moved upfront to Section~1 (from Section~2), 
to emphasize the fact that even though a timing environment was well configured, 
there still existed significant variability in the traditional measurement method.

\item Terminology: the definition of ``dual-PUT" is provided with its example in Section~2.

\item Experiments: the results from a new experiment are provided in Fig.~5 on page~4.

\end{itemize}

Many thanks to you for handling my manuscript.

\closing{Sincerely yours,}

\end{letter}

\end{document}