### 2017-02-23 subopt-tods16/paper_young.tex

\documentclass[prodmode,acmtods]{acmsmall}
\pdfminorversion=6

%Metadata Information
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2016}
\acmMonth{2}

\usepackage{url}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphics}
% \usepackage[total={6.5in,9in}, top=1.25in,left=0.9in]{geometry}
%\usepackage[text={6.5in,9in},centerpage,includefoot]{geometry}
% \usepackage{setspace}
\usepackage{color}
\usepackage{url}
\usepackage{balance}
\usepackage{amsmath}
\usepackage{amssymb}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{grey}{#1}}
\newcommand{\hilitey}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilitey{{\bf TODO:} {\em #1}}}
\long\def\shorten#1{}

%For last edit: revise and resubmit for SIGMOD'14
\long\def\shortenF#1#2{}%First is letter indicating what has been removed

%For double blind, keep the second argument. To turn off doubleblind, keep
%the first
\long\def\doubleblind#1#2{#2}
\def\azdb{\doubleblind{\hbox{\sc AZDBLab}}{\hbox{\sc DBLab}}}
\def\QatC{Q{@}C}
\long\def\comment#1{}
%c2j: Conference to Journal: first parameter is conference, second is journal
% For Conferences: c2j#1#2{#1}
% For Journals: c2j#1#2{2}
\long\def\c2j#1#2{#2}

\begin{document}

\title{%\vspace*{-0.75in}
A Causal Model of \hbox{DBMS} Suboptimality\\
}
\markboth{\ }{A Causal Model of \hbox{DBMS} Suboptimality}

% possible, but not really needed or used for PVLDB:
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as\textit{Author's Guide to Preparing ACM GIG Proceedings Using \LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}

% ****************** AUTHORS **************************************

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

%bug is vldb.cls gives an additional authors section for greater than 3...
\doubleblind{\numberofauthors{3}}{}
 %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
\doubleblind{% 1st. author
Richard T.~Snodgrass\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Arizona}\\
%       \affaddr{Tucson, AZ}\\
       \email{\small\tt rts@cs.arizona.edu}
\alignauthor
% 2nd. author
\alignauthor
Sabah Currim\\
       \affaddr{Alumni Office}\\
       \affaddr{University of Arizona}\\
%       \affaddr{Tucson, AZ}\\
%       %\vspace*{-0.05in}
       \email{\small\tt scurrim@email.arizona.edu}
% 3rd. author
\alignauthor
Young-Kyoon Suh\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Arizona}\\
%       \affaddr{Tucson, AZ}\\
       \email{\small\tt yksuh@cs.arizona.edu}
}{}}

\begin{abstract}
The query optimization phase within a \hbox{DBMS} ostensibly finds the fastest query
execution plan from a potentially large set of enumerated plans, all of
which correctly compute the specified query. Occasionally the optimizer
selects the wrong plan, for a variety of reasons. {\em Suboptimality} is
indicated by the existence of a query plan that performs more efficiently
than the DBMS's chosen plan, for the same query. From an engineering
perspective, it is of critical importance to understand the prevalence
of suboptimality and its causal factors.
We propose a novel structural causal model to explicate the
relationship between various factors in query optimization and
suboptimality. Our model associates suboptimality with the factors of
complexity of the query and optimizer and concomitant unanticipated
interactions among the components of the optimizer. This model induces a number of specific
hypothesis that were subsequently tested on multiple \hbox{DBMSes}. We
observe that \todo{Rick: check these}suboptimality prevalence
correlates positively with the number of operators available in the \hbox{DBMS}
(one proxy for optimizer complexity) and with the number of plans generated
by the optimizer (another proxy for optimizer complexity) and with
the number of correlation names (a measure of query complexity), providing
empirical support for this model as well as implications for fundamental
improvements of these optimizers. The results of this study also uncovered evidence of a potential fundamental
limitation of cost-based optimization.\todo{Rick:be more specific! Couple
  with contributions and conclusions}
This paper thus provides
a new methodology to study mature query optimizers, that of empirical
generalization, proposes a novel causal model for query suboptimality, and
tests several hypotheses deriving from this causal model.
\end{abstract}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures,
%performance measures]
\terms{}
\keywords{}
\acmformat{\doubleblind{Richard T.~Snodgrass, Sabah Currim, and Young-Kyoon
    Suh, 2017, A Causal Model of \hbox{DBMS} Suboptimality}{A Causal Model of
    \hbox{DBMS} Suboptimality}}

\doubleblind{\begin{bottomstuff}
This work was supported in part by NSF grants \hbox{IIS-0639106},
IIS-0415101, and EIA-0080123. 
%
%Author's addresses: G. Zhou, Computer Science Department,
\end{bottomstuff}}{}
\maketitle

\section{Introduction}\label{sec:intro}
DBMSes underlie all information systems and hence optimizing their performance is of
critical importance. The DBMS's query optimizer plays an important role. But
what if the optimizer {\em doesn't}: what if it selects the wrong plan?

This paper provides a thorough investigation into \hbox{DBMS} {\em
  suboptimality}: when the \hbox{DBMS} chooses a slower plan over a faster
plan for a given query. We systematically examine the factors influencing
the number of suboptimal queries. There could be multiple causes of the
suboptimality. One possible cause could be some peculiarity within the tens
of thousands of lines of code of that query optimizer. Another possible
cause could be the query's complexity. Prior research in other domains shows
that increasing complexity negatively influences
performance~\cite{campbell88,moody98}. A third possible reason could be some
fundamental limitation {\em within the general architecture of cost-based
  optimization} that will always render a number of queries suboptimal.

To better understand the impact of different factors on suboptimality of
query performance and the interaction between operators, especially in a
dynamic environment, an experimental approach is needed.
Our research introduces a novel approach
to better understand the factors influencing query performance in
\hbox{DBMSes}. Based on existing research and general knowledge of \hbox{DBMSes}, we
propose an innovative predictive model to better understand the presence and influence
of suboptimality in query evaluation. We use an experimental methodology on
a collection of \hbox{DBMSes} as subjects to test our hypotheses with respect to factors
influencing suboptimality, utilizing
experiment data collected over a \hbox{cumulative} 16,000 hours (over two
years) of query executions.
Our research falls within creative development of
new evaluation methods and metrics for artifacts, which were identified as
important design-science contributions~\cite{hevner04}.

\pagebreak
The key contributions of this paper are as follows.
\begin{itemize} 
\item We use an innovative {\em methodology} that treats \hbox{DBMSes} as experimental subjects.

\item We find that for a surprisingly large portion of queries, the plan
  chosen  by the query optimizer is not the best plan, for some cardinality
  of the underlying tables.

\item We propose a {\em predictive model} for \hbox{DBMSes} to better
  understand the factors causing suboptimality.

\item We test the seven quite specific hypotheses that arise from that
  model across a wide range of DBMSes, queries, and data.

\item We show that the hypothesized interactions are \todo{Rick:check after Sabah} supported,
  lending credibility to our particular model.

\item Through an innovative
  analysis, we track the net cumulative benefit of a succession of operators
  added to the DBMS, and show that current relational DBMSes may have
  already reached the stage where it is difficult to provide performance
  improvement via new operators, and in fact performance may degrade.

\item The predictive model and these experimental results suggest several specific
  engineering directions.

\item A subsequent analysis implies that a new approach, fundamentally
  different from that utilized over the paper forty years, may be required
  to get past a fundamental limitation uncovered in this research.
\end{itemize}

This paper takes a scientifically rigorous approach to an
area previously dominated by the engineering perspective, that of database
query optimization.  Our goal is to \hbox{understand} cost-based query optimizers
as a {\em general} class of computational artifacts and to come up with
insights and ultimately with predictive theories about how such optimizers,
again, as a general class, behave.  These theories can be used to further
improve \hbox{DBMSes} through engineering \hbox{efforts} that benefit from the fundamental
understanding that the scientific perspective can provide.

One might ask, shouldn't the task of optimizing queries be left to DBAs?  In
databases, and especially in data warehouses, the number of users writing
and running queries has been growing exponentially. This growth is aided, in
part, by the drag and drop query tools provided by the different
systems. For example, subject areas allow business users to write queries
without knowing anything about the underlying database structure. This,
coupled with constantly growing data presents new challenges for tuning. For
example, \doubleblind{the data warehouse deployed at the University of
  Arizona}{a large data warehouse we are familiar with} runs about
30,000 queries on a daily basis. 
Also, tuning a subject area or tables for one group of queries can
negatively impact the performance of other queries. Query optimization experts often take hours to tune and test existing canned
queries; the amount expended on this one system for manual query optimization
approaches \$100K/year. Therefore, we argue that it is important to
understand how existing query optimizers can be further improved, and
indeed, whether  fundamental limitation inherent in these optimizers exist
and whether such limitations are indeed already being encountered.

\begin{figure*}[t]
\centering
%originally 30pc
%\epsfig{figure=figures/plan769.pdf,width=36pc}
\includegraphics[width=30pc]{figures/subopt_flutter.pdf}
%\includegraphics[scale=0.42]{figures/subopt_flutter.pdf}
\caption{An Example of Suboptimality and Fluttering\label{fig:query769}}
\end{figure*}

We focus here on the effectiveness of query
optimization. The query optimization phase within a \hbox{DBMS} ostensibly finds the fastest
query execution plan from a potentially large set of enumerated plans, all of
which correctly compute the specified query. Occasionally the \hbox{``optimizer''}
selects a suboptimal plan, for a variety of reasons. We define {\em
  suboptimality} as the existence of a query plan that performs more
efficiently than the plan chosen by the query optimizer. From the engineering
perspective, it is of critical importance to understand the phenomenon of
suboptimality.

We study these aspects \c2j{}{across \hbox{DBMSes}} to identify the
underlying causes. We have developed a predictive causal model that
identifies four 
factors that may play a role in suboptimality. The ultimate goal is to {\em
  understand} a component within a \hbox{DBMS}, its cost-based optimizer, through the
articulation and empirical testing of a general scientific theory.

In Section~\ref{sec:related} we briefly summarize the vast amount of related
work in query optimization to establish the technical basis for our study.
The following section introduces the methodology we will follow, that of
{\em empirical generalization}~\cite{cohenbook}. We then present in
Sections~\ref{sec:model} and~\ref{sec:operationalization} a predictive,
causal model of suboptimality and state seven specific hypotheses derived
from that model. We then test these hypotheses across a number of queries,
data, and cardinalities, whose results provide strong support for the
validity of the proposed model.  These are the first scientific results that
we are aware of that apply {\em across} \hbox{DBMSes}, rather than on a
single, specific \hbox{DBMS} or on a specific algorithm. In
Section~\ref{sec:diminishing} presents a follow-on analysis that
demonstrates a fundamental limit to the number of operators that a
cost-based optimizer can support, and indicates that perhaps extant DBMSes
have hit that limit. This model has implications listed in
Section~\ref{sec:engineering} for research in engineering more efficient
\hbox{DBMSes}.

\section{Motivation}\label{sec:motivation}

Consider a simple select-project-join (SPJ) query, with a few attributes in
the SELECT clause, a few tables referenced in the FROM clause,
and a few equality predicates in the WHERE clause. This query might be an
excerpt from a more complex query, with the tables being intermediate
results.

%%%% query info: runid: 8917, dbms: pgsql, querynum:13, exp: op-1M-100q-6
%\vspace*{-2ex}
\noindent
\hspace{3ex}{\small\begin{verbatim}
        SELECT t0.id1, t0.id2, t2.id4, t1.id1 
        FROM ft_HT3 t2, ft_HT2 t1, ft_HT1 t0 
        WHERE (t2.id4=t1.id1 AND t2.id1=t0.id1)
\end{verbatim}
}

%\vspace*{-2ex}
\noindent
The optimizer generates
different plans for this query as the cardinality of the {\tt
  ft\_HT1} table varies, an experiment that we will elaborate later in depth.

The upper graph in Figure~\ref{fig:query769} represents the plans chosen by
a common DBMS as the cardinality of {\tt FT\_HT1} decreases from 2M tuples
to 10K tuples in units of 10K tuples. The \hbox{x-axis} depicts the estimated
cardinality and the y-axis  a plan chosen for an interval of
cardinalities. So Plan~P0 was chosen for 2M tuples, switching to Plan~P1 at
1,830,000 tuples, back to Plan~P0 at 1,640,000 tuples, and so on, through
the plan
sequence P0, P1, P0, P1, P2, P0, P2, and finally P3 at 40,000 tuples.

The lower graph in Figure~\ref{fig:query769} indicates the query times
executed at adjacent cardinalities when the plan changed, termed the
``query-at-cardinality'' (Q@C) time. For some transitions, the Q@C time at
the larger cardinality was also larger, as expected. But for other
transitions, emphasized in red ovals, the Q@C time at the larger cardinality
was smaller, such as the transition from plan P1 at 1,650,000 to P0 at 1,640,000
tuples. Such pairs identify suboptimal plans. For the pair at 720,000
tuples, P0 required 2.35sec whereas P1 at a larger cardinality required only
2.41sec. This query exhibits seven plan change points, two of which are suboptimal.

This query also illustrates an interesting phenomenon, in which the
optimizer returns to an {\em earlier} plan. Sometimes the
optimizer starts oscillating between two plans, sometimes even switching
back and forth when the cardinality estimate changes by a small
percentage. The example query showed returning to P0 twice and to P1 and to
P2 each once.
We call this phenomenon, in which
the query optimizer returns to a previous plan,
``query optimizer flutter,'' or simply ``flutter.''

We have found through our experiments that flutter and suboptimality are all
around us: {\em every} \hbox{DBMS} that we have examined, including several
open source \hbox{DBMSes} and several proprietary
\hbox{DBMSes}, covering much of the installed base worldwide, exhibit these
phenomena, even when optimizing very simple queries. In the Confirmatory
Experiment described in detail in Section~\ref{sec:experiments}, we started
with 6,967 query instances (a query run on a specific \hbox{DBMS}) after \doubleblind{our}{an} extensive query measurement
protocol~\cite{TTPv1,TTPv2} applied its extensive sanity checks. While about
20\%, or 1,491, of these query instances contained only one
plan, a few of the other query instances switched plans at almost every change in cardinality (we varied the
cardinality in increments of 10K tuples, a total of 200 cardinalities): see
Figure~\ref{fig:planchanges}. Slightly over half, or 3,933 query instances, exhibited
suboptimality somewhere along those 200 cardinalities; a few had
many changes to a plan that was in fact suboptimal, as indicated in
Figure~\ref{fig:suboptplanchanges}, across all four \hbox{DBMSes} considered.

\begin{figure}[h!]\centering
\caption{Cumulative percentage of queries \hbox{exhibiting} the \hbox{indicated}
  number of plan changes\label{fig:planchanges}}
\end{figure}

\begin{figure}[t]\centering
\includegraphics[width=0.60\textwidth]{figures/zi_ncpq.pdf}
\caption{Cumulative percentage of queries exhibiting the \hbox{indicated}
  number of changes to a {\em suboptimal} plan\label{fig:suboptplanchanges}}
\vspace*{-2ex}
\end{figure}

One oft-stated observation is that the role of query optimization is not to
get the {\em best} plan, but rather to get a plan that is acceptably
good. (Thus, the very term ``query optimizer'' is aspirational rather than
accurate.) Figure~\ref{fig:suboptcumulative} shows
the cumulative distribution of the relative amount of suboptimality (where
an $x$-value of 100 denotes that the query ran 100\% slower than the optimal
query, that is, twice as long). The good news is that 2,738 query instances,
or 67\% of the suboptimal queries, exhibited only a small degree of
suboptimality: less than 30\%. The challenge is that over fifth of all
queries (1,355) exhibited a significant amount of suboptimality ($\geq
30\%$). The relative slowdown can thus be quite large for some queries.

We started with 7,640 query instances, each of which is a particular query running on a
particular DBMS, cf.~Experiment 7 of Table~\ref{tab:run_stat}
in Section~\ref{sec:experiments}. After our protocol, we were left with
  6,967 query instances, of which 5,475 had at least one change point, so those are
  the ones we consider further. Of those, 1,382 had {\em no} suboptimality, so
4,093 (75\%) had some suboptimality.

1,951 (36\% of the query instances with a change point) have the higher cardinality
running at least 20\% faster than the lower cardinality. That means that
2,142 query instances (slightly over half of the suboptimal queries) were barely
suboptimal ($<$20\% slower) and about a third (1,355) were considerably
suboptimal (where the higher cardinality ran at least 30\% faster than the
lower cardinality).

\begin{figure}\centering
\includegraphics[width=0.60\textwidth]{figures/rel_diff.pdf}
\caption{Cumulative percentage of queries exhibiting the \hbox{indicated}
  percentage relative suboptimality\label{fig:suboptcumulative}}
\end{figure}

We emphasize four important points.
\begin{itemize}
\item First, we used a sophisticated query
measurement methodology that reduces the measurement variance, so that the
query plans we identify as suboptimal definitely are so.
\item Second, these
results are over four \hbox{DBMSes}, and thus, such phenomena are not dependent on
a particular implementation of cost-based optimization. Rather, they seem to
be common to {\em any} cost-based optimizer, independent of the specific
cardinality \hbox{estimation} or plan costing or plan enumeration algorithm or
implementation.
\item Third, suboptimal plans are {\em not} the result of poor
coding or of inadequate algorithms. We view query optimization in modern
\hbox{DBMSes} as an engineering marvel, especially given the complexity of the SQL
language and the requirements and expectations of \hbox{DBMS} users, who often
demand that important queries simply not get slower with a new release of
the \hbox{DBMS}. Rather, the prevalence of suboptimality observed here is a
reflection of the complexity of the task of query optimization.
\item Fourth, we wanted to understand whether a fundamental
limitation exists in the prevalent approach to query optimization utilized
by DBMSes generally, and certainly by the four DBMSes that we studies.
\end{itemize}

This is why we needed a new methodology. We want to understand
cost-based optimization deeply. This means that we need to go beyond the
examination of a single query optimizer, as is done in almost every paper
over the forty-year history of query optimization research, to study multiple instances of
that optimization architecture, in an effort to achieve generalizable
results.

Thus, in this paper,
we articulate a predictive causal model for how
and in what circumstances suboptimality arises and provide compelling
evidence via \hbox{hypothesis} testing that this model accurately characterizes the
\hbox{behavior} of query optimizers in general. This is what is meant by ``empirical
generalization''~\cite{cohenbook} and why it is needed to answer such questions. In
  the long history of research in database query optimization, or even of
  databases in general, our model and its hypothesis tests are the first
  predictive results that we are aware of that apply {\em across}
  \hbox{DBMSes}, rather than on a single, specific \hbox{DBMS} or on a
  specific algorithm. Such a DBMS-agnostic, though paradigmatic, causal model can
\hbox{provide} guidance to the community about where fundamental research is needed
and to the \hbox{DBMS} engineers about where to focus their efforts.

\section{Related Work}\label{sec:related}

\todo{Rick: For Science article, add history: IMS->System R,
  procedural->declarative, Selinger's query opt algorithm enabling rise of
  relational model}
\todo{Rick: For TODS, shorten history, probably remove SQL cites, just one
  sentence about central role of Selinger}
SQL has emerged as the {\em de facto} and {\em de jure} standard relational
database query and update language. It emerged as an ANSI and ISO standard
in 1986--7, with a major extension and refinement as SQL-92~\cite{Melton93} 
and an even larger extension as SQL:1999~\cite{Melton03} and refinement as
SQL:2008~\cite{SQL2008} and SQL:2011~\cite{Melton11}.

There has been extensive work in query optimization over the last 40
years~\cite{Ioannidis96,Jarke84}, during which a particular quite effective
paradigm had taken hold, in both open-source and proprietary DBMSes. In this
over-arching paradigm, query optimization and evaluation
proceeds in several general steps~\cite{Ramakrishnan03}. First, the SQL query is 
translated into alternative query evaluation plans based on the 
relational algebra via {\em query enumeration}. The cost
of each plan is then estimated and the plan with the lowest estimated cost
is chosen. These steps comprise query optimization, specifically {\em
cost-based query optimization}~\cite{Selinger}. The selected query plan is
then evaluated by the query execution engine which implements a set of
physical operators, often several for each logical operator such as
join~\cite{Graefe93}.

An influential survey~\cite{Chaudhuri98} does a superb job of
capturing the major themes that have pursued in the hundreds of articles
published on this general topic. Chaudhuri reviews the many techniques and
approaches that have been developed to represent the query plans, to
enumerate equivalent query plans, to handle some of the more complex lexical
constructs of SQL, to statistically summarize the base data, and to compute the
cost of evaluation plans. He also mentions some of the theoretical work
(which is much less prevalent) to understand the computational complexity of
these algorithms. Most of this research may be classified as adopting an
engineering perspective: how can we architect a query optimizer ``where 
(1)~the search space includes plans that have {\em low cost} (2)~the costing
technique is {\em accurate} (3)~the enumeration algorithm is {\em
  efficient}. Each of these three tasks is nontrivial and that is why
building a good optimizer is an enormous undertaking.'' \cite[page~35]{Chaudhuri98}

To determine the best query access plan, the cost model estimates the
execution time of each plan. There is a vibrant literature on this
subject~\hbox{\cite{Ioannidis03,Mannino88}}, including proposals for
histograms, sampling, and parametric methods. Again, most of these papers
are engineering studies, providing new techniques that improve on the
state-of-the-art through increased accuracy or performance. There have also
been a few mathematical results, such as ``the task of estimating distinct
values is {\em provably} error prone, i.e., for any estimation scheme, there
exists a database where the error is significant''~\cite{Chaudhuri98}.

An optimizer for a language like SQL must contend with a huge search space of
complex queries. Its first objective must be {\em correctness}: that the
resulting query evaluation plan produce the correct result for the
query. This objective must be ensured both by the initial SQL-to-relational
algebra translator and by the subsequent query enumerator. A secondary but clearly very important objective is {\em efficiency}; after
all, that is the raison d'etre for this phase. As is well known and has been
alluded to already, the name
for this phase is an exaggeration, as existing optimizers do not produce
provably optimal plans. That said, the
query optimizers of prominent \hbox{DBMSes} generally do a superb job of producing
the best query evaluation plan for most queries. This performance is the
result of a fruitful collaboration between the research community and
developers.

Early investigation of plan suboptimality resulted in approaches such as
dynamic query-reoptimization~\cite{Avnur,kabra98}, which exploit more
accurate runtime statistics that appear while a query is being executed, to
steer in-flight plan reoptimization. The very presence of such a radical
change to the normal optimize-execute sequence indicates that plan
suboptimality was of interest to some researchers.

However, even with great effort over decades, optimizers as a general class
are still poorly understood. As has been observed, ``query optimization has
acquired the dubious reputation of being something of a black
art''~\cite{Babcock05}. DeWitt has gone farther, stating that ``query
optimizers [do] a terrible job of producing reliable, good plans [for
  complex queries] without a lot of hand
tuning''~\cite[page~59]{winslett02}. And as we will see, suboptimality may
occur in sophisticated query optimizers even when considering only simple queries.

While this paper does not provide direct solutions to address
suboptimality, we envision that by following up on the implications of our proposed predictive model
for suboptimality, engineering practice, such as dynamic reoptimization just
mentioned, may benefit. We elaborate on this subject in
Sections~\ref{sec:root} and \ref{sec:engineering}, where we discuss the engineering implications
of our causal model.

\section{A Model of Suboptimality}\label{sec:model}

The purpose of query optimization is to generate optimal plans.  So
why would suboptimality occur in the first place? Query optimizers are
highly complex, comprised of tens or hundreds of thousands of lines of code. There
are several reasons for this complexity. First, an optimizer must
contend with the richness of the SQL language, whose definition
requires about 2000 pages~\cite{SQL2008}, with a multiple of
linguistic features. Second, an optimizer must contend with the
\hbox{richness} of the physical operators available to it. \hbox{DBMSes} have a
range of algorithms available to evaluate each of many algebraic
operators. Third, the optimizer must contend with an exponential
number of query evaluation plans. Kabra and DeWitt~\cite{kabra98}
identify several other sources of complexity: inaccurate statistics on
the underlying tables and insufficient information about the runtime
system: ``amount of available resources (especially memory), the load
on the system, and the values of host language
variables.''~(page~106). They also mention user-defined data types,
methods, and operators allowed by the newer object-relational
systems~\cite{Melton03}. Thus, the task of optimization is very
complex, with the result that the optimizers themselves consist of a
collection of ``components,'' that is, the rules or heuristics that it
uses during optimization, with each of these components being itself
complex.

We wish to understand the causal factors of suboptimality, through a
predictive model that explicitly states the interactions between these
causal factors. We test this model through experiments over tens\shorten{14,700} of thousand
of queries and hundreds of thousands of query executions\shorten{2.4M}, showing that there
is strong support for this model. We then extract engineering implications
from the model, suggestions for the most productive places to look to reduce
suboptimality and thus to improve existing query optimizers.

\begin{figure*}[tb]
\centering
%originally 30pc
%\epsfig{figure=figures/model.pdf,width=30pc}
\includegraphics[width=30pc]{figures/subopt_model.pdf}
\caption{Predictive Model of Suboptimality\label{fig:model}}
\vspace*{3ex}
\end{figure*}

Here we examine the hypothesized influence that each independent
variable will have on the one dependent variable, query suboptimality (with
some of the the influences mediated by one of the constructs). In
the next section we will operationalize these variables, explaining how
each is controlled or measured.

\subsection{Constructs in the Model}

The model concerns five general
constructs that we \hbox{hypothesize} will play a role in suboptimality: optimizer
complexity, schema complexity, query complexity, data complexity, and plan space complexity.
Two constructs include several specific variables that contribute to that
construct as a whole. Our model distills many of the widely-held assumptions
about query optimization. Our contribution is the specific \hbox{structure} of
the model and the specific operationalization of the
\hbox{factors} included in the model.

Our theory, which we will investigate in some detail and which is the
basis for our predictive model, is that suboptimality is due in part
to the complexity of the optimizer and concomitant unanticipated
interactions between the different components used within the 
optimizer with various sources of complexity in producing a good (hopefully
optimal) plan.  We argue that with the proliferation of concerns and
variability that an optimizer must contend with, it is extremely
difficult to ensure that for an arbitrary query that the optimizer
will {\em always} pick the right plan. There are many sources of complexity
that may challenge one or more components of the optimizer; our theory
implicates specific sources as contributing to observed
suboptimal plans.

The model depicted in Figure~\ref{fig:model} suggests specific factors that may
impact the prevalence of suboptimality. This model has one dependent
variable, {\em suboptimality}, on the far right. We observe this dependent
variable in our experiments by determining whether each particular query,
run on a particular \hbox{DBMS} and using a particular schema,
is suboptimal at any cardinality of the input table. In Section~\ref{sec:operationalization}, we
delve into the details of how we operationalize this and the other variables
we now examine.

\subsection{Variables in the Model}
The model has several independent
variables which influence suboptimality.

For the construct of {\em optimizer
complexity} we have one independent variable, ``\hbox{Number} of operators
available (in the \hbox{DBMS}).'' We can manipulate this variable by our choice of
\hbox{DBMS}: each \hbox{DBMS} has a set of operators available to its query
evaluator and available to the query optimizer to use with in query plans.

We have one variable for the construct of {\em schema complexity}:
``Presence of secondary indexes.'' If true, that means that every
non-primary-key attribute, for the table for which we vary the cardinality
(termed the {\em variable table}) as well as for the other three tables
(termed the {\em fixed
tables}), has a secondary index associated with it. The rationale is that the
presence of secondary indexes expands the number of possible plans: each
predicate in the query can often be mapped to one or more operators
utilizing that index.

For the construct of {\em query complexity} we have identified four
variables: ``Number of correlation names'', ``Presence of aggregate'',
``Presence of primary key attribute'', and ``Presence of subquery''. For each independent
variable, we have interventional control in our experiments, in that we can
manipulate the values of these variables through the construction of the
actual query to be optimized. The first is the number of correlation names
defined in the FROM clauses. The second is whether a single aggregate operator
(that is, either {\tt Max} or {\tt Avg}) appears in the SELECT clause.
The third
is whether a primary key attribute appears in at least one predicate in the
WHERE clause. The fourth is whether a single subquery appears in the
WHERE clause; that subquery evaluates to a single value that is
equality-compared with an attribute, in the where clause. The rationale is
that each factor may expand the number of plans or otherwise render the search for
the optimal plan more complex.

We have one variable for the construct of {\em data complexity}: that of
``Percentage of skewed data''.  Skew is generally defined as how values are
distributed with a column of a table. We refine this to ``how many duplicate
values are present,'' with zero skew meaning that there are no duplicate
values present and 100\% skew implies that there is but one value in the
entire column. The presence of skew complicates query time estimation, which
in turn complicates the search for the optimal plan.

In the middle of the figure is the construct of {\em plan space
complexity}. Given a particular query, its
complexity will impact the total number of candidate plans considered
by the optimizer. This variable is not directly
observable, again, especially for proprietary \hbox{systems}. However, we {\em
can} measure the number of plans actually generated by the optimizer
when presented with different cardinalities of the underlying tables.
We term this set of plans the ``effective plan
space'' and term the number of such plans, the cardinality of the
effective plan space, or ``{\em CEPS}''. Similarly, we cannot directly observe
the cost model utilized by the optimizer for the operators it considers, but
can classify the cost model of each operator as continuous (smoothly varying
with cardinality of its input(s) and without jumps) or discontinuous
(sometimes producing observable jumps in the predicted
cost), and can thus count ``the number of operators with
discontinuity.'' (Such operators have also been termed
``nonlinear''~\cite{Hulgen03}, though we focus on a more specific aspect of
discontinuity.) Finally, the variable ``Number of repeats'' {\em is}
directly measurable as the number
of times a plan is reused across the cardinality of the variable table. For
example, if the sequence of plans for a query as the cardinality increases
is $A ~ B ~ B ~ B ~ C ~ C ~ A ~ B ~ B~ B ~ C ~ B ~ C ~ C ~ C ~ B$, then the
``Number of repeats'' will be six: removing sequential duplicates results in $A ~ B ~ C
~ A ~ B ~ C ~ B ~ C ~ B$, with the last six distinct plans being duplicates
of the first three. Interestingly, CEPS plus the number of repeats gives you the
number of plans in the sequence, again, after removing sequential
duplicates, nine plans in this particular case. Thus, we associate with the latent construct
of plan space complexity three measurable (that is, indirect) variables. Our
  model stipulates that the construct, and thus the three associated with
  this construct, intervene between the constructs of optimizer complexity,
  schema complexity and query complexity and the construct of suboptimality.

Plan space complexity is an {\em intervening} construct, in that it is dependent
on some of the constructs on its left but is observable and thus exerts influence on the dependent
variable on its right. We can observe these variables within experiments and
indirectly influence their value but
cannot directly intervene to specify their value. For example, we can influence the CEPS through
manipulating the values for the independent variables of optimizer and query
complexity, but cannot directly set a value for CEPS within an
experiment.

\subsection{Interactions in the Model}\label{sec:interactions}
Given these six constructs and eleven specific variables depicted in
Figure~\ref{fig:model}, let's now examine the causal relationships between
these variables, depicted as directed arrow between variables (a line
originating or ending at a construct is interpreted as originating or ending at all
variables in that construct). Such relationships are specific {\em interactions} between
the the constructs (in particular, between their associated variables), as hypothesized by
this predictive causal model.

One causal factor of this model is the optimizer complexity. We hypothesize
that optimizer complexity has influence over suboptimality indirectly, via
plan space complexity. We hypothesize that an optimizer with a larger number
of available operators will generate more plans and hence increase plan
space complexity.

\vspace{0.6em}\noindent
{\bf Hypothesis 1}: Number of operators available will be positively correlated with, (a)~Number of
repeats, (b)~CEPS, and (c)~Number of operators with discontinuity.

\vspace{0.6em}We now turn to query complexity, a construct associated with four
independent variables. As with the optimizer complexity construct, we
include in our model an
indirect interaction through plan space complexity.

A higher value of each of these specific variables implies a more complex
query.  We expect a strong relationship between Number of correlation names
to CEPS (because it is well-known that the number of potential join
combinations is exponential to number of correlation names), to number of
repeats (number of repeats could partially track CEPS), and number of
operators with discontinuity (for the same reason).

We also expect a positive relationship between presence of aggregate (as
that will definitely add at least one operator to the query plan), presence
of primary key attribute, and
presence of subquery.

\vspace{0.6em}\noindent
{\bf Hypothesis 2}: Number of correlation names will be
strongly correlated with (a)~Number of repeats, (b)~CEPS, and (c)~Number of
operators with discontinuity. Presence of aggregate will be positively
correlated those those three variables (correlations d--f) and similarly
with Presence of primary key attribute
(correlation g--i) and Presence of subquery (correlations j--l).

\vspace{0.6em}
We don't feel that skewed data (in the ``data complexity''
construct) will impact plan space
complexity, but rather that it will negatively impact accuracy of plan cost
estimation, and thus increase suboptimality.

\vspace{0.6em}\noindent
{\bf Hypothesis 3}: Percentage of skewed data will be negatively correlated with Suboptimality.

\vspace{0.6em}
Let's now turn to the plan space complexity construct. We hypothesize a positive
correlation between the two variables associated with this construct. As the
number of plans considered by the optimizer increases, so should CEPS, which
could increase the number of operators with discontinuity that is observed.

The model also has two interactions within plan space
complexity, between 
CEPS and number of operators with discontinuity variables and between number of repeats.

We hypothesize that greater plan space complexity should make it more
difficult to optimize the query.  It is important to note that the
occurrences of suboptimality is not necessarily dependent on the presence
of a discontinuous plan operator. However, we predict that when
discontinuous plan operators appear in candidate plans, they may introduce
complexity to query optimization, especially at the cardinality where
discontinuity is possible. This is because the performance of such operators
is sensitive to the input size in a rather complex way. Any inaccuracy in
the estimation of plan statistics may lead the optimizer to select a suboptimal
plan.

\vspace{0.6em}\noindent
{\bf Hypothesis 4}: Plan space complexity (CEPS) and (a)~Number of operators
with discontinuity and (b) Number of repeats will be positively correlated.

\vspace{0.6em}
Similarly, a plan chosen by the optimizer once will be reconsidered at a
later cardinality and perhaps chosen, following
the past experience of selecting and using that plan among many other plans.
Even if many different plans are already used, 
for the same reason the optimizer may revisit a pool of 
the previously used plans and choose one of them than to explore other new plans, 
given the complexity of plan cost estimation. 
Then the larger CEPS, the more times the optimizer repeats using the same plans. 
We thus hypothesize that the number of repeats has a positive correlation with CEPS.

\vspace{0.6em}\noindent
{\bf Hypothesis 5}: The Number of repeats will positively correlate with CEPS.

\vspace{0.6em}
The schema complexity construct consists of the variable ``Presence of
secondary indexes.'' Such indexes provide opportunities for the optimizer to
consider more candidate plans that use these indexes, due to the additional
query evaluation operators now applicable. Those additional plans enable the optimizer to
possibly do a better job, while also adding complexity to the optimization
process. We hypothesize that this factor has a more complex role in the
model, serving as a {\em moderator} of two interactions previously
introduced. We hypothesize that the overall effect of the secondary indexes
is to increase the strength of the interaction between query
complexity to plan space complexity.

\vspace{0.6em}\noindent
{\bf Hypothesis 6}: Presence of secondary indexes will strengthen
the correlations between the Query Complexity construct, specifically, the
variables Number of
correlation names, Presence of aggregate, Presence of primary key
attribute, and Presence of subquery, and
the Plan Space Complexity construct, specifically, variables Number of repeats, CEPS, and Number
of operators with discontinuity (correlations a--l listed above in
Hypothesis 2).

\vspace{0.6em}
Finally, we hypothesize that queries with a high plan space complexity
present challenges to the query optimizer, and thus increase the chance that
the query optimizer picks a suboptimal plan. This is a fairly direct
connection for CEPS and Number of operators with discontinuity. For the
Number of repeats variable, a query exhibiting a high number of repeats has
more opportunities for suboptimality at some cardinality, just because the
number of repeats is an indication that the query optimizer is struggling.

\vspace{0.6em}\noindent
{\bf Hypothesis 7}: Suboptimality will be positively correlated with Plan
space complexity, that is, (a)~Number of repeats, (b)~CEPS, and (c)~Number
of operators with discontinuity.

\vspace{0.6em}
We have just {\em described} how
suboptimality might arise, through a theory and its elaborated causal model,
which implies seven specific hypotheses.
We now need to move to {\em prediction}.
How might we test such a model?

The first step to test this model is to {\em operationalize} each
variable. In the next section we describe explicitly how each variable is
defined. For the independent variables, we must be able to intervene, that is,
set their values before the experimental test commences. For the latent and
dependent variables, we need to be able to measure their values during each experiment.

\section{Variable Operationalization}\label{sec:operationalization}

In this section we specify how we operationalized each of the seven
variables in the model.  Recall that each independent variable is a property
of a \hbox{DBMS} (number of operators available), of the schema (presence
  of secondary indexes), of a query (number of correlation names,
presence of an aggregate, presence of primary key attribute, and
presence of subquery), of the data (percentage of skewed data), and of
the plan space (number of repeats, CEPS, and number of discontinuous
operators). There is 
one dependent variable of our model (suboptimality).

It is important to note that all manipulation must be done {\em
  outside} the \hbox{DBMS}. For proprietary systems we do not have access to the
internal code. We do not know ({\em cannot} know) all the plans that were
considered, nor the details of how the plans are selected. But such access
is not needed; indeed, to be able to study a phenomenon across many \hbox{DBMSes},
such access is not practical. But by designing the experiment to examine
the plans that each \hbox{DBMS} actually produces, and thus to examine phenomena
that can be externally visible, we can obtain valuable insights into general
classes of computational artifacts. Our ultimate goal is to make statements
that hold across cost-based optimizers in general, thereby moving up the
$y$-axis of Figure~\ref{fig:empirical}.

\begin{figure*}[t]
\centering
\includegraphics[width=0.60\textwidth]{figures/understanding.pdf}%0.40
\caption{Empirical Generalization\label{fig:empirical}}
\end{figure*}

\subsection{Optimizer Complexity}
By ``number of operators available in the \hbox{DBMS}'' we mean the \hbox{number} of operators
available for selection, projection, join and aggregate functions. Our
experiments intervene on this variable by selecting a particular \hbox{DBMS} on
which to evaluate each query. Across the available \hbox{DBMSes}, as the number of
operators available increases, the complexity of the optimizer increases
because it has to choose between more operators.

The {\tt EXPLAIN PLAN} facility specifies the operators(s) \hbox{employed} in that
plan. For each \hbox{DBMS}, we collect all the unique operators from all the plans
and count the number of these distinct operators, each used by for least one
query at at least one cardinality by that \hbox{DBMS}. The number of operators used
by a \hbox{DBMS} ranged from 8 to 53 across all the queries and data sets
used in the Exhaustive, Exhaustive with Keys, and Exploratory Experiments,
to be discussed in Section~\ref{sec:experiments}.

\subsection{Schema Complexity}
The presence of secondary indexes is easy to operationalize. We generate two
databases, one without any secondary indexes and one with a key specified
for each table on the non-key (that is, other than the first) attribute.

\subsection{Query Complexity}\label{sec:querycomplexity}
Query complexity is also relatively easy to operationalize. We randomly generate queries,
such as the example presented in Section~\ref{sec:motivation}. Each query
is a select-project-join-aggregate query, with a few attributes in the
SELECT clause, a few tables referenced in the FROM clause, a few
equality predicates in the WHERE clause, and zero or one aggregate
functions in the SELECT clause. As such, some of the complexities
mentioned by Kabra and \hbox{DeWitt~\cite{kabra98},} such as
user-defined data types, methods, and operators, are not
considered. Concerning presence of primary key attribute, if this independent
variable was set to 1, we ensured that there was at least one primary key
attribute in one of the comparisons in the WHERE condition.

The queries were generated
by a simple algorithm\shorten{, the pseudo-code for which is shown in
Figure~\ref{alg:querygen}}. 
\shorten{The query generator has four main components, which are {\tt
  generateSelectClause()}, {\tt generateFromClause()}, {\tt
  generateWhereClause()}, {\tt generateGroupByClause()}. , and {\tt buildAggregateStatement()}.}
The SELECT clause will contain from one to four
attributes, hence, an average of 2.5 attributes. The number of
correlation names
in the FROM clause varied from one to four, with duplication of tables
allowed (duplicate table names within a FROM clause implies a self-join).
In the queries that were generated, from one to four unique tables were
mentioned in the FROM clause. Somewhat fewer tables were mentioned than the
number of correlation names, as the presence of self-joins reduced the
number of unique tables referenced by the queries.

For the query in Section~\ref{sec:motivation}, the number of correlation
names is 3, the presence of an aggregate
is false (0), the presence of primary key attribute is false (0),
and the presence of subqueries is false (0).

%\pagebreak
\shorten{In {\tt generateWhereClause()}, in which all the joins will appear,
  we}We ensure that Cartesian products are
eliminated\c2j{.}{\shorten{({\tt cartesianPossible="false"})}}  We do this by
connecting all the correlation names that appear in the FROM statement via
equi-joins on random attributes. The comparisons are all
equality operators. To ensure that the queries are as simple as possible, we
do not include any additional predicates in the WHERE clause.  \c2j{}{This
  is realized by setting the attributes {\tt maxIsAbsolute} to {\tt true}
  and {\tt complexUsePercentage} to {\tt 100}.  Basically, ``complex''
  predicates eliminate the Cartesian product, and by setting complex
  predicates as ``absolute,'' no additional predicates are included except
  for those which are necessary for eliminating Cartesian product.} Also
for simplicity, we include neither disjunctions nor negations.

\shorten{In the \todo{Correct name: generateGroupByClause}, where the aggregate
functions and
grouping occurs, the presence or absence of a group by clause is
recorded. We expect the
complexity of the query to increase when the first aggregate function
is introduced. (We don't consider in this model whether
additional aggregate functions in the same SELECT clause add
complexity}

The presence of subquery was 0 or 1, with 0 indicating no
subquery. For each query needing a subquery, we picked a separate generated query and
rendered it as a subquery to replace an attribute in the where clause of the
original generated query. As an example when this variable was one, we
started with query qs1-2, shown here.

\begin{verbatim}
SELECT t0.id2, SUM(t1.id3)
FROM ft_HT2 t0, ft_HT2 t1
WHERE (t0.id2=t1.id1)
GROUP BY t0.id2
\end{verbatim}

We then generated another simple select-project (SP) 
query at random concerning the variable table 
and simply replaced one side (in this case, the right side) with the generated entire 
query as a subquery, to produce this final query.

\begin{verbatim}
SELECT t0.id2, SUM(t1.id3)
FROM ft_HT2 t0, ft_HT2 t1
WHERE (t0.id2 IN (SELECT t2.id3 FROM ft_HT1 t2)) 
GROUP BY t0.id2
\end{verbatim}
We thus have a maximum of only one level of nesting.

\subsection{Data Complexity}\label{sec:datacomplexity}
This independent construct has one variable: percentage of skewed data.

Skew has a very specific definition in statistics, 
involving elongating the left or right tail of a distribution, 
thereby moving the mean left or right of the median 
(in a symmetric distribution the mean = median = mode).

But we start with a distribution without a tail: 
the uniform distribution: the values from 1 to 2 million (2M). 
We consider this to be a skew of 0 (no skew). 
At the other end of the spectrum is one in which all the values are identical, 
or a skew of 1.0.

We define the skew quite simply: ``the reciprocal of the number of distinct
values,'' so $0 < \hbox{\em skew} \leq 1$.
For 2M distinct values, the skew would be $\frac{1}{2M}= 0.000005$, which is practically 0. 
For exactly one distinct value, the skew would be 1.0. For two distinct values, 
the skew would be 0.5. For ten distinct values, the skew would be 0.1.

We can generate the table of 2M rows by generating values sequentially from
1 to the number of distinct values.  This creates a ``span of values.'' We
repeat this for the second span if necessary, and on and on, until we have
2M values in all.

When varying the cardinality, we remove 10K values from the variable table
and then copy those tuples to a new table to ensure that every page is as
full as possible (that is, 100\% load factor).  This gives us a table of
1.99K tuples. (We then get a query plan for this table.)  We repeat this
removal process until the final cardinality reaches 10K.

The way we effect the 10K removal is as follows. 
The key idea is to remove individual spans until we've deleted 10K
values. Since we don't touch the remaining spans, the number of distinct
values does not change, and so the skew remains constant.

Let's first illustrate with a skew factor of 1.0. 
This translates to exactly 1 value for the entire table, or 2M spans.
To shorten, we remove 10K spans, equal to 10K values. 
This removal keeps the skew at 1.0, due to the unique values in the
remaining spans.
At the end, 10K spans, each with a single value, remain.

What about a skew factor of 0.1? 
This skew factor translates to 10 distinct values. 
So we generate 200K spans, each with 10 values. 
To shorten, we delete 1,000 spans, which is equivalent to removing 10K values. 
As before, this does not change the skew.
At the end, the table will have 1,000 spans with 10K values, retaining a
skew factor of 0.1.

Let's see an example at the other end of the spectrum, of a skew factor of
1/2M.  This skew factor, which is almost close to 0, gets translated to 2M
distinct values.  So we generate a single span of 2M values.  To shorten the
table, it makes no sense to remove this single span in its entirety. But we
can remove 10K values from this single span.  Note that this changes the
skew to 1/1.99M, then eventually to 1/10K, which is still very close to zero
(the skew has changed from .0000005 to .00001).

We use five skewness values: 0.000005 (which we term only slightly
misleadingly as ``no skew''), 0.001, 0.1, 0.5, and 1.0,
for this independent variable.

\subsection{Plan Space Complexity}\label{sec:discontinuity}

This explanatory construct includes three variables.

As discussed in Section~\ref{sec:model}, the ``cardinality of the effective plan
space'' (CEPS) is the number of plans selected as optimal for that query
being evaluated on one of the 200 cardinalities for the variable table. It
is ``effective'' because it was chosen, as opposed to all of the
plans that were considered but not chosen. (Recall that for proprietary
\hbox{DBMSes}, we do not have access to such plans.) Note that we count only
distinct plans. As we saw in Figure~\ref{fig:query769}, fluttering queries
return to a previous plan. This particular query has a CEPS of 4.

The number of repeats is just the number of times a plan associated with a
smaller cardinality is repeated, after removing sequential
duplicates.

We also wanted to evaluate the contribution of cost model non-linearity to
suboptimality. We found that it is possible to assemble, from outside the
\hbox{DBMS}, an approximation of the cost formula for that operator, and thus
directly observe whether it is non-linear. Specifically, the result of SQL's
{\tt EXPLAIN PLAN} (a result that is particular to each \hbox{DBMS}) includes the
{\em estimated cost} of each stage of the plan. That cost (estimate) is a
dependent variable, one that can be observed.

To operationalize the ``number of operators with discontinuity'' we
classify each \hbox{DBMS} operator as either continuous or discontinuous, to be
defined shortly.  We then examine the queries
to identify the plan change points, which provide the {\em effective plan
  space}, a set of plans chosen for one or
more cardinalities of the variable table. For each distinct query plan in the
effective plan space, we
count the number of discontinuous plan operators that appear in that plan. Some of
these plans may contribute no such operators, some may contribute one such
operator, and some may contribute several such operators. We then sum the
counts of the distinct plans in the effective plan space, yielding an
integer as the operationalization of ``number of discontinuous operators''
for each query.  This count per query varied from 0 to 85 for the
queries we tested in the confirmatory analysis.

To classify each operator as continuous or discontinuous, we use the data
from the Exhaustive Experiment. We
used a small subset of queries to be used later in testing the model, to
attempt to observe the same operators as encountered in that
study. We
collected all query plans at all possible cardinalities (200 in all) and
looked for ``jumps,'' that is, when the cost model for an operator in the
plan exhibited discontinuity. Jumps are determined from the estimated cost
extracted from each plan operator. We provide an example of a jump below. If
a jump is observed, we classify that particular plan operator as
discontinuous.

A {\em jump} is a pair of close cardinalities (in our case, separated by 10K
tuples) in which the query optimizer's cost model is {\em discontinuous},
that is, does not smoothly increase from the lower cardinality to the upper
cardinality. To identify jumps in the first step, we examine the slopes
between each pair of adjacent cardinalities (again, separated by 10K
tuples). We expect that the slope (that is, the first derivative) is
well-behaved and thus does not change much for adjacent cardinalities. 
A jump thus indicates a large deviation in the second derivative of the 
cost model across the two cardinalities.

The Exhaustive experiment gathers the cost model for each operator in each
plan for each cardinality (in our case, for cardinalities ranging from 10K
to 2M in steps of 10K, or 200 points).  Figure~\ref{fig:discontinuity}
presents for a single query\shorten{({\em query4}, exhaustive-10q-1,
  oracle)}, the {\em cost} of each of the hash-join operators utilized in
each plan at a particular cardinality. Each distinct point type depicts an
individual hash-join operator. Note that when two identical query plans
appear at two different cardinalities, we consider all the plan operators
found at the same position within these two plans to be identical. This
figure shows only a small portion of the 200 cardinalities, and is typical
(we generally observed jumps at low cardinalities for these particular queries and
relation sizes).  As shown by this figure, the hash-join operator
represented by `+' has two discontinuous jumps, both appearing below
cardinality 150K. The other hash-join operator represented by $\times$ has one
discontinuous jump.  Between the jumps, the behavior is linear. It is our
guess that
such a jump is caused by the transition between one pass of a disk-based hashing
technique to two passes (and indeed for one of the operators, perhaps the one
lower in the operator tree, the transition to three passes, hence, the two
jumps). That said, all
that matters for our predictive model is that operators that experience
such discontinuity in their cost models present opportunities for
suboptimality.

%\begin{figure}[t]\centering
\begin{figure}[t]
\centering
%\includegraphics[width=0.70\textwidth]{figures/discontinuity_400K.pdf}%0.40
\includegraphics[width=0.60\textwidth]{figures/discontinuity_500K.pdf}%0.40
\caption{An Example of Discontinuous Plan \hbox{Operators (Hash-Join)}}
\label{fig:discontinuity}
\end{figure}

\shorten{It is interesting that jumps occurred only at smaller cardinalities. This
might depend on the strategy of how the \hbox{DBMS} utilizes the main memory
buffer.  Assume that the \hbox{DBMS} attempts to use as little memory as possible.
One possibility is that when the cardinality is low, memory allocation is
done in small chunks. Hence as the cardinality grows, more passes are
required. However, when tables get larger, bigger
chunks of memory would be allocated, requiring the operator to take longer
to fill up the available memory. We again emphasize that our focus is on predicting
suboptimality. Perhaps memory allocation has a role in operator
discontinuity, which our model predicts to have a role in query
suboptimality.
}
To identify such jumps, we examine the second derivative (the change in the
first derivative, that is, the change in the slope). A jump will be indicated by a larger than normal second
derivative. By identifying a sudden change in the second derivative, we can
effectively spot the cardinality at which an operator in a plan becomes
discontinuous.

Formally, we compute the slope ({\em S}) of the {\em estimated} cost
(of the cost model, formalized as $C_{\hbox{\small\em card}}$,
where {\em card} is the input
cardinality) between each pair of adjacent cardinalities as
$S_{\hbox{\small\em card}} = (C_{\hbox{\small\em card} + 10K} -
C_{\hbox{\small\em card}}) / 10K$~.
This computes a series of slope values.
We then compute the standard deviation of all the slope
values.  For example, examining Figure~\ref{fig:discontinuity}, the slopes
are small except at three places, one for the green operator and two for the
red operator. By identifying the slope values that are greater than one
standard deviation over the average value, the discontinuous operators can
be identified: A ``discontinuous operator'' is one for which a jump is
observed in that operator in one of the plans for at least one of the input
queries.

\subsection{Suboptimality}\label{sec:suboptimality}
We now consider the one dependent variable.
How might suboptimality be observed? We have developed a system, \azdb, that
allows us to perform experiments to study this phenomenon of suboptimality.
\hbox{\azdb} submits queries to the \hbox{DBMS}, while varying
the cardinality of one of the tables,
requesting in each case the evaluation plan chosen by the \hbox{DBMS}.
This is done using the
{\tt EXPLAIN} SQL facility available in modern \hbox{DBMSes}. (The Picasso system
also used this facility to visualize the plan space chosen by a \hbox{DBMS}
optimizer~\cite{harish07,Haritsa10}.) We can then compare
the performance (execution time) of various plans for the query, to identify those
situations when a suboptimal plan was chosen, when in fact there was a
different plan that was semantically equivalent to the chosen plan (that is,
yielded the identical result) but which ran faster.

We modify the cardinality to produce multiple execution
plans for a given query. For one of the \hbox{DBMSes}, we can modify the stored
table statistics directly. For the other \hbox{DBMSes}, we had to do so indirectly,
by varying the size of the table and running the optimizer on tables of
different size. As we vary the cardinality, we
collects all the plans that the optimizer felt were appropriate for that
query.

Our definition of suboptimality assumes that
actual execution time for any query plan is {\em monotonic non-decreasing}, that is,
unchanged or increasing as the cardinality increases.  The intuitive justification is that at the higher
cardinality, the plan {\em has} to do more work, in terms of CPU time and/or
I/O time, to process the greater number of tuples. (For an operator having a discontinuous cost model, as the
anticipated input cardinality grows, there will be {\em jumps}, in which the
predicted cost is temporarily much greater. Section~\ref{sec:discontinuity}
explains further how we specifically
operationalize this property. Also note that we don't consider SQL
operators such as {\tt EXCEPT} that are not monotonic.)
We formalize this property as follows.

\vspace{1em}\noindent
{\bf Definition:} {\em Strict Monotonicity}: given a query $Q$ and an actual
cardinality $a$,

\quad\quad\quad\quad\quad$\forall p \in \hbox{\em plans} (p) \;\forall c > a \;( \hbox{\em time} (p,
c) \geq \hbox{\em time} (p,a) )$

\vspace{1em}\noindent
where {\em plans}($p$) is the set of plans generated by the optimizer and
{\em time}($p$, $c$) is the execution time of plan $p$ on the data set with
the varying table at cardinality $c$.\hfill$\blacksquare$

\vspace{1em}
\noindent
Note that the comparison is with the same plan $p$, occurring at higher
cardinalities.

To test this assumption, we ran an experiment that we term
``Monotonicity.'' This experiment considered 60 queries, chosen from the
pool of queries
generated for testing suboptimality, and timed them for cardinalities from 10K
to 2M tuples, in steps of 10K tuples (hence, we used 200 cardinalities), for
each \hbox{DBMS} (for one \hbox{DBMS} that was very slow, we started with 30K tuples, as
discussed in Section~\ref{sec:experiments}). We varied the cardinality of the variable table by starting with
the maximum size, running the queries, then deleting 10K tuples and repeating.
We performed an ``{\tt ANALYZE TABLE}'' function to force
the \hbox{DBMS} to update the table's statistics to be accurate before actually
evaluating the query (we did this for all experiments).
We expected that as the cardinality decreased,
the runtime would also monotonically decrease. However, due to the variance in query
time measurement observed even when the cardinality was identical, we encountered
spurious violations.

Assuming a normal distribution for our time measurements, 95\% of 
the distribution falls between $-2*\sigma$ and $+2*\sigma$. 
Therefore, to statistically
infer with a 95\% confidence interval that a violation occurred, 
we relaxed our definition of monotonicity to the following.

\vspace{1em}\noindent
{\bf Definition:} {\em Non-Strict Monotonicity}: given a query $Q$, an actual
cardinality $C$, and the standard deviation of the query executions for
cardinality C as $\sigma$, $Q$ is non-strict monotonic if $\forall p \in
\hbox{\em plans} (p) \;\forall c^{\prime} > C $,

\quad\quad\quad\quad\quad$\hbox{\em time} (p, c^{\prime}) \times (1+\sigma_{c^{\prime}}) \geq
\hbox{\em time} (p,C) \times (1-\sigma_{C})~.~~\hfill\blacksquare$

\vspace{1em}
As we will see in Section~\ref{sec:experiments}, we observed only 3,347
violations (0.73\%) of non-strict monotonicity, for the largest experiment
(Experiment 7: Confirmatory) across all
the \hbox{DBMSes} we studied, justifying our conclusion that the \hbox{DBMSes} under
study are indeed monotonic.

We can now turn to suboptimality. Recall that the monotonicity test examines
two adjacent {\QatC}s for which the {\em same plan} is observed. To detect
suboptimality, we look for adjacent {\QatC}s with {\em different plans}, termed
a ``change point,'' mentioned earlier. We look for such change points where the computed
query time at the {\em upper} cardinality is {\em smaller} than the computed
query time at the {\em lower} cardinality. Say the lower cardinality used
Plan {\em A} and the upper cardinality exhibited Plan {\em B}. Had the \hbox{DBMS}
query optimizer selected Plan {\em B} for the lower cardinality, the query time would have been smaller than that for
Plan {\em A}, which follows directly from the
monotonicity assumption. The conclusion is that for the lower cardinality, the
optimizer picked the less efficient plan, and thus, this query exhibits
\hbox{suboptimality}. Note that since this approach cannot consider plans
that were never chosen, it very likely misses some suboptimal plans (for
which there was a better plan not seen), and thus produces a
conservative estimate of suboptimality.

Our definition of suboptimality compares the computed runtimes 
and standard deviations at the cardinality just before the change
point (designated as $n-1$) and at the change point (that is, $n$).

The query is said to be suboptimal if
%\hspace*{3em}
$\hbox{\em time}_{n-1} - 0.5 \cdot \hbox{\em stddev}_{n-1} \geq \hbox{\em
time}_n + 0.5 \cdot \hbox{\em stddev}_n$~.\\
At each Q@C pair, Suboptimality is coded as four levels (0--3), based on the distance in 
standard deviations, up to three standard deviations (we chose this cutoff
because 99.7\% of all Q@C pairs are less than or equal to three standard
deviations). 
\shorten{\begin{itemize}
\item Not suboptimal, i.e., 0, when
$$\hbox{\em time}_{n-1} - 0.5 \cdot \hbox{\em
stddev}_{n-1} < \hbox{\em time}_n + 0.5 \cdot \hbox{\em stddev}_n $$
\item Low suboptimality, i.e., 1, when
$$\hbox{\em time}_{n-1} - 0.5 \cdot \hbox{\em
stddev}_{n-1} \geq \hbox{\em time}_n + 0.5 \cdot \hbox{\em stddev}_n$$
and
$$\hbox{\em time}_{n-1} 1 1.0 \cdot \hbox{\em stddev}_{n-1} < \hbox{\em time}_n +
1.0 \cdot \hbox{\em stddev}_n$$
\item Moderate suboptimality, i.e., 2, when
$$\hbox{\em time}_{n-1} - 1.0
\cdot \hbox{\em stddev}_{n-1} \geq \hbox{\em time}_n + 1.0 \cdot \hbox{\em
stddev}_n$$
and
$$\hbox{\em time}_{n-1} - 1.5 \cdot \hbox{\em stddev}_{n-1} < \hbox{\em time}_n +
1.5 \cdot \hbox{\em stddev}_n$$
\item High suboptimality, i.e., 3, when
$$\hbox{\em time}_{n-1} - 1.5
\cdot \hbox{\em stddev}_{n-1} \geq \hbox{\em time}_n + 1.5 \cdot\hbox{\em stddev}_n$$
\end{itemize}

\noindent}
We then sum this value over all Q@C pairs with different plans
(the change points) to arrive at a single integer for the query instance.

Note that while \azdb\ examines the plan at every cardinality, it only has
to actually execute the query at pairs of cardinality that are change
points.  Since many fewer {\QatC}s were involved, we could try many more
queries than the \hbox{Exhaustive} \hbox{Experiment}. In the Exploratory Experiment to be
described in Section~\ref{sec:experiments}, the value of suboptimality ranged
from 0 (no suboptimality) to 133, with the majority between 0 and 9. A full 56\% of
the queries were suboptimal.  \hbox{Because} the occurrence of large values of this
measure was so rare, we did a log transformation: $\log_{10} (1 + \mbox{\em
  subopt})$ in the confirmatory analysis.

\section{Testing the Causal Model}

\c2j{}{\subsection{Experimental Setup}\label{sec:setup}}
\shorten{As mentioned, we used \azdb\ to run the experiments. This infrastructure,
now comprising around 44K lines of code, handles all the details of
generating the sample data, generating the actual queries, collecting the
various plans generated by the optimizer, running the queries, and storing
the results in a ``labshelf,'' a set of tables managed by
a separate \hbox{DBMS}. This infrastructure also performs the many steps
required to ensure repeatability.

We have assembled a hardware lab of dedicated
machines, one \c2j{for each DBMS}{each for the three \hbox{DBMSes}} and one to run the \hbox{DBMS} used
to store the lab shelves. Having dedicated hardware allows us to worry
less about other processes running on the machines that could dirty the
results (these machines are {\em only} for experiments), and allows us to
run extensive experiments involving days or weeks of computation. We have
used this software and hardware lab to perform a number of experiments
examining the phenomenon of suboptimality.
}
\shorten{{\azdb} utilizes experiments defined in XML. It supports {\em experimental
  scenarios}, which are short (perhaps 100 lines of code) Java subclasses
that provide details of how to run kinds of experiments.}

The measurements were collected using Tucson Timing
Protocol\linebreak Version~1 (TTPv1)~\cite{TTPv1} and Version~2 (TTPv2)~\cite{TTPv2} on a suite of five machines, each an
\hbox{Intel} Core i7-870 Lynnfield 2.93GHz quad-core processor on a LGA 1156 95W
motherboard with 4GB of DDR3 1333 dual-channel memory and Western Digital Caviar
Black 1TB 7200rpm SATA hard drive, running Red Hat \hbox{Enterprise} Linux
Server release 5.8 (Tikanga) for TTPv1 and release 6.4 (Santiago) for TTPv2, with a kernel of 2.6.32-358.18.1.  The protocol
provided calculated query evaluation time, including computation and I/O time,
in msec. Both protocols were utilized exactly as specified.

No run violated the experiment-wide sanity checks. For the largest
experiment, the Confirmatory experiment 7 discussed below, approximately
10.5\% of the query executions and 5.1\%
of the queries-at-cardinality (Q@Cs) were dropped due to query execution
and Q@C sanity checks. As a result, excessive variation in calculated query
time was observed in only 0.003\% of the Q@Cs. A total of 0.35\% of Q@C adjacent
pairs violated relaxed monotonicity and 0.68\% of the Q@Cs violated strict
monotonicity, which is acceptable.

\doubleblind{In fact, over the course of the research described here, we
  found that currently available techniques did not offer sufficient query time
  precision, and so we developed TTPv1~\cite{TTPv1} to provide that needed
  precision. As our research in suboptimality proceeded, we realized that
  to get a actually recognized suboptimality for two adjacent cardinalities
  separated by only 10,000 rows (just a few pages), or even in the case of
  MySQL, 300 rows, we needed even finer resolution. This provided the
  impetus for the massive effort to develop TTPv2~\cite{TTPv2}, which did
  indeed provide the needed precision.}{We
  thank the developers of this protocol and of the software that enabled the
  running of many, many queries for providing us this software.}

In these experiments, we  installed each disk directly
on the machine that also runs the DBMS, ensured a cold cache (disk drive,
disk controller, O/S, and DBMS buffer), and discarded any sequence of query
executions that appear to be the result of query result caching. We also
ensured within-run plan repeatability.

In the following, we describe in detail the data used by our experiments and
the experimental scenarios we defined. More details on the data sets and
experiment scenarios used
can be found in Appendix~\ref{sec:app}.

\subsection{Data sets}\label{sec:datasets}
We generate our experiment data set randomly in each of the experiments. However, we
use seeds to control the random data generator so that it can produce
repeatable data as required.

Our experiment data set consists of relational tables. There are two types of
tables. The first is a ``fixed table'' that, once created and
populated, will never be modified in the future.  In contrast, the second
type is a ``variable table.'' We alter the cardinality\c2j{ of
  this table}{, physically, of such tables} as the experiments are being performed. 
\c2j{We used a}{We organized three configurations for the data sets. The first configuration
is a small} data set with four tables, each with four integer-typed
attributes. Each table is populated with one million rows, and thus the size
of each table is roughly 16Mbyte. We also produced a version of the data set with primary
keys (of the first attribute) for all tables.

\subsection{The Experiments}\label{sec:experiments}

We are interested in predicting the behavior of \hbox{DBMSes} through our
model. We selected four relational
\hbox{DBMSes}, some open source and some proprietary, that are representative of
the relational \hbox{DBMS} market. Each was used in its stock configuration.

In each experiment, we varied the
cardinality from 2M (maximum) to 10K (minimum), in increments of 10K. For
the one \hbox{DBMS} that was slower than the others and was timing out for
the majority of the queries when run between 10K and 2M, we reduced the
size of the tables and varied the cardinality from 60K (maximum) to 300 
(minimum), in increments of 300.

Utilizing JDBC to manipulate independent variables from outside the
  \hbox{DBMS} allows us to empirically generalize by moving up the $y$ axis of
  Figure~\ref{fig:empirical}, from one system to several systems and then to
  a general theory. We don't reveal the identity of the \hbox{DBMS} we studied, for two reasons. First, commercial \hbox{DBMSes}
include in their user agreements requirements not to release 
performance data. This is detrimental to science, but we have no choice but
to live with that restriction. However, in some sense the specific \hbox{DBMS}
doesn't matter, as we are studying phenomena about cost-based optimizers
{\em in general}, and so are interested in making statements that apply to
all the experimental subjects in our study.

\begin{table}[t]
\tbl{Experiments 1--7: Detailed Run Statistics\label{tab:run_stat}}
{%
\resizebox{140mm}{!}
{
\begin{tabular}{c|c|c|c|c|c|c|c}
& {\em Experiment}& {\em Protocol} & {\em Cumulative} & {\em Number of}&{\em
    Number} &{\em Number}& {\em Number of}\\
& & & {\em Hours} & {\em Query Instances}&{\em of Q@Cs} &{\em of QEs}&{\em Retained QEs}\\
\hline
1 & Monotonicity 		& --- & 38 & 60 & 12,000 & 12,000 & 12,000\\
2 & Initial Exhaustive 		& TTPv1 & 1,672 & 160 & 32,000 & 320,000 & 244,787\\
3 & Exhaustive 			& TTPv2 & 1,544 & 160 & 32,000 & 320,000 & 319,980\\
4 & Exhaustive with Keys 	& --- & 28 & 200 & 40,000 & --- & ---\\
5 & Initial Exploratory 	& TTPv1 & 560 & 780 & 8,842 & 88,420 & 68,891\\%40512+26367+2012
6 & Exploratory 		& TTPv2 & 1,663 & 1,200 & 12,560 & 125,600 & 114,377\\
7 & Confirmatory 		& TTPv2 & 11,375 & 7,640 & 99,558 & 995,580 & 890,631\\
\multicolumn{3}{c|}{\em Total}	& 16,880 & 10,200 & 236,960 & 1,861,600 & 1,650,666\\
\end{tabular}
}
}
\end{table}

Table~\ref{tab:run_stat} exhibits the statistics of running a number of 
queries in our experiments. All but the last column list the number of query
instances, etc., gathered by the protocol for each experiment. The last
column gives the number of query executions retained by the protocol. One of
the primary goals of TTPv2 was to reduce the number of QEs discarded due to
phantom processes that were indirectly detected. Experiments 3, 6, and 7
benefited from that protocol.

We performed six separate experiments, each looking at a different
aspect. It is important to emphasize that while the {\em queries} all came
from the same query pool and while the data sets were also shared by the
experiments, the {\em query executions} for the five experiments are
disjoint. As a side comment, we mention that for all four \hbox{DBMSes}, the query
plans generated by a \hbox{DBMS} for a particular \QatC\ of a particular query
varied between the experiments, but not between the query executions (QEs) of that query
instance at that cardinality, by virtue of the way we designed the
measurement protocol.

The first experiment, termed ``Monotonicity,'' was described in
Section~\ref{sec:suboptimality}. This experiment ran quickly, as it
only involved 12,000 QEs. That experiment helped us realize that we needed
to be much more sophisticated in our approach to timing queries.

\doubleblind{After we developed our six-step TTPv1 protocol,}{When TTPv1
    became available,} we performed our
second experiment, termed ``Initial Exhaustive,'' which more accurately tested the
monotonicity assumption, as also described in Section~\ref{sec:suboptimality}.
This experiment involved 160 query instances, 32,000 \QatC s, and thus
320,000 QEs, requiring 1,672 cumulative hours.  The (very small) percentage
of strict monotonicity violations observed was consistent with the remaining
variance of the query time measurement, concluding that none of
the \hbox{DBMSes} violated monotonicity. This also provides a validation of
our definition of suboptimality, which requires monotonicity.

We used the plans generated from Initial Exhaustive to classify operators as
continuous or discontinuous, as discussed in detail in
Section~\ref{sec:discontinuity}. Again we note that this experiment used a subset of
the set of {\em queries} (but {\em not} query executions) from the query pool,
to ensure that we see the same operators as encountered in that
study.

We later reran this experiment using TTPv2 (we term this simply
``Exhaustive''). Note that the number of retained QEs went up a lot (that is
the main benefit of TTPv2), while the number of hours required actually
went down (another benefit).

We also ran an experiment (number 3 in the table) on the Exhaustive query set but using the data
set with primary keys, termed ``Exhaustive with Keys.'' However, in this
case we did not actually execute 
the queries, but rather just examined the plans that were returned from the
\hbox{DBMS} to identify additional discontinuous operators. \comment{Thus this third
experiment was very fast.}

The fourth experiment was for initial exploratory analysis of a prior
version of the causal model. (This fourth experiment used
the first version of the Tucson Timing Protocol (TTPv1)~\cite{TTPv1}.)
The earlier model had fewer independent
variables yet also had more complex interactions. Specifically, the model
did not include the independent query complexity variables of presence of
secondary indexes, presence of subquery, nor the schema complexity
independent variable of presence of secondary indexes, nor the plan space
complexity independent variable of number of repeats. The prior model also
had presence of primary keys in the schema complexity construct rather than
the query complexity construct. In reformulating that independent variable,
we were able to remove a complex mediating moderator between the plan
space complexity and suboptimality constructs.

%% The following shorten corresponds to TTPv1.
The fifth experiment was an ``Initial Exploratory'' analysis of the causal model, run on
a representative sample of queries and data sets: (a)~600 queries of one
\hbox{DBMS}, (b)~120 queries of another \hbox{DBMS}, (c)~10 queries
from each other two \hbox{DBMSes}, all without primary keys defined, and (d)~10 queries
from each of the four \hbox{DBMSes} on the primary key data set, for exploration
across all combinations, a total of 780 query instances. These query sets
are enumerated in more detail in Appendix~\ref{sec:app}.
We ran the \hbox{DBMSes} on the {\em
  change points}: adjacent cardinalities (10K apart; 300 for one \hbox{DBMS}) with different
plans. The protocol retained about 78\% of the QEs.

%% The following corresponds to TTPv2.
We later reran the exploratory analysis using TTPv2, this time on a larger sample of queries and data sets:
(a)~200 queries of four \hbox{DBMSes} without primary keys defined and
(b)~100 queries from each of the four \hbox{DBMSes} on the primary key data set,
for exploration across all combinations, a total of 1,200 query instances (= 300 $\times$ 4 (DBMSes)).
We termed this sixth experiment ``Exploratory.''  We ran the DBMSes on the {\em
  change points}: adjacent cardinalities (10K apart; 300 for one DBMS) with different
plans. This experiment required 1,663 hours. The experimental
methodology retaining 12,100 {\QatC}s (3.7\% were dropped), covering 1,123 query
instances (6.4\% were dropped). Only 443 strict monotonicity violations (0.78\%) and 301 relaxed
ones (0.54\%) were observed. This exploratory analysis allowed us to refine
the operationalizations.

The seventh and final experiment was used for confirmatory analysis of the model and thus was
called ``Confirmatory.'' This was the most time-consuming of
the experiments. Here we
used (a)~800 queries on the data set without primary keys defined and (b)~510 of
those queries that had joins on the primary key attributes, on the data set
with primary keys, 
(c)~100 queries on the data set without skew and primary keys,  
(d)~100 queries with a subquery on the data set without primary keys,  
(e)~100 queries on the data set with primary keys and secondary indexes defined, 
(f)~100 queries with a subquery on the data set with primary keys defined, and 
(e)~200 queries with a subquery on the data set with primary keys and secondary indexes defined,
for a total of 1,910 queries and a total of 7,640 query instances (over the
four \hbox{DBMSes}).
Again, we ran the \hbox{DBMSes} at the 99,558 {\QatC}s that were observed,
roughly 13 per query. The protocol accurately timed
both sides of adjacent {\QatC}s (53,547 change points in all).

In this Confirmatory Experiment, we observed 3,347 strict monotonicity
violations (0.74\%), and 1,966 relaxed monotonicity violations (0.43\%) (out
of a total of 452,684 pairs
of two Q@Cs having identical plans and query instance), which provides
further confidence that monotonicity also applies to operators found in
queries over data in various contexts (as described above) and that our
operationalization of suboptimality is a valid one.

In the remainder of this section, we focus using the measured independent
and dependent variables in the Confirmatory experiment
to test the predictions that arise out of our causal model in Figure~\ref{fig:model}.

\subsection{Descriptive Statistics}
Several initial conclusions can be drawn from this confirmatory experiment, which was the
result of several years of programming effort and about 30 months of
experimental runs, summarized in Table~\ref{tab:run_stat}.

Let illustrate the descriptive statistics with the Confirmatory
experiment. In that experiment, we started with 7,640 query instances. These
resulted in 99,558 Q@Cs, with 10 each query executions (QEs). There were 890,631
retained QEs, which resulted in 6,983 retained query
instances (equivalently, 657 query instances were dropped by the
protocol). Additionally, 16 query instances were dropped {\em after} the
protocol because they were missing data for the only change point in the
query. Therefore, we were left with 6,967 queries.

First, perhaps surprisingly, 
more than half (3,933 queries out of the 6,967
query instances that emerged from the measurement protocol) exhibited suboptimality
somewhere in the range of cardinality of the varying table. 
{\em Every} \hbox{DBMS} exhibits suboptimality.

Secondly, most (3,370) of those suboptimal queries (86\%) had the maximum value of
suboptimality  (i.e., level 3) at some Q@C pair.
\shorten{a big number for a query could be many pairs, each at a small
  distance, or a few pairs, each at level 3. How many queries had at least
  one pair with subopt>=3? Answer: 670 in 5.20, 36 in 5.3, 531 in 6.0. So,
  total of 1237; query used: select count(*) FROM (select distinct experimentid, experimentname, dbms, runid,
querynum from CF_Analysis_SuboptDefn_Ver1 where subopt>=3);
query used 7_10: select count(*) FROM (select distinct experimentid, experimentname, dbms, runid,
querynum from Cnfm_Analysis_SuboptDefn where subopt_sd>=3);}

This phenomenon (query suboptimality) is likely to be a fundamental aspect
of either the algorithm (cost-based optimization) or the creator of the
algorithm (human information processing). Our model includes both effects.

Third, concerning the causal factors of suboptimality in our model,
\begin{itemize}
\item the cardinality of the effective plan space (CEPS) ranged from 1 to 24
plans across the cardinality range,
\item the number of discontinuous operators observed in plans for a query
averaged 9.4, with 85 being the maximum, and
\item the number of repeats ranged from 0 to 108, with the mean being 4.99.
\end{itemize}

\subsection{Correlational Analysis}\label{sec:corr}
We tested Hypotheses 1--7 using the strength and significance of correlations
of variables involved. These hypotheses 
predicts eleven positive relationships. Table~\ref{tab:hyptest} lists the hypotheses followed by the
correlation observed when testing each hypothesis.  (``NS'' denotes not
significant at the 0.05 level, the accepted standard for
significance. ``---'' denotes no prediction arising from the model.)
\todo{Sabah: update after filling out Table II:}As can
be seen, most of the predictions (nine) arising from the causal model are
supported and significant.  The two
exceptions are Hypotheses 1 and 2. Hypothesis 1 was not supported because
the correlation between number of operators available and suboptimality was negative,
while we predicted positive correlation.  \hbox{Hypothesis}~2 was not significant.

\subsection{Regression Analysis}\label{sec:regression}
\todo{Sabah: revisit this entire section, with confirmatory analysis numbers}
We ran a regression over the independent variables of the
model that predict suboptimality over the data from the Confirmatory experiment. 
Our model explained \% of the variance of the suboptimality dependent
variable.

We also did regressions on the causal variables for CEPS, number of repeats and for number of
operators with discontinuity. Our model explained \% of the
variance for CEPS, \% of the variance for number of repeats, and \% of the variance for
discontinuity. \shorten{need path analysis}

Hypothesis 6 predicts that the presence of secondary indexes will be a moderator,
strengthen the interactions between query complexity with plan space
complexity. It thus provides
predictions that the strength of such interactions will increase in the
presence of secondary keys. When
secondary indexes were not specified, our model explains 
\shorten{\% of the variance
of suboptimality,} \% of the variance of CEPS, and \% of the variance
of discontinuity, with all of the independent variables. When
secondary indexes were defined on the underlying tables, our model explains
\% of the variance of suboptimality, \% of the variance of CEPS, and
\% of the variance of discontinuity. These findings are all consistent with our
hypothesis predicting a positive moderation effect of primary keys.

Table~\ref{tab:hyptest8} shows that the strength of all interactions goes
up as predicted, except for ones involving aggregates. Recall
that the predictive model indicates weaker effects generally for aggregates,
as they are evaluated late in the query, and the primary key attributes are
not necessarily included in the grouping attributes.

\begin{table}[t]
\tbl{Testing Hypotheses 1--7: Correlations on the Confirmatory Study\label{tab:hyptest}}{%
%\begin{center}
\resizebox{85mm}{!}{
\begin{tabular}{c|c|c|c|c}
{\em Variable}& {\em Suboptimality}&{\em Repeats}&{\em CEPS} & {\em Discontinuity}\\
\hline
Operators in \hbox{DBMS} & ---&  {\bf H1a:}$-$0.16 & {\bf H1b:} & {\bf H1c:} \\
Correlation names & {\bf }  & {\bf H2a:} 0.54 & {\bf H2b:}  &{\bf H2c:} 0.49\\
Skew & {\bf H3:}  & --- & ---  & ---\\
Presence of aggregate & & {\bf H2d:}  & {\bf H2e:}  & {\bf H2f:} \\
Presence of primary key & & {\bf H2g:}  & {\bf H2h:}  & {\bf H2i:} \\
Presence of subquery & & {\bf H2j:}  & {\bf H2k:}  & {\bf H2l:} \\
Repeats& {\bf H7a:}  & --- & {\bf H5:}  & ---\\
CEPS &  {\bf H7b:} 0. & {\bf H4b:} &  --- &{\bf H4a:}  \\
Discontinuity & {\bf H7c:} 0.31 & --- & --- & ---\\
\end{tabular}
}
%\end{center}
}
\end{table}

\begin{table}[t]
\tbl{Testing Hypothesis 6: Interaction Strength on the Confirmatory Study\label{tab:hyptest8}}{%
%\begin{center}
\resizebox{85mm}{!}{
\begin{tabular}{c|c|c|c|c|c|c}
{\em Variable}& \multicolumn{2}{|c|}{{\em Repeats}}&\multicolumn{2}{|c|}{{\em CEPS}} & \multicolumn{2}{|c}{{\em Discontinuity}}\\
%\hline
{\em }&{Not SI\em }&{SI\em }&{Not SI\em }&{SI\em }&{Not SI\em }&{SI\em }\\
%&{\em }&{\em }&{\em }&{\em }\\
\hline
Correlation names &  & 0. &0. &0.  & 0. & 0. \\
Presence of aggregate  & 0. & 0. & 0. & 0. & 0. & 0. \\
Presence of primary key  &  & 0. & --- & --- & --- & --- \\
Presence of subquery  &  & 0. & --- & ---  & --- & --- 
\end{tabular}
}
%\end{center}
}
\end{table}

\shorten{
In our experiments, we noticed that some of the \hbox{DBMSes} might at
different times pick a different plan for the same query at the same
cardinality. We addressed this in our experiment by running the optimizer
(with a JDBC {\tt PreparedStatement}) and then executing the query ten times
in fast succession. Each time we did an {\tt EXPLAIN PLAN}  to ensure that
the originally-chosen plan was indeed used (it was).

That said, at the next cardinality, the optimizer went into action again,
and so could have chosen a {\em different} plan for that new cardinality.
Our supposition is that such a {\em heuristic} optimizer introduces
some variance in the cardinality dependent variable.

To test this, we ran regressions on just the \hbox{DBMS} that was demonstrably {\em
  not} heuristic, for the common case of primary keys. Interestingly, only
the presence of an aggregate and CEPS were significant at the 0.05 level, yet
our model explains fully 53\% of the suboptimality variance (and 56\% of the
CEPS variance and \todo{Sabah:??} of the discontinuity variance). This might imply
that heuristic optimization itself introduces perhaps 10--20% of the variance.

The size of the regression coefficients also provides useful
information. From the regressions run on all of the data, for the
statistically significant causal factors, the highest coefficients, all in
predicting the suboptimality dependent variable, were CEPS at 0.42, number
of tables involved at 0.23, number of discontinuous variables at 0.16,
number of operators available in \hbox{DBMS} at 0.06. Note though that number of correlation
variables is highly correlated with CEPS; the regression will apportion the variance of the dependent
variable among these factors.
}

\subsection{Summary of Model Testing}
In our experimental design, we started with a structural causal model that
encapsulates our theory for how cost-based query optimizers might select a
suboptimal plan for a query at a cardinality. This model implies the seven
specific hypotheses listed in Section~\ref{sec:model}. We then performed
six experiments,
\begin{itemize}
\item to refine our operationalizations: specifically discontinuity,
via the Initial \hbox{Exhaustive} and Exhaustive \hbox{Experiments}, and number of operators available, via the
Exhaustive with Keys \hbox{Experiment},
\item to test fundamental assumptions, specifically
monotonicity, via the Monotonicity \hbox{Experiment}, and
\item to test and make minor
refinements to our model: the Initial Exploratory and \hbox{Exploratory}
Experiments.
\end{itemize}
During this exploration, \doubleblind{once we completed the TTPv2 protocol,}{when the TTPv2 protocol became available,} we reran
\hbox{experiments} to avail ourselves of the increased precision due to a much
higher percentage of retained QEs and thus Q@Cs and query instances.

Throughout these six experiments, we were cognizant of the
possibility of Type 1 errors: false positives that lead one to believe a
relationship exists when it doesn't.
To control for such errors, we then performed the final Confirmatory 
Experiment on a completely different data set consisting of many more 
query instances, 7,640 in all, running on four \hbox{DBMSes} that each
utilize cost-based query optimization, to test
our refined model. Statistical inference is only possible in confirmatory
analysis, where the model and hypotheses are selected a priori.

\todo{Sabah: update with final results:}
Correlational analysis provides significant support for the model, except
for Hypotheses 1 and 2, with the implication that the influence of the
number of operators available remains largely unexplored. Via regression
analysis, the model explains 34\% of the variance of suboptimality (overall:
with primary keys a little lower and without primary keys a little higher
due to that moderating effect), 31\% of CEPS, and 35\% of discontinuity. In
all but one case, the causal influence of independent and latent variables is significant
($p<0.05$). The direction of the regression coefficients, again in all but one
case as predicted, also provides strong support for the model.

Now that the causal model has been found to be supported by the confirmatory
\hbox{analysis}, we turn to possible implications of this model.

\subsection{Identifying Root Causes of Suboptimality}\label{sec:root}

Our goal in this paper has been to understand cost-based query optimizers
as a {\em general} class of computational artifacts and to articulate and test
a predictive model \hbox{characterizing} how such optimizers,
again, as a general class, behave.  This model can be used to further
improve \hbox{DBMSes} through engineering \hbox{efforts} that benefit from the fundamental
understanding that the scientific perspective can provide.

\todo{Sabah: update with final numbers:}
Our model includes three causal factors of suboptimality: \hbox{DBMS} optimizer
complexity, query complexity, and plan space complexity. Of these factors,
the regression \hbox{coefficient} that was highest was for CEPS
(cf.~Table~\ref{tab:hyptest}: 0.51--0.54, normalized) The next \hbox{highest}
regression factor was number of tables involved, which is highly correlated
with CEPS (cf.~Table~\ref{tab:hyptest}: 0.54). These two observations imply that the
number of plans being considered is a major determinate of
suboptimality, indicating that choosing among many plans is hard, despite
many decades of research and development.

The next most influential factor is the number of discontinuous operators
(cf.~Table~\ref{tab:hyptest8}: 0.40--0.41, normalized),
\hbox{implicating} the cost model.

\shorten{Sabah, how does the following sound?}The one following that is the number of operators provided by the \hbox{DBMS}.
This factor is totally correlated with the \hbox{DBMS}, and so
is confounded with other specificities of the coding of each \hbox{DBMS}. That said,
it is still the case that as the number of operators goes up, across
all the \hbox{DBMSes} studied, the amount of suboptimality increases. This
\hbox{factor} also strongly impacts CEPS, and so indirectly impacts suboptimality
through that path.

These factors implicate two broad root causes of suboptimality across \hbox{DBMSes}:
(i)~the cost model and (ii)~the plan search process.

\section{Diminishing Returns?}\label{sec:diminishing}
In order to increase query performance, \hbox{DBMSes} are extended over time
with new \hbox{relational} query operators. Also over time, \hbox{DBMSes}
are extended with new storage and indexing structures, which themselves
elicit new query operators. For example, nested loop join was probably the
first implementation of this needed algebraic operator. B-tree index join
was only possible when such indexes were added (quite early in the history
of modern DBMSes). Adding hashing to a DBMS enables a hash join operator;
adding hash indexes enables (perhaps several variants of) hash-index join. Each
subsequent generation of the \hbox{DBMS} thus supports an ever-expanding
collection of operators, with the current incarnation the most recent
within a series of \hbox{DBMS} {\em generations}.

\subsection{A Gedanken Experiment}\label{sec:gedanken}
Consider a Gedanken experiment that examines the plan for each query at
each cardinality, that is, each Q@C, {\em for each \hbox{DBMS} generation}. In early
generations, there will be few plan changes for a given query as the cardinality varies
simply \hbox{because} there are few operators available. For
subsequent generations, some of the Q@Cs will be associated with different plans,
enabled by the new operators that were added.

In many (hopefully most) cases, a new plan selected by a generation will be
more efficient than the plan selected by the immediate previous
generation. After all, that is the very reason the new operator(s) were
added to that subsequent generation.  However, in the presence of
suboptimality, sometimes the new plan at that Q@C is {\em not} preferable,
as that \hbox{DBMS} generation's query optimizer selected the wrong plan.
Indeed, the query optimizer also evolves and improves with each new
generation, in part to minimize the chance of selecting the wrong plan.

Our predictive model in Figure~\ref{fig:model} suggests that suboptimality is causally impacted by the
independent variable of CEPS (cardinality of effective plan space), and CEPS
is naturally connected to the number of operators available\todo{Rick:
  connect with Section~\ref{sec:root} after Sabah updates that section; also
  discuss the exponential increase in CEPS possible}. A possible implication
is that as the evolution of generations of the \hbox{DBMS} add more and more
operators, suboptimality will also increase.

It is important to emphasize that our causal analysis to this point has been
{\em across multiple \hbox{DBMSes}}, each with a different set of
operators and even number of operators. Let's reprise from Section~\ref{sec:intro} on
page~\pageref{sec:intro}: Our goal, up to this point, has been ``to \hbox{understand} cost-based query
optimizers
as a {\em general} class of computational artifacts and to come up with
insights and ultimately with predictive theories about how such optimizers,
again, as a general class, behave.''

The present discussion has a quite different focus. Here we're talking about
the query optimizer {\em within} a single \hbox{DBMS}, across multiple
generations, each generation adding one or more operators.

Let's then refine our Gedanken experiment. Let's assume each subsequent
generation adds a single operator. Starting with a fixed set of queries,
for each DBMS, we run each generation on each Q@C and then sum up the query
times to compute a {\em per-generation (total) time} for that DBMS. We can
also sum over all four DBMSes, to see how this  {\em overall per-generation
  time} (a single number for each generation number) varies with generation.

The underlying question then becomes, does an additional operator made
available in a subsequent generation of the \hbox{DBMS} actually help or
hurt? More fully, is the predicted increase in suboptimality originating
from that added operator, a causal effect discovered and validated by the process
of {\em science}, specifically the application of \hbox{empirical} generalization illustrated
in Figure~\ref{fig:empirical}, compensated for by the increased performance
afforded by that operator, a benefit realized by the application of {\em
  engineering}, the articulation, elaboration, and implementation of new
storage structures and query evaluation algorithms?

Fortunately, the engineering has already been done: current DBMSes have at
their disposal multiple query operators, which have been perfected over the
decades. Adopting the engineering perspective would predict that the overall per-generation
time would {\em monotonically decrease as the generations add query evaluation
operators}.

Interestingly, adopting the scientific perspective reaches a different conclusion. Our
causal model in Figure~\ref{fig:model} predicts that as the number of
operators increases, the latent measures of plan space complexity increase,
which therefore also {\em increase} suboptimality. As the generations of the
\hbox{DBMS} contain successively greater numbers of operators, suboptimality
will also increase with generation number.  Given that each new operator
will improve a {\em shrinking}
subset of Q@Cs for any given query, while the suboptimality occurs in a portion over an
{\em expanding} subset of Q@Cs, our causal model implies that the curve of
per-generation time will fall with each successive generation, but then
start to level off as suboptimality becomes more prevalent (specifically,
the prediction is that the second derivative is either. Eventually, the
per-generation time will either asymptotically approach a horizontal line
(with the first derivative approaching 0 from below), or worse: the first
derivative changing to positive, with the per-generation time over the
queries we are studying actually {\em increasing} with \hbox{DBMS}
generation. In either case, the causal model predicts a point where {\em the
\hbox{increase} in execution time due to suboptimality obviates the decrease
enabled by the new operator}.

Put in simpler terms, engineering considerations alone imply that DBMSes can
continue to improve, whereas also adopting the scientific perspective predicts
that eventually each DBMS will hit a wall, beyond which improvement is not possible.

Does such a limit exist, and if so, how close are modern \hbox{DBMSes} to that limit?

\subsection{Simulating \hbox{DBMS} Generations}
While we do not have access to prior generations of our \hbox{DBMSes} (which is why
the previous discussion was in the form of a Gedanken experiment), we can
approximate this with an experiment using data already collected on the
current version available for each \hbox{DBMS}, to {\em simulate} the prior
generations, each successively having a fewer number of operators, and thus a smaller set of
realizable query plans. (The effect will probably be smaller in our
simulation, as we are nonetheless using the most recent query
optimizer in all of the simulated generations; only the available operators
will vary.)

In the next section, we'll explain how we characterize the generations of
each DBMS. For now assume that each DBMS is associated with a series of
generations, with each generation having one more operator. So generation 1
of the DBMS has just one operator, generation 2 has two operators, etc.

Our data consists of the 8,840 query instances and their 112,118 Q@Cs in
both the exploratory and confirmatory experiments (Experiments 6 and 7) of
the previous study. Specifically, we focus on consists of pairs of adjacent
Q@Cs (let's refer to them as the {\em lower} Q@C and the {\em upper} Q@C,
and thus we also have a {\em lower plan} and an {\em upper plan}) for the
same query at the {\em lower} and {\em upper} cardinalities, that are
adjacent (that is, separated by the minimum cardinality, either 10,000 or
300 rows), each with an actual execution time. All Q@Cs for that query and
for that DBMS having cardinalities between the upper Q@C of one pair and the
lower Q@C of the next pair are associated with that same plan. This will be
the upper plan, guaranteed by the process in which we chose those Q@Cs for
actually timing (recall from Section~\ref{sec:motivation} that we start from
the highest cardinality, that of 2M tuples, looking for changes in the
plan).

For this generational experiment, we associate with each Q@C a generation
(a positive integer) that
is the earliest generation of the DBMS that contains all the operators in the
plan. We also only consider Q@C pairs (with adjacent Q@Cs) where the lower
plan has a generation distinct from that of the upper plan.  Say the upper
generation is earlier than that of the lower plan, as illustrated in
Figure~\ref{fig:negback}. In this figure, as we scan from right to left in
decreasing cardinality, we first encounter Plan A from generation 2 at the
highest cardinality of 1320 rows (this is the
measured Q@C that has Plan A that is closest to the measurement at 910K rows), then a pair of Q@Cs with Plan A at 910K rows and Plan B at the
adjacent 900K rows, then later Plan A again at 10K rows.

\begin{figure*}[t]
%\centering
%originally 30pc
\includegraphics[width=30pc]{figures/negback.pdf}
\caption{An Example of a Negative Backward Delta\label{fig:negback}}
\end{figure*}

In this case, the thinking goes that, had we been
in the \hbox{DBMS} generation 2, the optimizer would have selected Plan A
throughout, because Plan B simply wasn't possible (as it involves an
operator not present in generation 2). Then when generation 3 was created by
adding an operator, the optimized chose Plan B for 900K. The reason Plan B
was chosen is that it is faster than Plan A at that cardinality, shown in
the figure by extrapolating the time down from 910K down to 900K (we will
revisit this extrapolation shortly).

In this particular case, as just mentioned, Plan B is more appropriate (faster) than Plan A,
but that is not the only possibility. Sometimes the later generation with
more operators available chooses the {\em wrong} plan, one that is slower
than the plan chosen by the earlier generation, due to the suboptimality
we've observed. (We'll examine an example shortly.)

The question then becomes, does a plan change by a subsequent generation
(\hbox{enabled} by the additional operator(s) in that generation) represent a win (runs faster)
or a loss (runs slower, because of suboptimality)? More broadly, do the Q@Cs
in the aggregate enabled by each succeeding generation continue to overcome
the increasing burden of suboptimality?

\subsection{Characterizing \hbox{DBMS} Generations}
To which generation do we assign each \hbox{DBMS} operator?

Note that we don't actually know the specific order in which the operators
were added to each \hbox{DBMS}. But even if we did, that order was somewhat
arbitrary, with a host of considerations going into those decisions over the
years. Given that we are using the same optimizer for each such defined
generation, we'll adopt a more systematic ordering of the operators. We
start by gathering all {\em single-operator} plans, all {\em two-operator} plans,
and so forth, and order the generation by the prevalence of their
\hbox{appearance} in these query plans.

Specifically, for each \hbox{DBMS} we designate the first
generation to contain the single operator that maximizes the number of plans
at change points, that is, maximizing the number of Q@Cs, containing just that
operator. So for example, all the plans generated by MySQL that have exactly
one operator involve just the Full Table Scan operator.  So no choice was
needed: we designate the
first generation as just having that one operator. Any plan with just that
operator can be constructed by MySQL generation 1, as well as by any
subsequent generation (as each generation includes that initial operator).

We then examine the plans containing exactly two operators.
Using MySQL again, there are two such : one with the Full Table Scan and
Full Table Scan with Join operators (888 Q@Cs) and
one with the Full Table Scan and Ref operators (112 Q@Cs). Full
Table Scan with Join is thus the operator added by Generation 2, given its
prevalence of Q@Cs. Each
subsequent generation adds that operator that maximizes the number of plans
that operator will eventual enable. So Generation 3 adds the Eq\_Ref operator,
as that operator enables 633 plans eventually. Generation 4 adds the Ref
operator and Generation 5 adds the Index operator. 
involve all five operators.

The generations of MySQL thus can be characterized from the Q@Cs we encountered: five distinct combinations of two
operators, covering 1086 Q@Cs, six combinations of three operators (1126 Q@Cs),
two combinations of four operators (104 Q@Cs), and exactly one
combination of all five operators (6 Q@Cs).\shorten{ Thus for MySQL our data implies
a simulation of five distinct generations, each adding an operator.

\begin{align*}
g_1 &= \{ \hbox{\em full table scan} \}\\
g_2 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join} \}\\
g_3 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join}, \hbox{\em eq\_ref} \}\\
g_4 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join}, \hbox{\em
  eq\_ref}, \hbox{\em ref} \}\\
g_5 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join}, \hbox{\em
  eq\_ref}, \hbox{\em ref}, \hbox{\em index} \}
\end{align*}}
For the four \hbox{DBMSes} in our study, the number of generations ranged from five
to thirteen.

\subsection{Number of Change Points Per Generation}\label{sec:CPQ}
As an illustrative example of how we can look at DBMS query optimizer
performance through the lens of change points, let's consider  the {\em
  maximum number of change points per query}, or {\em maximum CPQ}, from the
perspective of
\hbox{DBMS} generations.

For each query, we count the number of Q@Cs at each
generation, so for instance a query might have five Q@Cs at generation 1, seven Q@Cs at
generation 2, and 17 Q@Cs at generation 5. We then then take the maximum CPQ
over the queries (so perhaps 17 is the maximum for generation 5 over all
queries). Our hypothesis is that as the generation increases, the maximum
query flutter will increase, which then influences the maximum CPQ. The results are shown in Table~\ref{tab:maxcpq}.

\begin{table}[h]
\tbl{Max Change Points Per Query, Per Generation\label{tab:maxcpq}}{%
%\begin{center}
\begin{tabular}{c|cr}
{\em Generation}&{\em Maximum Change Points}\\
& {\em Per Query}\\
\hline
1 & 1\\
2 & 66\\
3 & 115\\
4 & 136\\
5 & 132\\
6 & 141\\
7 & 42\\
8 & 172\\
9 & 2\\
10 & 12\\
11 & 55\\
12 & 29\\
13 & 1\\
{\em Cumulative} & 172\\
\end{tabular}
%\end{center}
}
\end{table}

What we notice is that the maximum CPQ {\em does} increase with generation,
up through generation 8, after which it falls off dramatically. In
retrospect, this fall-off makes sense. There can be only 200 Q@Cs for a
query, because we look for query plan changes only every 10,000 tuples up to
a maximum of 2M tuples. As the generation increases, there is less ``room''
for more plans (as some fraction of the previously generated plans will
still be quite good or even optimal), and so less opportunity for flutter
which will show up as a large maximum CPQ. And indeed, at generation 8,
there is a query that has an astonishingly high number of change points,
172, all within that generation.

This analysis indicates that there is something concerning that seems to get
critical around generation 8.  And indeed, for this data set, we'll see in
Section~\ref{sec:hitthewall} that the efficacy of the query optimizer
bottoms out around that generation.

\subsection{Using Change Points}\label{sec:usingCP}
We now consider how to use data already collected, that is the {\em change
  points} discussed in Section~\ref{sec:experiments}: adjacent
cardinalities (10K apart; 300 for one \hbox{DBMS}) with different
plans. Specifically, how can change points be used to evaluate the
effectiveness of different generations of a \hbox{DBMS}?

Define $\hbox{\em ops}(p)$ for a given plan $p$ to be the set of operators
present in that plan, with some operators perhaps repeated in that plan. 
A generation $g$ (also a set of operators) is {\em applicable} to a plan $p$ if
$\hbox{\em ops}(p) \subseteq g$. By
definition, if a generation is applicable to a given plan, it is
applicable to all subsequent generations. The smallest such generation is
termed the {\em minimally applicable generation}, or {\em mingen}.
\shorten{
\shorten{We couldn't make this one work...
\todo{Young:replace in the above para with an actual SQL example query that was associated
  with two MySQL generations in a Q@C pair. use the generation numbers from
  the next para}

\todo{Rick: I found one example but the cardinality information 
was different than what's stated in the current prose...}

%%runid = 1097, querynum = 46, max_card = 60000, upper_card = 29100, lower_ard = 28800, upper_gen = 4, lower_gen = 5
%%max_cqt: 230, upper_cqt: 110, lower_cqt: 120, lower_extrapolated_cqt: 110
%% experimentname: op-pk-30K-100q-idx, pk: true, sec_idx: true
%% plan 0: ALL:Full Table Scan,ref
%% plan 1: eq_ref,index,ref
%% plan 0 at upper_card, plan 1 at lower_card
\noindent
\hspace{3ex}{\small\begin{verbatim}
        SELECT t3.id2, t3.id4, t0.id3, t2.id1 
        FROM ft_HT3 t2, ft_HT1 t1, ft_HT2 t3, ft_HT1 t0  
        WHERE  (t2.id2=t1.id2 AND t1.id2=t3.id1 AND t3.id1=t0.id4)
\end{verbatim}}
}
}
As change point consists of a pair of adjacent Q@Cs, we have two generations
to consider. Lets start with pairs for which $\hbox{\em
mingen}(\hbox{\em lower}) > \hbox{\em mingen}(\hbox{\em
  upper})$. An example is shown in Figure~\ref{fig:posback}. In this case,
the set of operators in Plan A for the Q@C at cardinality 900K requires at
least generation 3, whereas the set of operators for Plan B for the Q@C at
cardinality 910K requires but generation 2. (Note that the generations in
this example
are consecutive, but that is not required. Sometimes the generations within a
pair are wildly different.)

\begin{figure*}[t]
%\centering
%originally 30pc
\includegraphics[width=30pc]{figures/posback.pdf}
\caption{An Example of a Positive Backward Delta\label{fig:posback}}
\end{figure*}

To examine the wisdom of picking Plan A for 900K (whose time was measured as
the {\em higher} point shown (the one above the dashed line), we find the closest Q@C
also having Plan A. There are two illustrated here, at cardinalities 30K and
2000K, respectively, with the closer one at 30K. (If there is no such Q@C,
we can't do the extrapolation, and simply remove that change point from
further consideration.)

We use the slope of the dashed line between the measured query time of Plan
B at cardinalities 30K and 910K to get an estimated query time at 900k.
This realizes an estimate of how fast the query using Plan B would have run
on the database having the variable table with a cardinality of 900K. In
this particular case, Plan B looks like it would have run {\em faster} (in
fact, was was faster even at 910K in this particular example).

As illustrated in Figure~\ref{fig:discontinuity}, plans can be
discontinuous, an example being external merge sort. For that operator, as
the cardinality increases, the number of passes  will at some point by
incremented, but the time is linear for a set number of passes.
Such behavior will, for plan B present at
the two granularities of 30K and 910K, slightly overestimate the runtime of
that plan at 900K, and thus {\em underestimate} the penalty of going with
Plan A rather than Plan B. That said,
given that we are extrapolating from the actual run time at a very
close cardinality, our extrapolated estimate should be close.

We term this situation a {\em backward extrapolation}, because we are extrapolating
backward from a cardinality of 910K to one of 900K.

With this extrapolation, we can compute the {\em relative delta}, defined as the
measured time of Plan A (the chosen one) at the lower cardinality (900K) minus the
extrapolated time of Plan B at that same cardinality, divided by the
measured time of Plan A. The relative delta is thus scaled by original
measured time at that granularity, and thus has a maximum value of 1.

In this case, the relative delta is positive (the measured time for the plan
associated with the higher generation is greater than the extrapolated
time), which indicates that the optimizer chose the {\em wrong plan}: here,
Plan B should have been chosen.

On the other hand, a {\em negative} relative delta (where the measured
time for the plan associated with the higher generation is lower than the
extrapolated time, such as that illustrated in Figure~\ref{fig:negback})
implies that the additional operator(s) available to the minimally
applicable generation of the upper plan were indeed beneficial, in that that
plan was faster. (In such cases, we again divide by the larger
value, so the minimum is~-1.)

Summarizing, a
positive relative delta (extrapolated in either the forward or backward direction) implies the presence of suboptimality: the wrong
plan was chosen by the \hbox{DBMS} generation, perhaps due to the greater number of
operators available. A negative relative delta (forward or backward) implies the {\em absence} of
suboptimality: the chosen plan was faster at that cardinality than the one
chosen at the adjacent cardinality. (Note that this is a slightly more expansive definition
of suboptimality than that used earlier and illustrated in
Figure~\ref{fig:query769}.)

There are six orthogonal possibilities for each change point (that
is, a pair of adjacent Q@Cs), a total of 66,769 pairs/change
points from the Exploratory and Confirmatory Experiments. (There is also the
case of a query instance containing a lone Q@C, which we don't consider further.)

\begin{enumerate}
\item The pair of Q@Cs share the same generation: $\hbox{\em mingen}(\hbox{\em lower}) =
\hbox{\em mingen}(\hbox{\em upper})$  (62,903 Q@C pairs, 94.2\%), which we don't consider further.

\item The extrapolation yielded a computed query
time that was negative, which we also drop (10 pairs, 0.01\%).

\item The extrapolation from above and indicating
no suboptimality, termed a {\em negative backward relative
  delta}, as
  exemplified in Figure~\ref{fig:negback}, examined earlier (492 pairs,
  0.7\%).

\item The extrapolation, termed a {\em positive backward
    relative delta} and exemplified in Figure~\ref{fig:posback}, indicates
  a suboptimal plan (583 pairs, 0.9\%).

\item The extrapolation was a {\em negative forward relative delta},
  indicating no suboptimality (1218 pairs, 1.8\%).

\item A extrapolation was from below (consider Figure~\ref{fig:negback} but with
  Plan A from generation 5; we would then need to extrapolate from the
  closest Plan B, which is {\em at a smaller cardinality}), in a {\em forward
    direction}, to compute a {\em positive forward relative delta},
  indicating a suboptimal plan (795 pairs, 1.2\%).
\end{enumerate}
%1 of rel del: 2, -1 of rel del: 21 (neg backward:3, pos backward:18)
From these six possibilities, we thus retain (3) and (5), which indicate a
\hbox{suboptimal} plan at the higher generation (1710 pairs), and (4) and (6), which indicate a
non-suboptimal plan at the higher generation (1378 pairs).

\subsection{Trends Across \hbox{DBMS} Generations}\label{sec:trends}
The analysis in the previous section is for {\em a single pair of adjacent Q@Cs for a single query
running on a specific \hbox{DBMS}},
providing a {\em relative delta} for the minimally \hbox{applicable} generation. A
positive relative delta indicates that the later generation chose a
suboptimal plan; a negative relative delta indicates the later generation
did not. The relative delta is a percentage difference, and so is not \hbox{affected}
by the absolute magnitude of the run time nor by the cardinality in
question. Indeed, because it is a percentage difference, the relative delta
is not affected by the query nor even which \hbox{DBMS} is involved. We associate
this relative delta with the later generation, for it is that generation
which had the choice between the two plans for that query.

Consider how the {\em average} relative delta might behave across successive
generations.
The average relative delta is a characterization of the aggregate
impact of that generation, providing a quantitative estimate of the benefit
of adding that operator. (We use average so that each point is not impacted
by the number of pairs over which that point is computed.)  To emphasize,
the average aggregate delta aggregates over change points, queries, and even
DBMSes, providing a single average relative delta for each generation.

What does our causal model in Figure~\ref{fig:model}, supported by the
correlational and regression analyses of the Confirmatory Experiment in
Sections~\ref{sec:corr} and \ref{sec:regression}, say about this? That
causal model asserts that as the number of operators increases in
subsequent, the latent measures in plan space complexity increase, which
therefore impacts suboptimality, also in a positive direction.

Our Gedanken experiment in
Section~\ref{sec:gedanken} takes this behavior and predicts that
as the generations contain successively greater numbers of
operators, suboptimality will also increase.

If we plot the performance of the DBMS on the {\em y}-axis, say as the
total time for a workload consisting of a set of queries over a
prescribed data set, for a sequence of {\em DBMS generations} listed on the
{\em x}-axis, the average relative delta is in some way a characterization
of the {\em slope} of this relationship. A negative average relative delta
implies that the indicated generation is doing a good job, with less
suboptimality, and so the total workload execution time will go down; a
positive average relative delta implies that the suboptimal decisions are
dominating, indicated by the total workload execution time going up for that
generation.

We expect that the average relative delta for the first few generations will
be negative, reflecting new operators that improves some plans, a natural
result of the \hbox{efforts} of \hbox{DBMS} developers to increase performance over
successive generations of their \hbox{DBMS}.  With a negative slope, the
performance plot will decrease over successive DBMS generations.

That said, all is not rosy in this picture. Each new operator is applicable
to a successively smaller portion of the queries, and perhaps over a
successively smaller portion of the cardinality space.  As already noted,
our causal model predicts that the prevalence of suboptimality would
increase as the number of operators available increased.  It seems that even
with DBMS implementers doing smart things, these two considerations predict
that the average relative delta (itself a slope of the performance curve)
will {\em increase}, as suboptimality (a positive relative delta) becomes
more prevalent.

This analysis suggests then that the performance curve will thus start to
level off. That raises the
possibility of the performance curve either asymptotically approaching a horizontal line
(the first derivative approaching 0), or worse: the first derivative (the
average relative delta)
changing to positive, with the average time over the queries we are studying
actually {\em increasing} with \hbox{DBMS} generation.

So the question comes down to this:  {\em by
how much} is the increase in efficiency enabled by a new operator (for those
Q@Cs utilizing a plan containing that operator) greater than the decrease in
efficiency resulting from suboptimality (for a subset of those Q@Cs)
resulting from adding that operator? Is there a
point where the decrease due to suboptimality obviates the increase enabled
by the new operator: a DBMS generation where the average relative delta
becomes positive? How close might modern \hbox{DBMSes} be to that limit?

\subsection{Have Modern DBMSes Hit The Wall?}\label{sec:hitthewall}
We partition the relevant four sets of change points discussed in the
Section~\ref{sec:usingCP} into two groups: (i)~those
with a {\em positive} relative delta (either forward or backward), denoting a
{\em suboptimal decision} by the query optimizer at the later generation
(corresponding to a higher generation number)
and (ii)~those with a {\em negative} relative delta (either forward or
backward) relative delta, indicating a {\em non-suboptimal decision} by the
query optimizer at the later generation. (We can't state unequivocally that
the plan is optimal, because there may be yet another plan involving the
operators within that generation that is even faster.)

Let's first examine those change points for which the query optimizer made a
good decision, the non-suboptimal change points; see
Table~\ref{tab:non-subopt}.

The first thing to note is that this data aggregates the results over the
four \hbox{DBMSes}, which had generations ranging from five to 13. Thus this is the
first study we are aware of that compares the generational trends of \hbox{DBMSes}
over quite disparate code bases, thus getting at fundamental trends.

Overall, only a small number of change points, 1378, or about 2\% of the
total, satisfied the requirements listed above, including having a different
generation number for the lower and upper plans. In fact, there were no such
change points for the first three generations nor for the last
generation. The second column states the number of change points added by
that generation and the average relative delta across just those change
points. The last column states the {\em cumulative relative delta} for that
generation, defined as
the sum of the average relative delta for those change points associated
with plans having only operators in its generation, divided by the number of
change points. This is the relevant number for a generation, as it includes
all plans that would have been emitted by that generation, using the
operators at that generation's disposal.

While the numbers jump around quite a bit, especially for the first few
generations. The thing to focus on is the last column, where the cumulative
relative delta starting off at a low of
-0.28 (recall that low negative value is good, as it indicates the
additional operator was effective at lowering the execution time) and slowly decreasing to -0.23:
larger negative numbers trending to lower negative
numbers.

This general behavior matches our prediction arising from the structural causal model that it
gets harder to squeeze out performance gains as operators are added to the
\hbox{DBMS} over successive generations.

We now examine those change points for which the query optimizer made a poor
decision, in that we can surmise that there was a better plan (the one right
next to it in the adjacent Q@C pair). Table~\ref{tab:subopt} provides the
same information across the \hbox{DBMS} generations for the suboptimal change
points: those for which the
  relative deltas are positive, indicating a poor decision, as the query
  time increased. While there are
  still only a small number of change points, there are a greater number of
  {\em suboptimal} change points, with more relatively showing up at more
  recent generations. But more strikingly, the cumulative relative delta has
  the opposite behavior to the non-suboptimal change points: it {\em
    increases} over the generations, starting at 0.08 and ending at 0.25, a
  value {\em higher} than that of the non-suboptimal change points.

The last column brings the non-suboptimal and suboptimal together, stating
the {\em net cumulative relative delta}, gathering all of the change points
with plans that would have been generated by that DBMS generation
together. We see that the net starts out \hbox{at -0.24}, indicating that the new
operators are {\em decreasing} the query time for the
workload. Unfortunately, this happy situation starts to deteriorate: with
each successive generation, the improvement is less. This is as predicted:
the optimizer is struggling with both more options (plans over the available
operators) to select from and a diminished opportunity to make a significant
improvement.

\begin{table}[t]
\tbl{Non-Suboptimal Change Points\label{tab:non-subopt}}{%
%\begin{center}
\begin{tabular}{c|c|c|c}
{\em Generation}&{\em Number of} &{\em Average}&{\em Cumulative}\\
& {\em Change Points} & {\em Relative Delta} & {\em Relative Delta}\\
\hline
1 & --- & --- & 0.00\\
2 & --- & --- & 0.00\\
3 & --- & --- & 0.00\\
4 & 24 & -0.28 & -0.28\\
5 & 282 &-0.118 & -0.130\\
6 & 309 &-0.274 & -0.203\\
7 & 336 & -0.317 & -0.243\\
8 & 43 & -0.20 & -0.241\\
9 & 4 & -0.1 & -0.241\\
10 & 17 &-0.29 & -0.241\\
11 & 357 & -0.178 & -0.225\\
12 & 6 & -0.2 & -0.225\\
13 & 0 & --- & -0.225\\
{\em Cumulative}  &1378 & ---& -0.225\\
\end{tabular}
%\end{center}
}
\end{table}

\begin{table}
[t]
\tbl{Suboptimal Change Points\label{tab:subopt}}{%
%\begin{center}
\begin{tabular}{c|c|c|c|c}
{\em Generation}&{\em Number of} &{\em Average}&{\em Cumulative}&{\em Net Cumulative}\\
& {\em Change Points} & {\em Relative Delta} & {\em Relative Delta}& {\em Relative Delta}\\
\hline
1 & --- & --- & 0.00&---\\
2 & --- & --- & 0.00&---\\
3 & --- & --- & 0.00&---\\
4 & 2 & 0.2 & 0.2&-0.24\\
5 & 211 & 0.076 & 0.077&-0.045\\
6 & 86 & 0.16 & 0.102&-0.103\\
7 & 263 & 0.303 & 0.196&-0.0800\\
8 & 273 & 0.303 & 0.231&-0.0256\\
9 & 3 & 0.04 & 0.230&-0.0257\\
10 & 168 & 0.319 & 0.245&0.008\\
11 & 639 & 0.237 & 0.242&0.0296\\
12 & 53 & 0.35 & 0.245&0.0346\\
13 & 12 & 0.40 & 0.246&0.0360\\
{\em Cumulative} & 1710 & --- & 0.246&0.0360\\
\end{tabular}
%\end{center}
}
\end{table}

\begin{figure*}[t]
%\centering
%originally 30pc
\includegraphics[width=30pc]{figures/gen_benefit.pdf}
\caption{Cumulative Average Relative Delta over Generations\label{fig:benefit}}
\end{figure*}

At Generation 10, the
net actually becomes very slightly positive: meaning that the performance
curve has hit a minimum and is now on its way up, towards slower
performance. This trend increases in later generations.

Figure~\ref{fig:benefit} tells this story graphically, with the net benefit
in the center, having a least squares slope of -0.024, which is consistent with the
transition from a generation being net beneficial to actually being slightly
suboptimal around generation 10 and more so at later generations.

This highly aggregated result, extracting 3000-odd change-point pairs having
the specified properties of interest stated in Section~\ref{sec:trends} and
drawn from almost \hbox{100-thousand} Q@Cs implies that these four particular
\hbox{DBMSes}, taken together, might be close to or have already hit the wall, where
adding an operator actually slows down the average query.

And we recall the analysis in Section~\ref{sec:CPQ} in which a related
problem cropped up around generation 8.

It is important to emphasize that we can't say whether any of these DBMSes have actually
transitioned to where, for this class of queries, the errors of the
suboptimal change points overwhelm the benefits of the non-suboptimal change
points. Presumably, the DBMS vendors have done extensive tests to ensure
that the operator added to each generation did in fact effect a speedup on
the representative workloads that they use in evaluating their optimizer
enhancements. However, we {\em can} say that (a)~the trends observed
strongly point to a decreasing benefit and an increasing cost. as predicted
by the simple arguments made above, and (b)~if current DBMSes haven't yet
reached the point of diminishing returns, that possibility exists.

We should also reiterate all the provisos mentioned earlier. In all of these
experiments, we are looking at
quite simple queries, over a quite limited range of data, with only a small
percentage (3\%) of change points identified in our experimental
protocol. On the other hand, simpler queries are thought to be easier to
generate good plans than more complex queries (as Hypothesis 2 in
Section~\ref{sec:interactions} states), relational data is generally much less uniform
in its structure (e.g., we used only integer data), values (the values in our
tables are quite evenly distributed), schema (the schemas of our tables
are identical and extremely simple), and range of sizes (the smallest table is 1/200-th of the
largest table); all of these factors should independently, and certainly in concert, minimize the suboptimality.

\section{Engineering Implications}\label{sec:engineering}
We studied a particular
phenomenon, suboptimality, when the \hbox{optimizer} selecting a wrong
plan. This phenomenon is indicated by the existence of a query plan that performs more efficiently
than the DBMS's chosen plan, for the same query. From the engineering
\hbox{perspective}, it is of critical importance to understand the prevalence
of suboptimality and its causal factors. The genesis of our predictive model was a sense that suboptimality
is caused in part by the inherent complexity of these system and the
concomitant unanticipated interactions between various rules in the
optimizer.

Through a series of experiments managed by our laboratory instrument
management system, \azdb, carried out across several years, we uncovered
several surprising results that provide systematic clues as to where current optimizers come up
short and how they can be further improved.

\begin{itemize}
\item For many queries, a majority of the ones we
considered, the optimizer picked the wrong plan for at least one
cardinality, even when the cardinality estimates were completely accurate
and even for our quite simple queries.

\item A quarter of the queries exhibited significant suboptimality ($\geq 20\%$ of
        the runtime) at some cardinality.
\end{itemize}
\noindent
These two results indicate that there is still research needed on
this topic. Fortunately, the causal model helps point out
specifically where that research should be focused.

\begin{itemize}

\item Many queries exhibited {\em query fluttering}, in which the query
  optimizer returned to a previous plan at a higher cardinality.
\item Some queries exhibited significant {\em query thrashing}, with a plan change at
         almost every cardinality. While this phenomenon was first visualized
         by Haritsa et al.~\cite{harish07,Haritsa10} on some complex
         queries, we have shown that it is present even in a surprising
         percentage of simple queries.

\item Furthermore, some queries exhibited many changes
{\em to a suboptimal plan} as the \hbox{cardinality} was varied.
\end{itemize}
\noindent
These particular queries, as well as those of the right-hand side of
Figure~\ref{fig:suboptcumulative} exhibiting a large degree of suboptimality, can be a starting point for identifying the root
cause(s) of query thrashing. The phenomenon can be investigated initially on a
per-DBMS basis. Our methodology could then be
used to test proposed causal mechanisms of query thrashing across \hbox{DBMSes}, to
ascertain the generality of any proposed solutions.

\begin{itemize}
\item The causal model and our experimental results suggest that more
research is needed to improve the cost model of {\em discontinuous
operators}.

\item We also show that it may well be useful to explicitly take {\em \hbox{cardinality} estimate
  uncertainty} into account.

\item This research indicates that aggregates are {\em not} a problem, so
  that aspect of query optimization is in good shape.\todo{Rick: check with
    Sabah's results and modify around here if needed.}

\shorten{REVISE to characterize
  more precisely the relative
advantage or disadvantage of heuristic search\shorten{, and \hbox{(c)~research} is needed to
identify the inflection point where adding an operator decreases overall
\hbox{DBMS} performance}raise several unsettling concerns. These particularly simple
queries used only a few plans within any \hbox{DBMS}. Other more complex queries
will certainly use more plans. These results imply that as a \hbox{DBMS} becomes
more sophisticated, by adding operators, it also become more unstable, as
the prevalence of suboptimality grows. Perhaps there is a fixed level of
complexity beyond which a cost-based optimizer will be inexorably unstable.}
\end{itemize}

We see that costing of \hbox{discontinuous} operators is a
root cause of query suboptimality, and is thus particularly challenging to a
query optimizer. If the cost model is even a little bit off, the optimizer
might be on the wrong side of the ``jump,'' thereby selecting the wrong
plan.  That the presence of discontinuous operators has such a high
regression coefficient provides a quite specific guideline: more research is
needed to improve the accuracy of the cost model for such operators, such as
careful calibration that tunes the cost model with more accurate resource
knowledge, including the global memory capacity available, as well as to
improve the algorithms that allocate those buffer pages to specific
operators.

Concerning the plan search
process, another identified root cause of suboptimality,\shorten{one preliminary finding was that a heuristic
search process was implicated in 10--20\% of optimizer variance.} in cases
where the \hbox{DBMS} is not as sure about the cardinalities of the underlying
relations or the speed of the disk (e.g.,~if such relations migrated
frequently~\cite{reiss03}), 
perhaps the optimizer should\shorten{not try as hard to produce out the ``perfect'' plan. 
It might help if optimizers instead}
explicitly take uncertainty into account. Indeed, others have started to
\hbox{argue} that uncertainties in the query planning process should be
acknowledged and exploited~\cite{Babcock05}.

We mentioned dynamic query optimization in Section~\ref{sec:related}
earlier.  Dynamic query-reoptimization normally requires a significant
amount of information to be recorded during query execution, which can incur
non-negligible overhead on the overall query
performance~\cite{Avnur,kabra98}.  We envision that by utilizing the
proposed predictive model for suboptimality, it may be possible to enhance
reoptimization techniques such that given a particular query, a particular
data distribution, and a specific plan operator, just the important
statistics that affect the operator's performance can be identified and
should be recorded, thereby reducing the overhead of bookkeeping irrelevant
\hbox{information}.

Hence, the methodology introduced in this paper suggests fairly specifically where additional
engineering is needed (the cost model of discontinuous
operators and accommodating cardinality estimate uncertainty) and is not
needed (costing of aggregation).

The generational study in Section~\ref{sec:diminishing} though implies that
the challenge in daunting. That study validates the implications of the
causal model in Figure~\ref{fig:model}, which correctly predict  the almost
inexorable rise in net cumulative relative delta, which implies that the
suboptimizer will eventually hit the wall where it is no longer improving.

We emphasize that this section of this paper, considering engineering
implications of the underlying causal model, contrasts with the rest of the
paper, whose focus is on the science and on understanding these complex
systems at a fundamental level.  Good engineering should be built on solid
scientific results. This paper focuses on the latter.

\section{Summary}\label{sec:summary}

This paper studies an important
component of a \hbox{DBMS}, the query optimizer. This component is an amazingly
sophisticated piece of code, but is still not well understood after decades
of research and development.

This paper makes the following contributions in
an \hbox{attempt} to gain new understanding of this
component.\todo{Rick:make consistent with list at the beginning of the paper}
\begin{itemize}
\item Shows that even for simple queries, over a simple schema and
  relatively small range of sizes, the prevalence of {\em query
  suboptimality}, {\em query flutter}, and {\em query thrashing}, three problems that have not been systematically investigated across \hbox{DBMSes}, is
  high, and thus there is still research needed on this mature topic of
  query optimization.

\item Introduces a new {\em methodological
perspective} that treats \hbox{DBMSes} as experimental subjects within {\em empirical generalization}.

\item Proposes {\em operationalizations} of several relevant measures that apply
  even to proprietary \hbox{DBMSes}, as well as an overarching {\em predictive
    model} that attempts a causal explanation of suboptimality, encoding
   some of what is known about query optimization.

\doubleblind{\item Utilizes a novel {\em research infrastructure}, \azdb, to express, run, and
  analyze experiments.}{}

\item Tests {\em seven hypotheses} deductively derived from the predictive causal model. A
correlational analysis and a regression analysis
provided {\em strong support} for our model, across \hbox{DBMSes}, thus
qualitatively confirming what was informally known.

\item Uncovers compelling {\em evidence} (a)~that
suboptimality correlates with two operationalizations of query complexity,
(b)~that suboptimality correlates with two operationalizations of plan space
complexity, (c)~that query complexity is a contributor to plan space
\hbox{complexity,} and (d)~that schema complexity, as operationalized by the
presence of primary key attributes, moderates these three
interactions.\todo{Rick:Check again}

\shorten{\item A separate analysis suggested that heuristic plan selection
  appears to also be a source of suboptimality.}

\item For the kinds of queries we looked at, the factors that we
identified in our model: optimizer complexity, query complexity, and plan
space complexity, in concert predicted a significant portion of the variance
of suboptimality\todo{Rick: state. Do we want to keep the following?}.  It is doubtful that any other factor, as yet unknown,
will itself predict as much variance as the factors we studied in this
paper. That said, it is certain that there remain several unknown causal factors;
identifying those factors will undoubtedly also have important engineering
implications.

\item Articulates for the first time a possible {\em upper bound} on the
  number of operators a \hbox{DBMS} may be able to support, given that the
  empirical evidence suggests that additional operators speed up a smaller
  and smaller portion of the query/cardinality space while incurring an
  increasing chance of suboptimality over the remaining space, which is
  growing.

\item Applies a novel experiment over pairs of adjacent Q@Cs to show that
  such an upper bound is probably and may have already been reached with one
  or more our our subject DBMSes.

\item Identifies {\em specific directions for engineering interventions}.

\item Provides a {\em path toward scientific progress} in the
  understanding of a key enabling technology. 
  It is important to emphasize that our model doesn't apply to just one implementation of the
algorithm or to one \hbox{DBMS}.  Rather, it is quite broad, applying to any \hbox{DBMS}
with a cost-based optimizer.

\end{itemize}
This paper thus suggests a framework of casual model elaboration and
directed engineering efforts.

\section{Future Work}\label{sec:future}
There are several directions this work could be take. \todo{RICK IS HERE}With the model refinements
we propose here, additional directed pointers to engineering efforts should
emerge.\todo{Rick:DOn't say ``we'', but just say the research needs to be done}

\todo{Rick: keep following shorten?}\shorten{restate: Again, the purpose of our work is to present
  a methodological perspective that can be further improved to study more
  complex interactions present in the input to query optimizers.}

\shorten{Operationally, doing experiments over four \hbox{DBMSes} was challenging. We
had to install and configure each of these \hbox{DBMSes} and get our
experimental apparatus to work uniformly across these systems. But in terms
of testing the predictions, four is a small number, which meant that we had
insufficient data to resolve Hypothesis 2 and had a low correlation for
Hypothesis 1. We plan to augment \azdb\ to enable other \hbox{DBMSes} to be
studied; our theory makes specific predictions on what we may observe with
these other systems. Specifically, the interaction between number of \hbox{DBMS}
operators and both plan complexity and suboptimality is difficult to tease
apart, as that independent variable is entirely correlated to \hbox{DBMS}. Adding
\hbox{DBMSes} to our empirical study should help understand this interaction
better. In addition, we plan to apply path analysis, to separate direct and
indirect effects.}

We want to look into query flutter and thrashing in greater detail, as those phenomena
provide concrete indicators of problematic optimizer behavior. One
possible methodological approach is to utilize SQL optimizer hints, such as
``{\tt \*+ SEMIJOIN}'' in MySQL and ``{\tt enable\_hashjoin(false)}'' in PostgreSQL,
to encourage the optimizer to produce more plans at a given cardinality,
that can then be timed to make more explicit the entire plan space (recall that
CEPS, the cardinality of the effective plan space, is no greater and
probably much smaller than the cardinality of the plan space).

The investigations in this paper were based on very simple SPJ (select-project-join)
queries.  We wish to also consider {\em schema complexity}: 
foreign keys and indexes and {\em query complexity}: complex predicates, subqueries,
and user-defined data types, methods, and operators. \c2j{}{We also want
  to manipulate \hbox{DBMSes} internally (at least for those that are open-source),
  turning on and off the rules and observing suboptimality.} Our causal
model is extensible, in that we can add other factors, as long as their proper
operationalization can be established, and additional causal links.
So for
sub-queries, we can add nesting level, type of sub-query (scalar,
correlated, etc.), and number of ``effective joins''  (as some queries can
be rewritten into joins with the outer level).
It would also be interesting to study how a suboptimal subquery can affect
the suboptimality of the containing query.  For instance, is it true that if
many subqueries are themselves suboptimal, does that causally impact whether
the overall query is suboptimal? Finally, we may be able to determine
whether certain query rewrite techniques are employed within each \hbox{DBMS},
introducing another ``\hbox{DBMS} complexity'' factor into our model.

Kabra and
\hbox{DeWitt} have identified another source of complexity: inaccurate
statistics on the underlying tables and insufficient information about the
runtime system: ``amount of available resources (especially memory), the
load on the system, and the values of host language
variables.''~\cite[p.~106]{kabra98}.  Might there be other unanticipated
interactions, that are unknown simply because they haven't been
looked for?

\shorten{During our long-running experiments, the operating systems on which
the experimental subject \hbox{DBMSes} are installed were upgraded. This event caused
a significant increase in the number of phantom processes observed
during query executions, forcing us to eliminate many query executions for
statistical analysis, in that our model does not yet factor in a proper
measurement for the effects of phantom processes. But our proposed model
can be enhanced so that various additional parameters, such as operating system
configurations, can be examined under regression analyses, once these
parameters are properly operationalized.}

By employing an extensible causal model, many complex factors can be studied
via a systematic, statistically sound, scientific manner to better
understand the causal factors and their interactions.

In addition to the refinements of the causal model just discussed, our
research has provided specific directions for implementation interventions:
refining the cost model of discontinuous operators, improving buffer
allocation algorithms, and accommodating cardinality estimate uncertainty.

\todo{Rick: revise: talk about generational study, how we're hitting a wall,
  and how it might be necessary to fundamentally revisit query optimization,
  to come up with an entirely new approach that is less impacted by number
  of operators and by CEPS and thus avoids the flutter phenomenon.}

The methodology introduced in this paper and the causal model that results
has suggested both the scope of the problem of query suboptimality and a
number of specific engineering efforts that can now be carried out. Our
ultimate goal is a refined causal model that can fully explain how query
suboptimality arises in cost-based optimizers, thereby enabling engineering
solutions that reduce and eventually
effectively eliminate query suboptimality.

\doubleblind{\todo{Rick:Consider also mentioning somewhere Young's ergalics study
  on another kind of thrashing.}}{}

\doubleblind{\section{Acknowledgments}
This research was supported in part by NSF grants \hbox{IIS-0639106},
IIS-0415101, and EIA-0080123. We thank Rui Zhang for his help in initiating
this research, Ricardo Carlos, Preetha Chatterjee, Pallavi
Chilappagari, Jennifer Dempsey, David Gallup, Kevan Holdaway, Matthew Wong
Johnson, Andrey
Kvochko, Siou Lin, Adam Robertson,
Lopamudra Sarangi, Linh Tran, Cheng Yi, and Man Zhang for
their contributions to the \hbox{\azdb} and Phil Kaslo, Tom Lowry, and John
Luiten for constructing and maintaining our experimental instrument\c2j{}{,
  a laboratory of six machines and associated software}.
}{}

%\section{Electronic Appendix}
%The online electronic appendix to this article can be accessed in the ACM
%Digital\linebreak Library. That appendix lists...
Appendix~\ref{sec:app} provides further details on the experiments.

%\vfill\pagebreak
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.

\bibliographystyle{abbrv}
\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{99}

\vspace{0.1em}
\bibitem
%{Avnur:2000:ECA:342009.335420}
{Avnur}
R.~Avnur and J.~M.~Hellerstein, ``Eddies: Continuously Adaptive Query
Processing'', in {\em Proceedings of the ACM SIGMOD Conference},
pp.~261--272, 2000.

\bibitem
%[Babcock and Chaudhuri 2005]
{Babcock05}
B.~Babcock and S.~Chaudhuri, ``Towards a Robust Query Optimizer: A
Principled and Practical Approach,'' in {\em Proceedings of the ACM SIGMOD
  Conference}, pp.~119--130, Baltimore, Maryland, 2005.

\bibitem
% [Campbell, 1988 #829]
{campbell88}
D.~J.~Campbell, ``Task Complexity: A Review and Analysis,'' {\em Academy of Management}, 13(1), pp.~40--52, 1988.

\bibitem
%[Chaudhuri 1998]
{Chaudhuri98}
S.~Chaudhuri, ``An Overview of Query Optimization in Relational Systems,'' in {\em Proceedings of the ACM PODS Conference}, pp.~34--43, Seattle, WA, 1998.

\shorten{\bibitem
%[Cohen 1995]
{cohenbook}
P.~Cohen, {\bf Empirical Methods for Artificial Intelligence}, MIT Press, 1995.
}

\bibitem{TTPv1}
S.~ Currim, R.~T.~Snodgrass, Y.~-K.~Suh, and R.~Zhang,
``DBMS Metrology: Measuring Query Time,'' in {\em Proceedings of the ACM
  SIGMOD Conference}, pp.~421--432, June 2013.

\bibitem{TTPv2}
S.~ Currim, R.~T.~Snodgrass, Y.~-K.~Suh, and R.~Zhang,
``DBMS Metrology: Measuring Query Time,'' ACM {\em Transactions
on Database Systems}, 42+8 pages, October, 2016.

\shorten{\bibitem
%[Davies 1973]
{davies}
J.~T.~Davies, {\bf The Scientific Approach}, Academic Press, New York, 1973.

\bibitem
%[]
{gamma95}
E.~Gamma, R.~Helms, R.~Johnson, and J.~Vlissides, {\bf Design Patterns:
  Elements of Reusable Object-Oriented Software}, Addison-Wesley Publishing
Company, 1995.
}
\bibitem
%[Graefe 1993]
{Graefe93}
G.~Graefe, ``Query Evaluation Techniques for Large Databases,'' {\em ACM
  Computing Surveys} 25(2), pp.~73--170, June 1993.

\bibitem
%[Harish, Darera, and Haritsa 2007]
{harish07}
D.~Harish, P.~Darera, and J.~R.~Haritsa, ``On the Production of Anorexic
Plan Diagrams,'' in {\em Proceedings of the VLDB Conference},
pp.~1081--1092, 2007.

\bibitem{Haritsa10}
J.~R.~Haritsa, ``The Picasso Database Query Optimizer Visualizer,'' {\em
  PVLDB} 3(2):1517--1520, 2010.

\bibitem
% [Hevner et al 2004]
{hevner04}
A.~R.~Hevner, S.~T.~March, J.~Park, and S.~Ram, ``Design Science in
Information Systems Research,'' {\em MIS Quarterly} 28(1):75--105, 2004.

\bibitem{Hulgen03}
A.~Hulgen and S. Sudarshan, ``AniPQO: almost non-intrusive query optimization
for nonlinear cost functions,'' in {\em Proceedings of the VLDB Conference},
pp.~766-777, 2003,

\bibitem{Ioannidis96}
Y.~Ioannidis, ``Query Optimization,'' {\em ACM Computing Surveys}
23(1):121--123, June 1996.

\bibitem
%[Ioannidis 2003]
{Ioannidis03}
Y.~Ioannidis, ``The History of Histograms (abridged)'', in {\em Proceedings
  of the VLDB Conference}, pp.~19--30, September 2003.

\bibitem{SQL2008}
ISO, ``ISO SQL:2008 International Standard,'' 2008.

\bibitem{Jarke84}
M.~Jarke and J.~Koch, ``Query Optimization in Database Systems,'' {\em ACM
  Computing Surveys} 16(2):111--152, June 1984.
\bibitem
%[]
{kabra98}
N.~Kabra and D.~J.~DeWitt, ``Efficient mid-query re-optimization of
suboptimal query execution plans,'' in {\em Proceedings of the ACM SIGMOD Conference}, pp.~106--117, 1998.

\bibitem
%[Mannino et al.~1988]
{Mannino88}
M.~V.~Mannino, P.~Chu, and T.~Sagar, ``Statistical Profile Estimation in
Database Systems,'' {\em ACM Computing Surveys}, 20(3), pp.~192--221, 1988.

\bibitem
%[Melton 2003]
{Melton03}
J.~Melton, {\bf Advanced SQL:1999}, Morgan Kaufmann, 2003.

\bibitem{Melton11}
J.~Melton (editor), ISO/IEC 9075, Database Language SQL:2011 Part 2:
SQL/Foundation, December, 2011.

\bibitem
%[Melton and Simon 1993]
{Melton93}
J.~Melton and A.~R.~Simon, {\bf Understanding the New SQL: A Complete
  Guide}, Morgan Kaufmann, 1993.

\bibitem
%[Moody 1998]
{moody98} D.~L.~Moody, ``Metrics for Evaluating the Quality of Entity
Relationship Models,'' {\em Proceedings of International Conference on
  Conceptual Modeling}, pp.~211--225, Springer, Singapore, 1998.

\bibitem
%[Ramakrishnan \& Gehrke 2003]
{Ramakrishnan03}
R.~Ramakrishnan and J.~Gehrke, {\bf Database Management Systems}, Third
Edition, 2003.

\shorten{\bibitem
%[Reddy and Haritsa 2005]
{reddy05}
N.~Reddy and J.~R.~Haritsa, ``Analyzing Plan Diagrams of Database Query
Optimizers,'' in {\em Proceedings of the VLDB Conference}, Trondheim,
Norway, pp.~1228--1240, September, 2005.
}
\bibitem
%[Reiss and Kanungo 2003]
{reiss03}
F.~R.~Reiss and T.~Kanungo, ``A Characterization of the Sensitivity of Query
Optimization to Storage Access Cost Parameters,'' in {\em Proceedings of the
  ACM SIGMOD Conference}, pp.~385--396, 2003.

\bibitem
%[Selinger 1979]
{Selinger}
P.~G.~Selinger, M.~M.~Astrahan, D.~D.~Chamberlin, R.~A.~Lori, and
T.~G.~Price, ``Access Path Selection in a Relational Database System,''
in {\em Proceedings of the ACM SIGMOD Conference}, pp.~23--34, 1979.

\bibitem
%[]
{winslett02}
M.~Winslett, ``David DeWitt Speaks Out,'' {\em ACM SIGMOD
  Record} 31(2), pp.~50--62, June 2002.

%\bibitem
%%{Kabra:1998:EMR:276304.276315}
%{Kabra}
%N.~Kabra, D.~J.~DeWitt, ``Efficient Mid-Query Re-Optimization of Sub-Optimal Query Execution Plans'', in {\em Proceedings of the ACM SIGMOD Conference},
%pp.~106--117, 1998.

%\bibitem
%%{Avnur:2000:ECA:342009.335420}
%{Avnur}
%R.~Avnur, J.~M.~Hellerstein, ``Eddies: Continuously Adaptive Query Processing'',
%in {\em Proceedings of the ACM SIGMOD Conference}, pp.~261--272, 2000.

\end{thebibliography}

\newpage
\appendix
\section{Details on the Experiments}\label{sec:app}
Table~\ref{tab:run_stat} lists the run statistics of the seven experiments
  used in this paper. In this appendix we provide more detailed information
  on the experiments.

Table~\ref{tab:run_stat2} gives some of those details. We'll walk through
the columns in succession.

The third column states the data set used in each experiment, that is, the
specific tables being queried. There are four data sets, named A, B, C, and
D.

We first discuss the features shared between the four data sets. As
introduced in Section~\ref{sec:motivation}, the queries referenced tables
{\tt ft\_HT1}, {\tt ft\_HT2}, {\tt ft\_HT3}, and {\tt ft\_HT4}. All four
tables contain four columns, each of type integer. The specific values of
the rows for all but the first column depend
on the percentage of skewed data. Section~\ref{sec:datacomplexity} provides
the algorithm for generating the values for different values of skew; this
algorithm is used in columns two--four, which for any row will have
identical values. The first column holds a unique integer starting from 1
and going to 60K or 2M, for use in an optional primary key.

There was one version of the
last three tables, for use with MySQL, with cardinality 60K, and one version
for the rest of the DBMSes, with cardinality 2M.

We generate 200 versions of {\tt ft\_HT1}, termed the {\em variable table}. For MySQL, these version contain
300, 600, 900, 1200, $\ldots$, 59,700, and 60,000 rows; for the rest of the
DBMSes, these versions contain 10,000, 20,000, 30,000, $\ldots$, 1,970,000,
and 2M rows, as introduced in Section~\ref{sec:motivation}.

We now differentiate the data sets, elaborating on the discussion in Sections~\ref{sec:datasets}--\ref{sec:experiments}. Data Set C is perhaps the simplest to
describe: it specifies no primary key, has no duplicate rows, and has no skew
(of course, for any of the four tables). Data Set A differs from Data Set C
only in that there is skew.  As summarized at the end of
Section~\ref{sec:datacomplexity}, we use five skewness values: 0
(approximately), 0.001, 0.1, 0.5, and 1.0.

Data Set B is similar to Data Set A, adding the specification of the first
column as the primary key. And Data Set D is similar to Data Set B, adding
the specification that the other three columns should each be associated
with a secondary index, only only that one column.
We see the confirmatory examined a much larger variation of data sets than
the exploratory studies.

The next column of Table~\ref{tab:run_stat2} concerns the {\em Lab
  Shelf}. \hbox{\azdb} utilizes the metaphor of a bookshelf of lab
notebooks. Here, each shelf is associated with a version of \hbox{\azdb} itself. For the experiments in this paper, we used at various times over
the last three years lab shelves (that is, program versions) 5.19, 5.2,
6.0, and 7.1. Versions 5.19 and 5.2 were very similar; both implemented
TTPv1.  Version 6.0 also implemented TTPv1, but collected more query
measures that were not relevant for this paper. Version 7.1 implemented
TTPv2.

The \hbox{\azdb\ system also includes support} for {\em experiment scenarios}, which
are Java code that actually performs the experiment, such as varying the
cardinality and running different queries on the data. The only difference
in the scenario code across these experiments was in accommodating the
details of the data set (that is, creating secondary indexes and data skew)
and adding robustness features, such as using exponential backoff when the
network connection to the DBMS was temporarily lost.

The bottom line is that while the lab shelf and experiment scenario varied
somewhat, the only important aspect was the {\em Protocol} column of
Table~\ref{tab:run_stat}.

We now turn to the fifth column, ``what was examined?'' Here there are just two
possibilities, all 200 cardinalities or just the cardinalities at which the
query plan changed. The sixth column, ``what was timed?'', indicates that experiments three and four, Exhaustive and
Exhaustive with Keys, described in Sections~\ref{sec:discontinuity} and \ref{sec:experiments}, just
collected query plans, and thus took much less time to run.

The last column states how many Q@Cs the experiments measured (each with 10
QEs), termed {\em raw}, and how many Q@Cs were retained after the protocol
implied by the fourth column and listed explicitly in the third column of
Table~\ref{tab:run_stat} dropped query executions and Q@Cs via its many
  sanity checks.

\begin{table}[t]
\tbl{Experiments 1--7: Detailed Run Statistics\label{tab:run_stat2}}
{%
\resizebox{140mm}{!}
{
\begin{tabular}{c|c|c|c|c|c|c}
& {\em Experiment}& {\em Data Sets} & {\em Lab Shelves} & {\em What was} &{\em Was query}&{\em Number of}\\
&                 & {\em Used}      &                   & {\em Examined?}&{\em Timed?}   &{\em Retained (Raw)}\\
&&&&&& {\em Q@Cs}\\
\hline
1 & Monotonicity 	& A & 6.0     &all cardinalities&yes& 12,000 (12,000)\\
2 & Initial Exhaustive 	& A & 5.19 + 6.0&all cardinalities&yes& 27,948 (32,000)\\
3 & Exhaustive 		& A & 7.1     &all cardinalities&no & 29,515 (32,000)\\
4 & Exhaustive with Keys & B& 6.0     &change points   &no& --- \\
5 & Initial Exploratory & A + B&5.19 + 5.2 + 6.0&change points&yes& 8,171 (8,842)\\
6 & Exploratory 	& A + B&7.1   &change points   &yes& 12,100 (12,560)\\
7 & Confirmatory 	&A + B + C +D&7.1&change points&yes& 94,502 (99,558)\\
\multicolumn{2}{c|}{\em Total}         &            & &               &   & 184,236 (196,960)\\
\end{tabular}
}
}
\end{table}

\todo{Rick: I (Young) updated the following. Please review it.}
The following sets of queries were used in the seven experiments.

\begin{description}
\item[QSa] 100 queries over the four tables, generated as described in Section~\ref{sec:querycomplexity}
\item[QSb] 100 queries, generated the same way
\item[QSc] 100 queries
\item[QSd] 100 queries
\item[QSe] 100 queries
\item[QSf] 100 queries
\item[QSg] 100 queries
\item[QSh] 100 queries
\item[QSi] 100 queries
\item[QSj] 100 queries
\item[QSl] A primary key query set consisting of 230 queries from {\em
  QSa}--{\em QSf}
\item[QSm] A primary key query set consisting of 160 queries: 40 queries
  drawn from each of {\em QSg}--{\em QSj}
\item[QSn] A primary key query set consisting of 110 new queries
\item[QSo] A query set consisting of 100 queries without aggregates (for Data Set C)
\item[QSp] A subquery query set consisting of 100 queries, each with a subquery
\item[QSq] A subquery query set consisting of 100 queries, each with a subquery
\item[QSr] A query set consisting of 390 queries drawn from {\em QSl} + {\em QSm}
\item[QSs] A primary key query set consisting of 100 queries drawn from {\em QSr}
\end{description}

Experiment 1 (Monotonicity) used the first 50 queries from QSa
for one DBMS plus the first six queries from QSa and two queries each from
{\em QSd} and {\em QSe} for MySQL, for a total of 60 query instances.

Experiment 2 (Initial Exhaustive) used the first 50 queries from {\em QSa}
for the three other DBMSes plus the ten queries for MySQL from Experiment 1, for
a total of 160 query instances. Experiment 3 (Exhaustive) used the same 160 queries.

Experiment 4 (Exhaustive with Keys) used the first 50 queries from {\em QSa}
for all four DBMSes, for a total of 200  query instances.

Experiment 5 (Initial Exploratory) used the (100) queries from {\em
  QSa}--{\em QSf} (for
one DBMS), the first 20 queries from {\em QSa}--{\em QSf} (for another DBMS), the first
10 (primary key) queries from {\em QSa} (for all four DBMSes), and the first 10
(non-primary key) queries from {\em QSa} (for the other two DBMSes), for a total of 780 query instances.

Experiment 6 (Exploratory) used {\em QSa} and {\em QSb}, plus the first 100 (primary
key) queries from {\em QSr}, for all four DBMSes, for a total of 1200 query
instances.

Experiment 7 (Confirmatory) used {\em QSc}--{\em QSj}, 
along with {\em QSr} except the (first 100) queries included in Experiment 6, 
{\em QSn} for primary key (for two runs, or 220 queries), 
{\em QSo} for no data skew, 
{\em QSp} for primary key and subquery, 
{\em QSp} for subquery, 
{\em QSp} and {\em QSq} for primary key and secondary index and subquery, 
{\em QSs} for primary key and secondary index, 
all across all four DBMSes, for a total of 7,640 query instances.

\shorten{
\newpage
\todo{This is old material}

\input{OldTests.tex}

\input{Repeatability.tex}

\input{Measuring.tex}

\input{ExperimentSetup.tex}

\input{Discontinuous.tex}

\input{Results.tex}
}
\end{document}

% LocalWords:  PVLDB