### 2017-02-23 mlppi-comsis17/cover-letter.tex

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Young-Kyoon Suh}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{
%Editors-in-Chief of ACM Transactions on Computer Systems (TOCS): \\
%\begin{tabular}{ll}
%\vspace{.1in}
%
%Prof. Todd C. Mowry\\
%Computer Science Department\\
%Carnegie Mellon University\\
%5000 Forbes Avenue\\
%Pittsburgh, PA 15213 \\
%USA\\ 
}

\begin{flushleft}
{\large\bf Young-Kyoon Suh}
\end{flushleft}
\medskip\hrule height 1pt
%\begin{flushright}
%\hfill Department of Computer Science, The University of Arizona, Tucson, AZ 85721 USA\\
%\hfill Tel: +1-520-370-1865
%(Please understand we cannot reveal our affiliation, as the manuscript should be double-blind.)
%\end{flushright} 
\begin{flushright}
\hfill Supercomputing R\&D Center, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Korea Institute of Science and Technology Information (KISTI), 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Republic of Korea, 34141, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Tel: +82-42-869-0725
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Email: yksuh@kisti.re.kr
\end{flushright} 
%\vfill % forces letterhead to top of page

%\vspace{-0.5in}

\opening{Dear Prof. Vladimir Kurbalija (or Editor-in-Chief):}

\noindent An earlier version of this submission was published in Proceedings of the 21st ACM \hbox{International} Conference on Information and Knowledge Management (CIKM). 
This version to be submitted to Computer Science and Information Systems includes the following substantial extensions.

\begin{itemize}

\item Expanded Algorithms: we expand the description of the pre-processing, initial and optimized phase algorithms shortened in the earlier version. Furthermore, this submission contains three additional algorithms as follows:

\begin{itemize}

\item splitting overlapping ranges in Algorithm~2 on page~4,
\item checking the allowed field count limit on the split ranges on in Section 2.1.4 on page~5,
\item computing scan cost in Algorithm~3 on page~6, and
\item computing query cost in Algorithm~5 on page~7.

\end{itemize}

\item Complexity Analysis: we conduct the asymptotic running time analysis on each phase in Section 3 on page 8.

\item Experiments: we provide additional runtime statistics on the workloads used in the experiments in Table 1 on page 10. 
In addition, we compare the quality of a human expert's recommendation with that of our wizard in Figure 5 on page 10.

\item Expanded Literature Review: the related work section on page~9 elaborates in detail on the similarities and differences between our wizard and existing tools.

\end{itemize}

\closing{Sincerely yours,}

\end{letter}

\end{document}