### 2017-02-28 mlppi-comsis17/paper.tex
ComSIS working draft

\documentclass[runningheads]{comsis2}

\usepackage[ruled,linesnumbered]{algorithm2e}
%\usepackage{algorithm}
%\usepackage{amssymb,amsthm,amsmath}
\usepackage{amssymb}
\usepackage{balance}
\usepackage{subfigure}
\usepackage{float}
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
\usepackage{latexsym}
\usepackage{multirow}
\usepackage{color}
\usepackage{url}
\usepackage{comment}
\usepackage[dvips]{graphicx}          % graphicx package for including ps files

%% Necessary definitions for the running heads
\def\journalissue{Computer Science and Information Systems ?(?):??--??}
\def\paperidnum{DOI: N/A}
\setcounter{page}{1}

\title{MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads}
%\titlenote{An earlier version of this article was presented as a short paper at ACM CIKM'12.}
%% Use this if the title is too long for the running heads
%\titlerunning{An Automated Multi-level Partitioning Tool on Analytical Workloads}

\author{Young-Kyoon Suh\inst{1} \and Alain Crolotte\inst{2} \and Pekka Kostamaa\inst{2}}
%Correspondence to Young-Kyoon Suh, now at Korea Institute of Science and Technology Information
%% Use this the list of authors is too long for the running heads
%\authorrunning{Correspondence to Young-Kyoon Suh, now at Korea Institute of Science and Technology Information}

\institute{Department of Computer Science, The University of Arizona, Tucson, AZ 85721, USA\\ 
  \email{yksuh@cs.arizona.edu}
  \and
  Teradata Corporation, El Segundo, CA 90245, USA\\
  \email{\{alain.crolotte,pekka.kostamaa\}@teradata.com}
%  \and
%  Korea Institute of Science and Technology Information, Daejeon 34141 Republic of Korea\\ 
%  \email{yksuh@kisti.re.kr}
}

\def\form#1{$\langle{#1}\rangle$}
\def\range#1{$[{#1}]$}
\def\openrange#1{$({#1})$}
\def\lopenrange#1{$({#1}]$}
\def\ropenrange#1{$[{#1})$}
\def\product#1{$(#1)$}
\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{grey}{#1}}
\newcommand{\hilitey}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilitey{{\bf TODO:} {\em #1}}}
\long\def\shorten#1{}
\long\def\comment#1{}
\def\MAX{\mbox{\sl MAX}}
\def\LIMIT{\mbox{\sl LIMIT}}
\def\getQueryCost{\mbox{\sl getQueryCost}}
\def\getScanCost{\mbox{\sl getScanCost}}
\def\BigO{\mbox{\sl O}}
\def\NULL{\mbox{\sl NULL}}
\def\SolA{\mbox{\sl A}}
\def\SolB{\mbox{\sl B}}
\def\SolC{\mbox{\sl C}}
\def\SolS{\mbox{\sl S}}
\def\Qone{\mbox{\sl q1}}
\def\Qtwo{\mbox{\sl q2}}
\def\Sone{\mbox{\sl s1}}
\def\Stwo{\mbox{\sl s2}}
\def\SolSone{\mbox{\sl S1}}
\def\SolStwo{\mbox{\sl S2}}
\def\SolSthree{\mbox{\sl S3}}
\def\SolSfour{\mbox{\sl S4}}
\def\SolSfive{\mbox{\sl S5}}

\begin{document}

\maketitle

\begin{abstract} 
Typically, it is a daunting task for a database administrator (DBA) to figure out how to partition a huge fact table accessed by query workloads for better performance. To relieve such a burden, we introduce an intelligent partitioning tool to recommend an optimized partitioning on the fact table. This tool uses a greedy algorithm for search space enumeration. This space is driven by predicates of a given query workload. The tool takes advantage of the cost model of a query optimizer to prune the search space. The tool resides completely on a client and interacts with the optimizer via APIs. Thus, there is no overhead to instrument the optimizer code. Furthermore, the predicate-driven method can be applied to any clustering or partitioning scheme. We show that the tool's recommendation outperforms a human expert's solution. We also demonstrate that the recommendation scale very well with increasing workload and growing fact table.

\vspace{6pt}\textbf{Keywords:} Star Schema, Fact Table, Multi-Level Partitioning.
\end{abstract}

%\section{Introduction}
%
%The text of the paper\dots

\section{Introduction}
\label{sec:intro}

Tables and materialized views in Teradata are hash distributed based 
on a user-specified column or set of columns called {\em primary index}. 
Each virtual processor, a unit of parallelism called {\em AMP} in the \hbox{Teradata} \hbox{relational} 
database management system (DBMS), 
receives a subset of the data and stores it in hash order.

Teradata users typically choose primary index fields, 
so that the data is evenly distributed among the AMPs. 
The primary index fields are also chosen to \hbox{reflect} join fields 
in workloads to accomplish cheaper local joins that do not require 
data shuffling.

PPI (Partitioned Primary Index)~\cite{sinclair:ppi} is an optional horizontal 
partitioning scheme applied locally on each AMP's data. 
Note that in other database products, this type of partitioning would 
most likely to be called {\em clustering}. 
In the rest article we use the term \hbox{{\em partitioning}} only, not clustering, to avoid confusion. 
Database \hbox{administrators} (DBAs) usually choose the columns for PPI, 
based on join fields and single table \hbox{predicates} to optimize the queries included 
in the workload. 
The PPI columns are used to physically cluster data with the same values 
together in contiguous data blocks. 
This allows ``partition elimination'' in scans and joins for performance improvement.
PPI can be specified as single or multiple (or nested) levels. 
This type of partitioning scheme is known as 
Multi-Level PPI ({\em MLPPI)}~\cite{klindt09mlppi}.

By allowing non-qualified partitions to be eliminated, MLPPI can reduce 
significantly the amount of data to be scanned to answer a query. 
But a large number of partitions can create significant overhead, 
particularly in joins and table maintenance operations 
such as inserts and deletes, so that the selection of partitions can be usually a balancing act.

Figure~\ref{fig:exam} exemplifies an MLPPI table with potential partitioning schemes. 
This example is a subset of the {\tt LINEORDER} table from the Star Schema Benchmark (SSB)~\cite{oneil:ssb}. 
Note that the full table with modified data types was actually used in our experiments.

\begin{figure}[t]
\begin{center}
{\small
\begin{tabular}{l}
{\tt CREATE TABLE LINEORDER(} \\
\hspace{0.1in}{\tt LO\_ORDERKEY INTEGER,} \\
\hspace{0.1in}{\tt LO\_QUANTITY INTEGER,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT INTEGER} \\
{\tt )} \\
{\tt PRIMARY INDEX ( LO\_ORDERKEY )} \\
{\tt PARTITION BY (} \\
\hspace{0.1in}{\tt CASE\_N(} \\
\hspace{0.2in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN),} \\
\hspace{0.1in}{\tt CASE\_N(}\\
\hspace{0.2in}{\tt LO\_QUANTITY < 25,} \\
\hspace{0.2in}{\tt LO\_QUANTITY >= 25 AND LO\_QUANTITY <= 30,} \\
\hspace{0.2in}{\tt LO\_QUANTITY > 30 AND LO\_QUANTITY <= 35,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN)} \\
{\tt );} \\
\end{tabular}
}
\end{center}
\vspace{-0.2in}
\caption{An Example of MLPPI Table\label{fig:exam}}
\end{figure}

In the definition of {\tt LINEORDER} in the figure, the primary index 
is {\tt LO\_ORDERKEY}. 
The \hbox{primary} index \hbox{dictates} the AMP on which a row will be located, 
while the partitioning of data will be dictated by the values and 
ranges associated with {\tt LO\_DISCOUNT} and {\tt LO\_QUANTITY}. 
{\tt LO\_DISCOUNT}, for example, has two ranges for \hbox{values} 
greater than or equal to 7 and one ({\tt NO CASE OR UNKNOWN}) 
for all the other values, while \linebreak {\tt LO\_QUANTITY} has four ranges. 
As a result, the relational table will have a total of 2$\times$4 = 8 partitions as follows.

\begin{center}
\begin{tabular}{|c|p{6cm}|}\hline 
Partition & Condition \\ \hline
1		& {\scriptsize  {\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY < 25}} \\ \hline
2		& {\scriptsize  {\tt LO\_DISCOUNT >= 7 \&\& 25 <= LO\_QUANTITY <= 30}} \\ \hline
3		& {\scriptsize {\tt LO\_DISCOUNT >= 7 \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
4		& {\scriptsize {\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY no case}}\\ \hline
5       & {\scriptsize {\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY < 25}} \\ \hline 
6		& {\scriptsize {\tt LO\_DISCOUNT no case \&\& 25 <= LO\_QUANTITY < 30}} \\\hline								  	        
7		& {\scriptsize {\tt LO\_DISCOUNT no case \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
8		& {\scriptsize  {\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY no case}} \\ \hline
\end{tabular}
\end{center}

Whatever partition set is chosen, 
the selection must be semantically correct; namely, 
the mapping of rows to partitions must be an injection. 
In other words, the constraints must form a covering of 
the entire range, so that a row will belong to exactly one and one partition. 
The Teradata optimizer then applies \hbox{partition} elimination for queries 
that specify conditions on {\tt LO\_DISCOUNT} and/or {\tt LO\_QUANTITY}. 
For \hbox{example}, only partitions 1 and 2 are needed for the query 
``{\tt SELECT * FROM LINEORDER WHERE LO\_DISCOUNT >= 7 AND} {\tt LO\_QUANTITY <= 30}.'' 
Similarly, partitions 1 and 5 are sufficient to answer the query 
``{\tt SELECT * FROM LINEORDER WHERE LO\_QUANTITY < 25}.'' 
\linebreak The respective sizes of the partitions are a factor of the data distribution. 
To see MLPPI in greater detail, check our references~\cite{sinclair:ppi,klindt09mlppi}.

Next, using the full version of {\tt LINEORDER} as defined in SSB~\cite{oneil:ssb} 
we provide examples of partitioning schemes that 
a DBA may define based on a small query set. 
Consider a query set $Q$ consisting of two queries $q$1 and $q$2 from the SSB set~\cite{oneil:ssb}, as shown below.

\begin{center}
{\small
\begin{tabular}{rl}
$q$1:	& {\tt SELECT SUM(l.LO$\_$EXTENDEDPRICE*l.LO\_DISCOUNT)} \\ 
		& {\tt FROM LINEORDER l, DDATE d} \\
		& {\tt WHERE l.LO\_ORDERDATE = d.D\_DATEKEY} \\
		& {\tt AND d.D\_YEAR = `1993'} \\
		& {\tt AND l.LO\_DISCOUNT IN (1, 4, 5)} \\
        & {\tt AND l.LO\_QUANTITY <= 30} \\ 
		&						   \\	        
$q$2:	& {\tt SELECT c.C\_NATION, SUM(l.LO\_REVENUE)} \\ 
		& {\tt FROM CUSTOMER c, LINEORDER l} \\
		& {\tt WHERE l.LO\_CUSTKEY = c.LO\_CUSTKEY} \\
		& {\tt AND c.C\_REGION=`EUROPE'} \\
		& {\tt AND l.LO\_DISCOUNT >= 7} \\
		& {\tt AND l.LO\_QUANTITY >= 25 AND l.LO\_QUANTITY <= 35} \\
		& {\tt GROUP BY c.C\_NATION} \\
		& {\tt ORDER BY revenue desc}. \\
\end{tabular}
}
\end{center}

Let's now focus on the {\tt LINEORDER} fact table and 
the predicates involving {\tt LINEORDER} fields only. 
There are five constraints identified by the query number 
and the sequence number of the constraint in each query, as illustrated below.
%\vspace{-0.05in}	
\begin{center}
\begin{tabular}{|c|l|}\hline 
$q$1.1 & {\tt LO\_DISCOUNT IN (1,4,5)} \\ \hline
$q$1.2 & {\tt LO\_QUANTITY <= 30} \\ \hline
$q$2.1 & {\tt LO\_DISCOUNT >= 7} \\ \hline
$q$2.2 & {\tt LO\_QUANTITY >= 25}\\ \hline
$q$2.3 & {\tt LO\_QUANTITY <= 35} \\ \hline 
\end{tabular}
\end{center}
%\vspace{-0.05in}	
At this point the DBA needs to consider options 
based only on two fields and the five constraints. 
There are many possibilities from a {\em fine-grained} partition set 
to a loser definition.

Considering for the time being the column {\tt LO\_DISCOUNT}, 
one solution is to identify each value 
for the {\tt IN} predicate in $q$1.1 
and use $q$2.1 as is. This yields the following partitioning expression 
for {\tt LO\_DISCOUNT}:
\begin{center}
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 1,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 4,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
\end{center}
The above expression minimizes the size of the partitions 
by focusing exactly on the values required to satisfy the constraints, 
but creates a large number of small partitions.

Another possibility is to look at the maximum and minimum 
values in the {\tt IN} set and to build an {\tt AND} clause equivalent to 
a between clause yielding the following:
\begin{center}
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 1 AND LO\_DISCOUNT <= 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
\end{center}

\noindent The above partitioning solution decreases the number of partitions, 
compared to the previous solution but still focuses sharply on the constraints 
with still relatively small partitions.

Another possibility is to use the partitioning associated with {\tt LO\_DISCOUNT} shown 
in the beginning of this section with only two partitions.

Similar considerations apply to the partitioning associated with {\tt LO\_QUANTITY} 
this time with ranges and covering problems difficult to deal with. 
The resulting partitioning scheme for the table includes both {\tt LO\_DISCOUNT}
and {\tt LO\_QUANTITY}, so that the number of possible combinations is the product of 
the potential combinations for each field. 
Also, there are two queries, and one partitioning scheme may be better for one query or the other. 
As a result, the DBA will be faced with a daunting combinatorial search problem 
and no clear basis to decide on which combination is the best. 
This state of affairs begs for a {\em tool} to assist the DBA.

In the sequel, we introduce an automated tool, called {\em MLPPI wizard}. 
Indeed, the \hbox{wizard} is the first physical database design tool 
developed at Teradata.
The tool is based on a novel technique using a greedy algorithm for 
search space enumeration. 
This tool based on a general framework allowing general expressions, 
ranges and case expressions for \hbox{partition} \hbox{definitions} 
is particularly well-suited for an MLPPI definition. \linebreak
The {\em predicate-driven} \hbox{technique} used by the tool can be applied 
to any clustering or partitioning based on simple fields and expressions or complex SQL predicates. 
The wizard also borrows the optimizer's cost model 
to prune the search space and reach a final solution.

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.5]{architecture}
%\vspace{-0.2in}
\caption{The MLPPI Wizard Architecture\label{fig:arch}}
\end{figure*}

Figure~\ref{fig:arch} illustrates how our tool recommends 
the final MLPPI customized for a given workload consisting of 
a set of queries and corresponding weights. 
The wizard passes through a series of phases: {\em Preprocessing}, 
{\em Initial}, and {\em Optimized} Phases. 
In the preprocessing phase the tool produces the most granular partitioning (MGP) based on the predicates collected from every query.
In the initial phase the wizard refines the passed MGP 
by merging a pair of partitions with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}) 
and then yields an initial (feasible) MLPPI based on the refined partitions.
In the \hbox{optimization} phase, the tool makes a further attempt to optimize the initial MLPPI. 
We provide in greater \hbox{detail} each phase, 
using the query set $Q$ consisting of $q$1 and $q2$ in the rest of the article.

\shorten{
An initial MLPPI is produced through the top two phases reflecting practical constraints by field or partition. 
It is further optimized in the last phase.
In the preprocessing phase, 
for each query its predicate(s) is extracted via an API call (on {\tt explain}) made to the server. 
Then, the wizard produces the most granular partitioning (MGP) based on the 
predicates collected from every query. 
If the MGP has too many fields over a field count limit, 
then the wizard \hbox{incrementally} throws away a field from the MGP until the limit.

In the initial phase, the passed MGP gets reduced if it includes more partitions than a partition count limit allowed by an MLPPI. 
The refinement can be \hbox{incrementally} done by merging a pair of partitions 
with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}), 
until the limit. 
After the phase, an initial (feasible) MLPPI is produced based on the refined partitions.

In the optimization phase, the tool makes a further attempt to optimize the initial MLPPI. 
Similar to the initial phase, more partitions can be reduced by merging ones with the 
least {\em query cost} (as will be described in Section~\ref{sec:opt_phase}). 
If the ongoing solution cannot be further improved, the tool makes the final MLPPI recommendation. 
We will provide more details about each phase later, 
using the query set $Q$ consisting of $q$1 and $q2$ 
in the rest of the article.}

To the best of our knowledge, there is no existing industrial tool 
to solve the {\it {\bf multi-level partitioning}} problem except our wizard. 
The wizard is contained completely on the client side, as opposed to 
previous work~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} 
requiring optimizer code extension. 
Our tool simply uses existing APIs to simplify the queries, capture fact table 
predicates and costs, and uses these items to make a recommendation. 
This makes the wizard extensible and portable to different releases 
of the Teradata DBMS server.

The article is a substantial extension of prior work~\cite{Suh12}, 
and organized in the following way. 
Section~\ref{sec:algorithms} presents the multi-level \hbox{partitioning} 
algorithms for the wizard. 
In turn, Section~\ref{sec:analysis} conducts an analysis of the complexity of 
the algorithms. 
Section~\ref{sec:experiment} reports the performance evaluation results. 
Section~\ref{sec:related_work} states how \hbox{Teradata} MLPPI wizard 
is different from some existing tools proposed by other vendors. 
Lastly, Section~\ref{sec:conclusion} summarizes our discussion.

\section{The MLPPI wizard}
\label{sec:algorithms}

In this section we describe in detail each phase of the wizard using the query set $Q$ provided in Section~\ref{sec:intro}.

\subsection{The Preprocessing Phase}
\label{sec:pre_phase}

Algorithm~\ref{algo:pre_phase} sketches 
the steps in the preprocessing phase. 
This phase consists of a total of six steps.

\vspace{-.1in}

\begin{algorithm}	
\caption{The Preprocessing Phase}
\label{algo:pre_phase}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$Q$ (input query set) }
\Output{$R$ (non-overlapping range set), $M$ (query-to-range-set map)}
\hspace{.05in} Query Simplification (on $Q$)\\
\hspace{.05in} Range Extraction (from $Q$) \\
\hspace{.05in} Non-Overlapping Range ($R$) Construction \\
\hspace{.05in} Field Count Limit Check (on $R$) \\
\hspace{.05in} Query-to-Range-Set Map ($M$) Construction \\
\hspace{.05in} Partition Count Limit Check (on $R$) \\
}
\end{algorithm}

%\vspace{-.3in}

\subsubsection{Query Simplification}

The first step of this phase is to simplify the predicate(s) of each query 
by removing redundant conditions. 
For instance, if a query predicate has the condition of 
``{\tt LO\_DISCOUNT IN (1, 4, 5) AND LO\_DISCOUNT IN (2, 3, 4, 5)}'', 
then the predicate can be simplified as ``{\tt LO\_DISCOUNT IN (4, 5)}.'' 
This can be done via an API call to the DBMS server. 
In the running example this simplification is not required.

%\vspace{-.3in}

\subsubsection{Range Extraction}

The following step is to gather the simplified predicates and 
then obtain ranges from the collected predicates. 
This can also be done through an API call to the server. 
The query predicate is of the form  
\begin{center}
\begin{tabular}{l} 
\form{\tt variable} \form{\tt op} \form{\tt constant},
\end{tabular}
\end{center}
where \form{\tt variable} is a field from {\tt LINEORDER}, 
\form{\tt op} is in 
\begin{center}
\begin{tabular}{l} 
\{{\tt $ = $ }, {\tt $ < $}, {\tt $ <= $}, {\tt $ >= $}, {\tt $ > $}, 
{\tt IN}\},
\end{tabular}
\end{center}
and \form{\tt constant} represents a constant value(s). 
All \form{\tt op}s are self-explanatory. 
In particular, `{\tt IN}' is a list predicate implying `{\tt OR}' operator.
In the running example, 
we can collect a set of predicates $P$ = \{$p$1,$p$2,$p$3,$p$4\} from $Q$, where

%\vspace{-.15in}

\begin{center}
\begin{tabular}{l} 
$p$1: {\tt l.LO\_DISCOUNT IN (1, 4, 5)} \\ 
$p$2: {\tt l.LO\_DISCOUNT >= 7} \\ 
$p$3: {\tt l.LO\_QUANTITY <= 30} \\ 
$p$4: {\tt l.LO\_QUANTITY >= 25} {\tt AND} {\tt l.LO\_QUANTITY <= 35}. \\ 
\end{tabular}
\end{center}

%\vspace{-.15in}

In turn, the wizard constructs a bi-directional {\em map} between a query and 
associated predicates, so that for each query the corresponding predicate(s) can be 
found directly in the map, and vice versa. 
In the example, a map, called $M$1, is built 
between $Q$ and $P$, and $M$1 is filled with the following entries:

$M$1:
\begin{center}
\begin{tabular}{ll} 
\form{q1, \{p1, p3\}} \\
\form{q2, \{p2, p4\}}.
\end{tabular}
\end{center}

Next, we gather fields referenced by predicates in $P$, extract 
ranges from the predicates, and group the ranges on each of the fields. 
Assume $I$ to be the list of fact table fields referenced by $P$. 
In the workload $Q$, we can see that there are two fields used by $P$: 
{\tt LO\_DISCOUNT}, {\tt LO\_QUANTITY}. 
These are added to $I$.

After that, we construct a set of distinct ranges on each field in $I$, using the predicates in $P$. 
Then, the whole range set becomes a kind of two-dimensional array, called $R$. 
The array $R$ captures the range representation of the predicates. 
Each entry under a field in $R$ represents a pair of (start and end) values forming the range. 
In the range representation, infinity ($\infty$) or -infinity (-$\infty$) can 
be used for unbounded ranges. 
``['' or ``]'' are used for closed ranges, and ``('' or ``)'' for open ranges. 
In particular, {\tt IN} predicate is represented by multiple ranges. 
For instance, the above $p$1 can be represented by the following three ranges 
\range{1, 1}, \range{4, 4}, \range{5, 5}; however, the last two consecutive 
ranges can be consolidated as \range{4, 5}. 
In the continuing example the range set $R$ is formed by the entries below.

$R$:
\begin{center}
\begin{tabular}{c|c} 
 {\tt LO\_DISCOUNT} & {\tt LO\_QUANTITY} \\ \hline
\range{1, 1} & \lopenrange{\infty, 30}\\
\range{4, 5} & \range{25, 35} \\
\ropenrange{7, \infty} &  \\
\end{tabular}
\end{center}

Then, each predicate can be mapped to its associated ranges in $R$. That is, 
we can have another map, called $M$2, between $P$ and $R$. 
Let $R$[$i$, $j$] denote the interval value for the $i$-th field and $j$-th range in $R$. 
$R$[1, 3], for instance, indicates range \ropenrange{7, \infty} under {\tt LO\_DISCOUNT}. 
In the running example, $M$2 is constructed as follows:

$M$2:
\begin{center}
\begin{tabular}{l}
$\langle${$p$1}, \{$R$[1, 1],$R$[1, 2]\}$\rangle$ \\
$\langle${$p$2}, \{$R$[1 ,3]\}$\rangle$ \\ 
$\langle${$p$3}, \{$R$[2, 1]\}$\rangle$ \\
$\langle${$p$4}, \{$R$[2, 2]\}$\rangle$. \\
\end{tabular}
\end{center}

\subsubsection{Non-Overlapping Range Construction}

Since there could be multiple predicates on the same field across queries, 
extracted ranges may overlap each other. 
The overlapping ranges must be broken without any common portion, so that 
a pair of consecutive, \hbox{non-overlapping} ranges can be 
considered for a merge in further phases.

Algorithm~\ref{algo:split_overlap} describes the way of breaking 
the \hbox{overlap} of ranges.

%\vspace{-.1in}

\begin{algorithm}[h]
\caption{Splitting Overlapping Ranges}
\label{algo:split_overlap}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$R$ having overlapping ranges}
\Output{$R$ with consecutive, non-overlapping ranges}
	\ForEach{field $i$ $\in$ $R$}{	
		$L$ $\leftarrow$ Get the range set of $i$ from $R$ and sort by start value \\
		$j$ = 0 \;
		\While{$j$ $<$ $|L|$-1}{
			$r_{j}$, $r_{j+1}$ $\leftarrow$ Adjacent ranges in $L$ \\
			\If{$r_{j}$.$end$ $\geq$ $r_{j+1}$.$start$}{
				Make a split by modifying values of $r_{j}$ and $r_{j+1}$. \\
				Insert into $L$ an intermediate range if any. \\
				Re-sort ranges in $L$ and reset $j$ to 0. \\
			}\lElse{$j$++;}
		}
		Update $R$ with the final $L$. 
	} 
}
\end{algorithm}

\vspace{-.1in}

\noindent For each field, the tool gets its associated ranges and sorts them 
by start value. 
It then checks whether or not adjacent ranges overlap each other.
If that is the case, then we make a split between the ranges. 
In the example, the split is applied on the 
overlapping ranges, \lopenrange{\infty, 30} and \range{25, 35}, under {\tt LO\_QUANTITY}. 
Subsequently, a common, intermediate range (\range{25, 30}) is newly created and inserted into the range set 
belonging to the {\tt LO\_QUANTITY} field, in order to fill the gap. 
Then, $R$ in the example is shown as follows.

\
{\it $R$}:
\begin{center}
\begin{tabular}{c|c} 
 {\tt LO\_DISCOUNT} & {\tt LO\_QUANTITY} \\ \hline
\range{1, 1}      & \openrange{\infty, 25} \\
\range{4, 5}      & \range{25, 30} \\
\ropenrange{7, \infty} & \range{31, 35} \\
\end{tabular}
\end{center}

The update of $R$ affects the existing map $M$2. 
In the running example, $p$3 and $p$4 influenced by the split get mapped to 
new range sets, \{$R$[2, 1], $R$[2, 2]\} and \{$R$[2, 2], $R$[2, 3]\}, respectively. 
Of course, the range sets mapped to $p$1 and $p$2 referencing {\tt LO\_DISCOUNT} remain unchanged. 
As a result, in the example we obtain the updated $M$2 as follows:

$M$2:
\begin{center}
\begin{tabular}{l}
$\langle${$p$1}, \{$R$[1, 1],$R$[1, 2]\}$\rangle$ \\
$\langle${$p$2}, \{$R$[1, 3]\}$\rangle$ \\ 
$\langle${$p$3}, \{$R$[2, 1], $R$[2, 2]\}$\rangle$ \\
$\langle${$p$4}, \{$R$[2, 2], $R$[2, 3]\}$\rangle$. \\
\end{tabular}
\end{center}

%\vspace{.1in}

\subsubsection{Field Count Limit Check}

%The DBMS or user may impose a limit on the number of fields in the final 
%MLPPI recommendation. 
At present, the Teradata DBMS has a limit (64) of the fields 
that can be used in an MLPPI definition.
If the number of fields present in $R$ exceeds the field count limit, 
we determine which field(s) should be thrown away to satisfy the limit. 
(The star schema fact table we use has much fewer fields than the limit, 
and thus, this step will not be executed.)
To choose victim fields, the tool computes the weighted sum of 
{\em query cost} of queries regarding each field. 
The sum can be obtained by adding up every query cost on an MLPPI using 
only the ranges under the field. 
(We will cover how to measure the query cost in Section~\ref{sec:opt_phase}.) 
We incrementally discard a field with the highest query cost sum 
until the limit is reached. 
Accordingly, we can update the existing $M$2. 
One might say that the MLPPI may not be exploited if too many ranges, 
exceeding a partition count limit (or 65,536 as of now) are found in one field. 
But we assume that such an extreme case is not expected. Even if 
it happens, we can utilize the merges of consecutive ranges to make 
the MLPPI feasible.

\subsubsection{Query-to-Range-Set Map Construction}

Now, the tool can create a bi-directional {\em query-to-range-set} map, called $M$, 
between $Q$ and $R$, using the existing maps $M$1 and $M$2. 
It leverages a transitive property from $M$1 to $M$2.
Once $M$ is constructed, 
the intermediate maps ($M$1 and $M$2) are not used in the rest of phases. 
$M$ is shown as below:

{\it $M$}:
\begin{center}
\begin{tabular}{ll} 
$\langle${$q$1}, \{$R$[1, 1], $R$[1, 2], $R$[2, 1], $R$[2, 2]\}$\rangle$ \\
$\langle${$q$2}, \{$R$[1, 3], $R$[2, 2], $R$[2, 3]\}$\rangle$.
\end{tabular}
\end{center}

In the subsequent phases $M$ gets used for computing costs and updated 
along with a merge of ranges, and $R$ is refined for an MLPPI recommendation.

\subsubsection{Partition Count Limit Check}

In this regard, the range set $R$ can be used to define an MLPPI using each field as one level. 
However, if the current number of partitions by $R$ is greater than 
the partition count limit, 
then it is not possible to make a feasible MLPPI based on ranges in $R$.

The actual partition count limit is ample enough to pass the running example, 
but to continue our discussion, assume that in our discussion the limit is {\it 15}.
From $R$ in the running example, we can obtain a total of (number of ranges in {\tt LO\_DISCOUNT}) 
$\cdot$(number of ranges in {\tt LO\_QUANTITY}) = (3+1) $\cdot$ (3+1) = {\it 16}
partitions. 
Because the total partitions surpass the limit, 
the tool should pass through the initial phase in which 
fewer partitions than the limit are made. 
(Otherwise, the wizard will proceed to the optimized phase immediately.) 
Note that one more range is counted per field in the calculation. 
The added range is equivalent to the ``{\tt NO CASE OR UNKNOWN)}'' case in Figure~\ref{fig:exam}.

The tool is now ready to proceed to the next phases, 
with $R$ (input range set, or MGP) and the prepared $M$ (query-to-range-set map). 
%In the following subsections, we describe how the MLPPI solution for $Q$ 
%can be eventually made using $M$ and $R$.

\subsection{The Initial Phase}
\label{sec:init_phase}

In this phase, we incrementally merge a range pair in $R$ to reduce partitions. 
The merge continues until the number of ongoing partitions 
drop below the partition count limit. 
Once reaching the limit, we can have a feasible MLPPI with 
the remaining partitions.

By and large, overall I/O cost may be increased by the merge. 
Since every row is hash-sorted in each AMP of the Teradata DBMS server, 
typically the full scan on a partition is done to retrieve 
all the rows matching a given query. 
In the case of a merged partition, we may read the non-qualifying rows 
that would not be seen before the merge, thereby paying more I/O to answer 
the query.  
To minimize the merge overhead, we pick up the range pair incurring 
the least I/O increase in a heuristic fashion.

To choose the desirable range pair for a merge, 
the tool leverages the scan cost of a query on a range pair.
Algorithm~\ref{algo:comp_scan_cost} represents how 
the scan cost can be computed on the range pair that \hbox{influences} a query. 
The scan cost represents the I/O cost to answer the query when the range pair is merged. It is 
modeled as the {\em number of blocks} that are to be read 
for the target query. To compute the scan cost of a query $q$, we use 
a corresponding scan cost query ($s$), and 
it can be constructed as follows: 
\begin{center}
$s$: {\tt SELECT * FROM $F$ WHERE} {\it CP},
\end{center} 
where $F$ is a fact table, and 
{\it CP} indicates the predicates  
restored from the ranges mapped to $q$ from $M$.

\begin{algorithm}[t]
\caption{Scan Cost Computation}
\label{algo:comp_scan_cost}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$q$ (query), $rp$ (range pair), $M$ (query-to-range-set map)}
\Output{Scan cost for answering $q$ when considering a merge of $rp$}
	$r_{m}$ $\leftarrow$ Consolidate $rp$ \;
	$L$ $\leftarrow$ Copy the range set mapped to $q$ from $M$\; 
	Delete ranges in $rp$ from and insert $r_{m}$ into $L$. \\
	$s$ $\leftarrow$ ``{\tt SELECT * FROM FACT$\_$TABLE WHERE }'' \;
	\ForEach{range $r$ $\in$ $L$}{	
		Restore a predicate $p$ from $r$. \\
		Add $p$ to {\tt WHERE} clause of $s$.
	}
	Make an API call with $s$ to the server. \\
	Extract the spool size (in bytes) from the result. \\
	$blc$ $\leftarrow$ Calculate the block counts from the spool size. \\
	Update $q$'s scan cost to $blc$. \\
	{\bf return} $blc$ \;
}
\end{algorithm}

If $q$ turns out to be affected by the merge of a range pair, 
then the range set associated with $q$ in $M$ is temporarily updated 
by removing the parent ranges and adding the merged one. 
The altered range set is remapped to $q$ in $M$ and then 
translated to the equivalent predicates, 
so that {\it CP} for $s$ based on the predicates can be built.
Unless the merge range pair influences $q$, then 
{\it CP} can be simply built based on the 
predicates restored from the existing ranges mapped to $q$.

To help understand the scan cost construction, let us 
consider a range pair ($rp$) of \hbox{$R$[2, 2]} and $R$[2, 3] in the running example. 
$rp$ produces the merged range, \range{25, 35}, \hbox{under} {\tt LO\_QUANTITY}. 
The merge by $rp$ affects both $q$1 and $q$2 in the workload.

Thus, the range sets of $q$1 and $q$2 on {\tt LO\_QUANTITY} are 
altered to \{\openrange{\infty, 25}, \range{25, 35}\} and \{\range{25, 35}\}, 
respectively. Of course, no change is made to the existing range sets 
of the queries on {\tt LO\_DISCOUNT}. 
Therefore, the corresponding scan cost queries for $q$1 and $q$2 
can be built as below:
\begin{center}
\begin{tabular}{rl}
$\Sone$:	& {\tt SELECT * FROM LINEORDER l} \\
		& {\tt WHERE ((l.LO\_DISCOUNT = 1)} \\
		& {\tt OR (l.LO\_DISCOUNT >= 4} \\ 
		& {\tt AND l.LO\_DISCOUNT <= 5))} \\
		& {\tt AND l.LO\_QUANTITY <= 35}  \\ 
$\Stwo$:	& {\tt SELECT * FROM LINEORDER l} \\
        & {\tt WHERE l.LO\_DISCOUNT >= 7} \\
        & {\tt AND (l.LO\_QUANTITY >= 25 AND} \\ 
        & {\tt l.LO\_QUANTITY <= 35)}.
\end{tabular}
\end{center}

It is possible to combine all the (consecutive) ranges in the (temporarily) altered set of $q$1, 
and hence, a single predicate ({\tt l.LO\_QUANTITY <= 35}) can be built.

The wizard sends such a scan cost query to the server via an API call, 
and it extracts from the server's response the {\em spool size} (in bytes), or the size of scan query result. 
Then, the tool computes as scan cost the {\em block counts} using the spool size. 
If the merge of a range pair does not impact on any query in the workload, 
the existing (previously computed) scan cost of queries 
will be reused for the range pair. 
That is, the wizard will only recompute 
the scan cost of a query {\em iff} the query is affected by 
the merge of two consecutive ranges.

In this way, the weighted sum of scan cost of queries, $T_{s}$ can be computed 
for each range pair. 
$T_{s}$ can be defined as follows:
\begin{center}
\begin{tabular}{c} 
$T_{s}$ = $\sum_{i=1}^{n}$($sc_{i}${$\cdot$}{$w_{i}$}),
\end{tabular}
\end{center}
where $n$ is the number of queries in $Q$, $sc_{i}$ is the scan cost of a query $q_{i}$ in $Q$, and 
$w_{i}$ denotes the (\hbox{non-negative}) weight associated with $q_{i}$.

Once all range pairs are examined, the tool chooses the range pair 
with the least $T_{s}$ for a merge. 
If several range pairs end up with the same least $T_{s}$, 
then the tool applies the heuristic of favoring the range pair that produces 
fewer number of partitions when merged, to make 
it faster to reach the partition count limit.

In the example, suppose that the running range pair $rp$ produces the least $T_{s}$, 
and thus, the tool merges $rp$. 
Thus we can update both $R$ and $M$ as follows.

$R$:
\begin{center}
\begin{tabular}{c|c} 
{\tt LO\_DISCOUNT} & {\tt LO\_QUANTITY} \\ \hline
\range{1, 1} & \openrange{\infty, 25} \\
\range{4, 5} & \range{25, 35}		  \\
\ropenrange{7, \infty} &  \\
\end{tabular}
\end{center}

$M$:
\begin{center}
\begin{tabular}{ll} 
$\langle${$q$1}, \{$R$[1, 1], $R$[1, 2], $R$[2, 1], $R$[2, 2]\}$\rangle$ \\
$\langle${$q$2}, \{$R$[1, 3], $R$[2, 2]\}$\rangle$
\end{tabular}
\end{center}

Note that we see that $q$2 may have the most customized partition based on the updated $M$. 
Only the single partition formed by ($R$[1, 3]$\cap${$R$[2, 2]}) is sufficient 
to retrieve all the qualifying 
rows for $q$2; thus, the other partitions can be simply eliminated. 
In the meantime, to answer $q$1, we need to read the four partitions formed by 
the top two ranges of each field. 
Unfortunately, because the partitions formed by (({$R$[1, 1]{$\cup$}$R$[1, 2]})$\cap$($R$[2, 2]) 
contain the non-qualifying rows for $q$1, $R$ cannot provide $q$1 with as much benefit as $q$2. 
However, $R$ can be a good compromise to satisfy both queries, in that 
the MLPPI derived by $R$ can potentially minimize the total execution cost 
of $Q$.

\begin{algorithm}[t]
\caption{The Initial Phase}
\label{algo:init_phase}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$M$ (query-to-range-set map), $R$ (input range set)}
\Output{$R$ with \# partitions $\leq$ partition limit}
	$W$ $\leftarrow$ Query weights \\
	\While{(\# partitions by $R$ $>$ partition limit)}
	{
		\ForEach{a range pair ($rp$) in $R$}
		{	 
			$T_{s}$ $\leftarrow$ 0 \;
			\ForEach{$q$ in $M$}
			{	
				$w$ $\leftarrow$ Get $q$'s weight from $W$ \\
				$L$ $\leftarrow$ Get the range set mapped to $q$ in $M$ \\
				\uIf(\tcp*[f]{Affected}) {(($rp$ $\cap$ $L$) $\neq$ $\varnothing$)}
				{
					$T_{s}$ += $w{\cdot}$({\it getSC}($q$,$rp$,$M$)) \tcp*[r]{See Algo.~\ref{algo:comp_scan_cost} regarding {\it getSC}()}
				}
				%\lElse{ $T_{s}$ += $w{\cdot}$($q$'s existing scan cost)}
			}
			Associate $T_{s}$ with $rp$.
		}
		Find $rp$(s) with the least $T_{s}$. \\
		If a tie happens, select the $rp$ to make fewer partitions when consolidated. \\
		Update $M$ and $R$ by the chosen $rp$.
	}
	{\bf return} $R$\;
}
\end{algorithm}

Algorithm~\ref{algo:init_phase} represents the initial phase 
that has been proposed so far. 
The algorithm starts with a query-to-range set map ($M$) and an input range set ($R$). 
Until the number of the partitions in $R$ falls below the partition limit threshold, 
the algorithm determines the range pair yielding the least total execution cost ($T_{s}$) 
and updates $M$ and $R$. At the end, the algorithm produces a final recommendation based on $R$.
%\vspace{-.1in}

Now that the total partition counts ($4${$\times$}$3$=$12$) by $R$ falls  
below the assumed limit ($15$), the wizard 
proceeds to the optimized phase with $R$ without repeating the initial phase.

\subsection{The Optimized Phase}
\label{sec:opt_phase}

%We verified the effectiveness of the tool by running it on a workload based
In this regard, using $R$ we have an initial MLPPI recommendation for the {\tt LINEORDER} fact table.
The recommendation can be used sufficiently, but 
having fewer partitions, induced by further merges, 
can enhance the overall workload performance for the following reasons. 
First, multiple file contexts by many partitions can incur 
huge overhead that impacts on the query optimizer. 
There also exists an operational threshold 
that the optimizer can handle the maximum number of the partitions at a time.
Therefore, further reducing partitions greatly helps the optimizer to manage these partitions.

A similar algorithm, as suggested in the initial phase, 
can be applied to this optimized phase. 
Instead of the scan cost, the {\em query cost} can be used thanks to a feasible MLPPI. 
The query cost is modeled as the estimated processing time 
of a query on a ``faked'' fact table applying the MLPPI definition 
reflecting the merge of a range pair. 
This cost estimation technique is similar 
to the ``\hbox{what-if}" approach~\cite{chaudhuri1998autoadmin}.

\begin{algorithm}[t]
\caption{Query Cost Computation}
\label{algo:comp_query_cost}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$q$ (query), $rp$ (range pair), $M$ (query-to-range-set map), $R$ (input range set)}
\Output{Query cost to answer $q$ when considering a merge of $rp$}
	Create an empty shadow table $H$ having the same definition as the fact table $F$, 
	including indexes and all constraints (check and referential integrity constraints). \\
	Propagate all (field and index) statistics of $F$ to $H$. \\
	\If{$F$ has materialized views (MVs)}{
		Create the equivalent MVs on $H$. \\
		Propagate the statistics of the MVs on $F$ to those of $H$. \\
	}
	$r_{m}$ $\leftarrow$ Merge $rp$ \;
	$L$ $\leftarrow$ Copy ranges associated with $q$ from $M$\; 
	Remove ranges in $rp$ from and add $r_{m}$ to $L$. \\
	$R'$ $\leftarrow$ Update $R$ with $L$ \\
	Alter $H$ using a fictitious MLPPI with $R'$. \\
	Construct and send to the server an $H$-based shadow query $h$ equivalent to $q$. \\	
	$ept$ $\leftarrow$ Estimated processing time of $h$ extracted from the server response \;
	Update $q$'s query cost to $ept$. \\
	{\bf return $ept$} 
}
\end{algorithm}

Algorithm~\ref{algo:comp_query_cost} illustrates the steps for computing 
the query cost of a query $q$, 
given a range pair ($rp$) affecting $q$ when merged. 
First, we create an empty shadow table $H$ with the same definition as $F$, 
including all indexes and constraints such as check and referential integrity 
ones. Next, we propagate all statistics of $F$ to $H$. 
This covers field and index statistics. 
If $F$ has materialized views (MVs), then we create the equivalent MVs on $H$ and subsequently, 
propagate the statistics of the MVs on $F$ to the new MVs on $H$. 
After that, we alter $H$ by an MLPPI with $R$' applying the merge of $rp$. 
The wizard then builds a {\em shadow} query $h$ equivalent to the input query $q$ 
but replacing $F$ in $q$ with $H$, and makes an API call to the {\tt EXPLAIN} tool with $h$. 
Eventually, the tool extracts from the result and returns the estimated elapsed time 
of $h$ as the query cost of $q$. 
As done in the initial phase, the query cost of a query is recomputed only if the merge affects the query.

For each range pair the wizard computes the weighted sum ($T_{q}$) of query cost of queries. 
$T_{q}$ can be similarly defined as $T_{s}$, described in Section~\ref{sec:init_phase}.
%and thus the definition of $T_{q}$ is omitted.
Let the current least $T_{q}$ be $T$ and the existing least $T_{q}$ be $T_{p}$. 
If $T$ $<=$ $T_{p}$, then for the next iteration $T_{p}$ is updated to $T$. 
In turn, the tool merges the range pair that produces $T$ and updates $M$ and $R$ along with the merge. 
The wizard repeats the process until 1) no ranges in $R$ remain, or 2) pre-defined iterations are reached. 
The optimized phase is finished if $T$ $\geq$ $T_{p}$.

\begin{figure}[t]
\begin{center}
\begin{tabular}{|l|} \hline
{\tt ALTER TABLE LINEORDER} \\
{\tt MODIFY PRIMARY INDEX} \\
{\tt PARTITION BY(} \\
\hspace{.1in}{\tt CASE\_N(} \\
	\hspace{.2in}{\tt LO\_DISCOUNT $\geq$ 1 AND LO\_DISCOUNT $\leq$ 5,} \\
	\hspace{.2in}{\tt LO\_DISCOUNT $\geq$ 7,} \\ 
	\hspace{.2in}{\tt NO CASE OR UNKNOWN),} \\
\hspace{.1in}{\tt CASE\_N(} \\
	\hspace{.2in}{\tt LO\_QUANTITY $<$ 25,} \\
	\hspace{.2in}{\tt LO\_QUANTITY $\geq$ 25 AND LO\_QUANTITY $\leq$ 35,} \\
	\hspace{.2in}{\tt NO CASE OR UNKNOWN)} \\ 
{\tt );} \\ \hline
\end{tabular}
\end{center}
\caption{An MLPPI Recommendation for Workload $Q$\label{fig:mlppi}}
\end{figure}

In the running example, suppose that in the first round, ranges $R$[1, 1] and $R$[1, 2] are chosen for a merge. 
Then, the wizard produces a merged range \range{1, 5} \hbox{under} {\tt LO\_DISCOUNT} and proceeds to the next round. 
If a range pair selected for the subsequent merge fails to improve $T_{p}$, 
then, the tool exits the optimized phase and \hbox{produces} the final MLPPI recommendation for the given workload $Q$ 
as shown in Figure~\ref{fig:mlppi}.

Algorithm~\ref{algo:opt_phase} exhibits the proposed optimized phase, 
which also applies the same heuristic to favor a range pair 
that produces fewer partitions to break a tie.  
The order of the different partitioning levels may have some impact on 
the execution time. A range condition on lower levels like the query 
in Section~\ref{sec:intro} 
``{\tt SELECT * FROM LINEORDER WHERE LO\_QUANTITY < 25}'' 
requires scanning non-consecutive partitions. 
This may \hbox{incur} some overhead at run time. 
To reduce this \hbox{effect}, we sort the partition \hbox{levels} based on number of 
partitions in descending order \hbox{before} making a final \hbox{recommendation}. 
The solution in \hbox{Figure}~\ref{fig:exam} follows this heuristic 
making the two-partition case in the first level 
followed by the four-partition case in the second level. 
The order in Figure~\ref{fig:mlppi} can go \hbox{either} way 
since both levels have the same number of \hbox{partitions}.

\begin{algorithm}[t]
\caption{The Optimized Phase}
\label{algo:opt_phase}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$M$ (query-to-range-set map), $R$ (input range set)}
\Output{MLPPI such that the total query cost is minimal}
	$W$ $\leftarrow$ Query weights \\
	$T_{p}$ $\leftarrow$ Total query cost sum on an initial MLPPI by $R$ \\
	\While{(Pre-defined iterations or $R$ $\neq$ $\varnothing$)}{
		\ForEach{range pair $rp$ in $R$}{	 
			$T_{q}$ $\leftarrow$ 0 \;
			\ForEach{$q$ in $M$}{	
				$w$ $\leftarrow$ Get $q$'s weight from $W$ \\
				$L$ $\leftarrow$ Get the range set mapped to $q$ in $M$ \\
				\uIf(\tcp*[f]{Affected}) 
				{(($rp$ $\cap$ $L$) $\neq$ $\varnothing$)}{
					$T_{q}$ += $w${$\cdot$}({\it getQC}($q$,$rp$,$M$,$R$)) 	
					\tcp*[r]{See Algo.~\ref{algo:comp_query_cost} regarding {\it getQC}()}				
				}
				\lElse{ $T_{q}$ += $w${$\cdot$}($q$'s existing query cost)}
			}
			Map $T_{q}$ to $rp$.
		}
		$T$ $\leftarrow$ The least $T_{q}$ that has been so far seen.  \\
		\lIf{$T$ $>$ $T_{p}$}{ {\bf break} \;}
		\Else{
			$T_{p}$ $\leftarrow$ $T$ \tcp*[r]{Query Cost Sum Update} 
			Find $rp$(s) with $T$. \\
			If a tie happens, select the $rp$ to make fewer partitions when consolidated. \\
			Update $M$ and $R$ by the chosen $rp$. \\
		}
	}
	{\bf return} an MLPPI by $R$ \;
}
\end{algorithm}

\vspace{-.1in}
%\vspace\fill

\section{Analysis}
\label{sec:analysis}

In this section we analyze the time complexity of 
the proposed preprocessing, initial, and optimized phases.
For the time complexity we compute the logical running time of the wizard, using the {\em number of API calls} 
made to the server during the phases.

%In this section the complexity of the proposed algorithms 
%is analyzed.

\subsection{The Preprocessing Phase}

\begin{lemma}
\label{lemma:pre_phase}  
The running time complexity for building a query-to-range map is 
$\BigO$($N${$\cdot$}$C${$\cdot$}$V$), where $N$ is the number of queries 
in a workload, $C$ is the total number of fields in the fact table, 
and $V$ is the max number of values in a field. 
\end{lemma}

{\bf Proof}. Let us consider a worst case such that given a workload of 
$N$ queries, each query references all the fields of the fact table, and 
each field is associated with at most $V$ values by the predicates of the queries. 
Each query may have a single-value range. 
A range consisting of only a single value per field can be mapped to each query. 
Thus, the running time complexity for the map construction 
is $\BigO$($N${$\cdot$}$C${$\cdot$}$V$). \qed

{\bf Comment:} In practice, our experiments showed that the bound 
$\BigO$($N${$\cdot$}$C${$\cdot$}$V$) is overly pessimistic.

\subsection{The Initial and Optimized Phases}

\begin{lemma}
\label{lemma:init_phase}  
The running time complexity for the initial or optimized phases is 
$\BigO$($M^{2}${$\cdot$}$N$), 
where $M$ is the number of range pairs, and $N$ is the number of queries in 
a given workload. 
\end{lemma}

{\bf Proof}. Suppose that the merge of every range pair influences all the queries. 
Every iteration, either the scan or query costs need to be recomputed 
for each query. 
In the first round the \hbox{re-computation} cost is paid as many times as $M${$\cdot$}$N$, 
and a chosen range pair is merged. In the subsequent round 
$M$-1 range pairs remains, and the following cost amounts to ($M$-1){$\cdot$}$N$. 
In the worst case we may end up consolidating all range pairs, thus having no partition on 
a target table. The total calls to the server can be increased up to

\begin{center}
$M${$\cdot$}$N$+($M$-1){$\cdot$}$N$+$\cdots$+$N$ = ($\sum_{i=1}^M i$){$\cdot$}$N$ = $\frac{M(M-1)}{2}${$\cdot$}$N$. 
\end{center}

\noindent Hence, the running time complexity is $\BigO$($M^{2}${$\cdot$}$N$). $\qed$

%\vspace{0.1in}

Our experiments showed that $\BigO$($M^{2}${$\cdot$}$N$), the running time complexity, 
is overly pessimistic. 
In \hbox{practice}, the number of calls made was much fewer than the \hbox{theoretic} bound, 
since it was very rare for a range pair to be associated with all queries in a workload. 
In \hbox{Section}~\ref{sec:stat} we will show how many API calls take place on the workloads 
used for evaluation.

\section{Experiment}
\label{sec:experiment}

In this section, we describe our environment settings, 
report the statistics measured in our experiments, and 
demonstrate the performance of the MLPPI wizard (WIZARD), compared with 
that of no partitioning and the partitioning by a human expert (EXP).

\subsection{Environment Settings}
\label{sec:env_set}

The MLPPI wizard was implemented as a prototype 
on top of the Teradata DBMS server. 
It was written in Java. 
The performance of the wizard was evaluated 
on the Teradata DBMS server machine running Unix.

When it comes to workload generation, we took advantage of 
our simple star schema query \hbox{generator}. 
The generator assumes that 1) available operators, 2) fields, 
and 3) the minimum and \hbox{maximum} values of the fields are already known. 
For each query, the generator first randomly selects the number of single-table 
predicates to create. 
Then, it arbitrarily chooses a field and a specific operator 
for each predicate. 
If {\tt IN} operator on a chosen field is picked up, the query generator 
determines the number of values to add to the {\tt IN} list and then chooses random values within the min-max value 
range of the field. 
In this way the generator builds a query involving the generated predicates.

A generated query is based on a template that joins {\tt LINEORDER} and {\tt DDATE} with constraints defined by the predicates. 
This template is common in customer cases like reports and form 
templates. 
Using the query generator we generated two workloads consisting of 10 queries (10Q) and 20 queries (20Q).

For the query workloads we populated the fact table with a scale of 1TByte (1TB) and 3TBytes (3TB).

\begin{table*}[t]
\caption{The statistics obtained on the used workloads\label{tab:workload_stat}}
\begin{center}
{\scriptsize
\begin{tabular}{|l|l|l|l|l|} \hline
		   			   							   & \multicolumn{2}{|c|}{Init. Phase} & \multicolumn{2}{|c|}{Opt. Phase} \\ \cline{2-5}
		   			   							   & 10Q & 20Q & 10Q & 20Q \\ \hline
{\# Input Partitions}  							   & 1,008,000 & 110,739,200 & 51,840 & 59,520 \\ \hline
{\# Input Range Pairs} 							   & 441 & 4,180 & 24 & 48\\ \hline
{\# Total Iterations}  							   & 14 & 55 & 1 & 1 \\ \hline
{\# Total API Calls (Worst Calls by Lemma~\ref{lemma:init_phase}})   & 1,305 (1,944,810) & 31,915 (349,448,000) & 78 (5,760) & 388 (46,080) \\ \hline
{\# Average API calls per Range Pair} 		  	   & 2.96 & 7.63 & 3.25 & 8.08 \\ \hline
{\# Average number of range pairs compared per iteration} 	   & 31.5 & 76.0 & 24 & 48 \\ \hline
{\# Maximum number of range pairs compared per an iteration} & 38 & 103 & 24 & 48 \\ \hline
{\# Minimum number of range pairs compared per an iteration} & 25 & 49 & N/A & N/A\\ \hline
%{\# max. API Calls made per range pair} 		   & 4.92 & 16.14 & & \\ \hline
%{\# min. API Calls made per range pair} 		   & 1.92 & & & \\ \hline
{\# Average API calls made per iteration}		   & 93.21 & 580 & 78 & 388 \\ \hline
{\# Maximum number of API calls made per an iteration} 	   & 123 & 791 & 78 & 388 \\ \hline
{\# Minimum number of API calls made per an iteration} 	   & 73 & 381 & N/A & N/A \\ \hline
\end{tabular}
}
\end{center}
\vspace{-0.3in}
\end{table*}

\subsection{Statistics Report}
\label{sec:stat}

Table~\ref{tab:workload_stat} summarizes the WIZARD's execution statistics, 
which were obtained while the recommendations for the \hbox{workloads} were produced. 
The statistics includes input partitions, 
range pairs, number of API calls made, and total iterations 
observed in each phase.

About 1 million partitions were initially derived for 10Q whereas roughly 0.11 billion partitions for 20Q. 
Although the workload size doubled, the total number of partitions was increased exponentially. 
Much more predicates with different bindings produced two orders of magnitude more ranges in 20Q than those of 10Q. 
While the huge input partitions were made, the number of collected range pairs per field was relatively small.

The total number of iterations in the initial phases 
tended to be proportional to the input \hbox{partitions} 
\hbox{generated} from the workloads. 
Note that there \hbox{happened} only one iteration in the optimized phase. 
The reason was that since the partitions produced by the \hbox{initial} phase were customized enough, 
no more \hbox{iterations} were \hbox{necessary} for further optimization.

As mentioned in Section~\ref{sec:analysis}, the total calls made to 
the server per workload were by far fewer than the theoretical bound shown in parentheses. 
The worst call counts were calculated regarding the number of queries and range pairs observed in each phase. 
The average number of calls per range pair was limited within 10. %across workloads and phases.

\begin{figure*}[t]
\vspace{-.2in}
\centering
\subfigure[1TB]{
	\includegraphics[scale=0.6]{1T_total_time}
	\label{fig:1T_eval_res}
}
\subfigure[3TB]{
	\includegraphics[scale=0.6]{3T_total_time}
	\label{fig:3T_eval_res}
}
\vspace{-0.1in}
\caption{Total Elapsed Time\label{fig:total_elapsed_time}}
\end{figure*}

\begin{figure*}[t]
\vspace{-.2in}
\centering
\subfigure[1TB]{
	\includegraphics[scale=0.6]{1T_quality}
	\label{fig:1T_quality}
}
\subfigure[3TB]{
	\includegraphics[scale=0.6]{3T_quality}
	\label{fig:3T_quality}
}
\vspace{-0.1in}
\caption{Recommendation Quality Comparison\label{fig:recomm_qual}}
\vspace{-.1in}
\end{figure*}

\subsection{Performance Evaluation}

The main focus of our demonstration is to see the performance (quality) 
of MLPPI recommendations made by the MLPPI wizard (WIZARD), 
compared with the performance of no partitioning (NO PPI) and EXP (partitioning 
by a human expert).

%Our evalution results are exhibited in Figures ~\ref{fig:eval_res} 
%and ~\ref{fig:improvement}.
%Figure~\ref{fig:eval_res} shows the times (in seconds) taken 
%to run the 10Q or 20Q workloads over the 1TB or 3TB fact tables. 
%Based on the run results, Figure~\ref{fig:improvement} displays 
%how much improvement could be achieved by the EXP or the WIZARD 
%recommendations against the NO PPI one.

Figures~\ref{fig:total_elapsed_time} and \ref{fig:recomm_qual} exhibit our evaluation results.
Figure~\ref{fig:total_elapsed_time} shows the times (in seconds) taken 
to run the 10Q and 20Q workloads over the 1TB and 3TB fact tables. 
Based on the run results, Figure~\ref{fig:recomm_qual} shows 
how much improvement was gained by the EXP and WIZARD solutions, compared with the NO PPI solution.

Overall, the WIZARD recommendations were very effective at partitioning 
the fact tables by leveraging the predicates in the workloads. 
On average, the \hbox{total} execution time of the workloads on the WIZARD solutions was 
about 3.5 and 2 times faster than those of NO PPI and EXP, respectively. 
Also, the quality of the \hbox{WIZARD} solutions outperformed that of the EXP solutions; 
the WIZARD solutions yielded an average of about 77\% improvement on the NO PPI solutions.

Furthermore, the WIZARD recommendations 
scaled very well with growing workload size and table size, 
as illustrated in Figure~\ref{fig:recomm_qual}. 
Although we doubled the workload scale from 10Q to 20Q, 
the performance improvement of the WIZARD solution stayed the same in Figure~\ref{fig:1T_quality} 
and got almost negligibly decreased in Figure~\ref{fig:3T_quality}. 
This result showed that our WIZARD \hbox{solution} was scalable over increasing workload size.

We also increased the table scale from 1TB to 3TB. 
For 10Q the performance improvement (about 77\%) was almost the same over the increasing scales, 
and for 20Q it was decreased from 76\% to 74\%, which was \hbox{almost} negligible.
The WIZARD partitioning \hbox{recommendation} scaled well with increasing table size.

In sum, our experiment results attest the effectiveness of the proposed algorithms for WIZARD 
and show the superiority of the recommendations of WIZARD.

\shorten{Regarding the scale of 1TB in Figure~\ref{fig:1T_quality}, 
although we doubled the workload scale (from 10Q to 20Q) 
the performance improvement of the WIZARD solution compared with that of NO PPI was not decrease although we doubled the workload scale (from 10Q to 20Q).
In Figure~\ref{fig:3T_quality}, for the 3TB scale the improvement of WIZARD was slightly decreased, but it was negligible.
In sum, the performance of the WIZARD solution was scalable over increasing workload and table scale.

%Also, we scaled up the table from 1TB to 3TB on the fixed workload, but 
%the \hbox{recommendation} quality was not degraded at all. 
These results attest the effectiveness of our algorithms, 
and show the superiority of our recommendations.}

\section{Related Work}
\label{sec:related_work}

Physical database design~\cite{finkelstein1988physical,labio1997physical,rozen1991framework,Zilio1998} 
has been discussed in \hbox{academic} research and industrial sectors in the past years.
The major DBMS vendors (e.g., IBM, MS, and Oracle) have driven much of the work. 
Their specific interests have been in automating the physical design for table 
partitioning~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated,sheet2009oracle,saa2007oracle,rao2002automating}, indexes/materialized views~\cite{agrawal2000automated,Agrawal04:sqlserver,dash2011cophy,Kimura11,itwsqlserver,valentin2000db2,zilio2004recommending}, 
and integration~\cite{Zilio04:db2design}.

%Lightstone's work~\cite{Lightstone04:db2auto} introduces the MDC (Multidimensional 
%Clustering) Advisor that automates the selection of MDC 
%keys customized for a specific workload on DB2 Universal Databases to use MDC structure. 
%DB2 Advisor~\cite{Zilio04:db2design} is an integrated commercial 
%tool that captures a variety of aspects of physical database design 
%by automation in a comprehensive way. 
%Oracle Partitioning Advisor~\cite{sheet2009oracle} for recommending 
%table partitionings is  integrated into 
%the SQL Access Advisor~\cite{saa2007oracle}.

%Since data clustering significantly influences on the performance of 
%multidimensional queries, the MDC advisor helps DBAs that may have 
%nontrivial troubles in finding proper clustering dimensions and 
%granularity. 
%To enumerate and determine possible clustering dimensions, 
%his work relies on {\it what-if} query cost modeling, data sampling and a search algorithm.

%DB2 Advisor~\cite{Zilio04:db2design} is an integrated commercial 
%tool that captures a variety of aspects of physical database design 
%by automation in a comprehensive way. 
%It makes an attempt to cover a variety of design choices-partitioning, 
%MDC, materialized views and indexes. 
%DB2 Advisor leverages the DB2 query optimizer for yielding a recommendation, and 
%takes a hybrid searching approach for considering different features.

%Oracle has a tool, called Partitioning Advisor~\cite{sheet2009oracle}, 
%to make a partitioning recommendation for tables. 
%The tool is now integrated into 
%the SQL Access Advisor~\cite{saa2007oracle} that recommends indexes and materialized 
%views. 
%However, its technical details are not yet released in public. 
%
%Also, MS SQL Server provides some similar, automated tools~\cite{agrawal2000automated,agrawal04:integrating,
%itwsqlserver} for table partitioning, materialized views or indexes fitting to a given workload.
%The optimizer cost model is exploited to acquire estimates for 
%candidate materialized views or indexes. 
%In particular, Agrawal's work~\cite{agrawal04:integrating} presents 
%an approach of integrating vertical and horizontal partitioning schemes 
%into their physical database design, considering scalability.

Some of the tools in IBM DB2~\cite{Lightstone04:db2auto}, 
Oracle~\cite{sheet2009oracle}, and MS SQL Server~\cite{agrawal04:integrating} 
appear to be similar to our MLPPI wizard. 
However, in light of problem scope and approach, the wizard 
is fundamentally different from the existing tools excluding Oracle Partitioning \hbox{Advisor}~\cite{sheet2009oracle} 
providing no published technical details.

DB2 MDC Advisor~\cite{Lightstone04:db2auto} actually tackles 
a different problem of automatically recommending the most well-suited MDC 
keys for a given workload. 
Agrawal's work~\cite{agrawal04:integrating} 
discusses another problem of merging \hbox{single-level} range partitionings on objects such as tables and indexes. 
This article addresses the multi-level \hbox{partitioning} problem. 
Hence, the existing solutions cannot be directly applied to our MLPPI wizard.

Regarding the approach, DB2 MDC Advisor~\cite{Lightstone04:db2auto} uses 
the search space driven by {\em fields}. In contrast, our search 
space is driven by query-predicates, which is superior because 
\hbox{predicates} are more customized and specific to a workload than fields. 
Agrawal's horizontal partitioning scheme~\cite{agrawal04:integrating} 
also uses the search space driven by simple range predicates, but 
his technique has a \hbox{shortcoming} as follows. 
His work produces a solution for each individual query and attempt 
to merge the solutions. But this approach cannot reach 
an optimized solution in a global \hbox{perspective}.
Our wizard generates the whole search space upfront and in turn merges partitions, leading to a globally-optimized \hbox{solution}. 
Moreover, only a single column is considered in his work~\cite{agrawal04:integrating}, 
whereas our tool deals with multiple fields.

Implementations of previous tools~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} \hbox{required} instrumentation for optimizer code.
These instrumentations are needed to facilitate the required 
information for the physical design tools API calls. 
The \hbox{instrumentation} code need to be enhanced and tested for new database 
releases that add complexity and additional cost for software upgrades. 
The Teradata optimizer has a rich set of existing APIs originally coded 
for system and workload management tools. 
The APIs are sufficient to avoid the costly optimizer code change.

Also, Nehme's work~\cite{nehme2011automated}, deeply-integrated with 
optimizer, reveals a concern about the quality of the recommendations 
made by some tools, shallowly integrated with optimizer. 
But we observed in our experiments that the quality of 
the wizard's solutions was much superior to that of the base solutions. 
In addition, some might argue that in our loosely-coupled approach 
the cost to invoke the optimizer might be significant. 
But the measured call counts were much fewer than 
the theoretical bound, since most calls were made only when queries were affected by a merge.  
%If the remote communication cost is expensive, the tool can also be run, 
%{\em co-located} with the server.

%The existing work requires their database engine to be extended 
%to support their features. 
%This strong coupling may impose a nontrivial burden on their servers whenever 
%a new feature of the existing tools is desired. 
%As illustrated in Figure~\ref{fig:arch}, our tool is completely outside 
%the DBMS server. Thus, we are free of the constraint. 
%A literature~\cite{nehme2011automated} claims that 
%such loosely-coupled approaches might suffer from the reduced performance of a tool 
%and the poor quality of the partitioning recommendations. 
%The powerful Teradata APIs, however, help the wizard receive from the server 
%everything needed for yielding a MLPPI recommendation. 
%Moreover, as demonstrated in Figure~\ref{fig:perf_eval}, the quality of the 
%partitioning solutions by the wizard was also much superior to that of the base solutions.
%Some might argue that the cost of invoking the optimizer could be significant as well. 
%Again, the actual calls are mainly made only when queries are affected by a merge. 
%As a result, they were much fewer than 
%the theoretical bound as shown in Table~\ref{tab:workload_stat}.  
%To further save the remote communication cost, the tool can also be run, 
%{\em co-located} with the server. 
%Therefore, we believe that the shallowly-integrated strategy perfectly matches 
%with the wizard.

Lastly, there has been some previous work~\cite{nehme2011automated,rao2002automating,tatarowicz2012lookup} 
regarding table partitioning in multi-node \hbox{systems}, but 
our problem is discussed in the context of a single node system, as in the existing work~\cite{agrawal04:integrating}. 
Database cracking~\cite{idreoskm07}
%, self-reorganizing column values in accordance with query workload, 
assumes a single node environment, but 
it does not address our multi-level partitioning problem.

\section{Summary}
\label{sec:conclusion}
%We verified the effectiveness of the tool by running it on a workload based

Given workloads, it is difficult for DBAs to select appropriate fields 
in partitioning the fact table due to large search space. 
DBAs cannot easily determine how granular partitions should be made 
for the workloads.

To address this concern, we presented the MLPPI wizard to recommend a fact-table partitioning for a given star schema workload. 
Since \hbox{query-predicates} are exploited 
to capture necessary ranges for fields and define partitions, 
the MLPPI \hbox{recommendation} can be very optimized, customized to the workload.

We proposed the wizard's algorithms consisting of the three phases. 
The wizard incrementally reduced its search space by merging 
the range pair with the least scan or query cost, and eventually reached an MLPPI recommendation. 
In addition, we analyzed the running complexity for the initial and optimized phases. 
Despite the theoretically high bound, in practice, the wizard made much fewer calls in the phases.

We measured the performance of the recommendation by the wizard using 
our workloads.  
We demonstrated that the produced MLPPI solutions by 
the wizard could reduce the total elapsed time by more than a factor of two, 
compared with those of no partitioning approach and partitioning 
done by a human expert.

\section*{Acknowledgments} 
This work was carried out while the first author visited Teradata. 
The first author is now at Korea Institute of Science and Technology Information (yksuh@kisti.re.kr).
%\clearpage
%\vspace\fill

%1. Agrawal, R., Srikant, R.: Fast algorithms for mining association rules. In: Proceedings of the 20th International Conference on Very Large Databases. pp. 487–499. Morgan Kaufmann, San- tiago, Chile (1994)
%2. Bruce,K.B.,Cardelli,L.,Pierce,B.C.:Comparingobjectencodings.In:Abadi,M.,Ito,T.(eds.) Theoretical Aspects of Computer Software, Lecture Notes in Computer Science, vol. 1281, pp. 415–438. Springer-Verlag, Berlin Heidelberg New York (1997)
%3. Garcia-Molina, H., Ullman, D.J., Widom, J.: Database Systems: The Complete Book. Prentice Hall, New Jersey, USA (2002)
%4. van Leeuwen, J. (ed.): Computer Science Today. Recent Trends and Developments, Lecture Notes in Computer Science, vol. 1000. Springer-Verlag, Berlin Heidelberg New York (1995)
%5. Ribicˇre,M.,Charlton,P.:Ontologyoverview.MotorolaLabs,Paris(2002),[Online].Available:
%http://www.fipa.org/docs/input/f-in-00045/f-in-00045.pdf (current October 2003)
%6. Wang,X.,Bettini,C.,Brodsky,A.,Jajoida,S.:Logicaldesignfortemporaldatabaseswithmul-
%tiple granularities. ACM Transactions on Database Systems 22(2), 115–170 (1997)

%\begin{thebibliography}{99}
%\expandafter\ifx\csname url\endcsname\relax
%  \def\url#1{\texttt{#1}}\fi
%\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
%\expandafter\ifx\csname href\endcsname\relax
%  \def\href#1#2{#2} \def\path#1{#1}\fi
%
%{
%\bibitem{Suh12}
%{Y.-K.~Suh, A.~Ghazal, A.~Crolotte, and P.~Kostamaa}, {A New Tool for Multi-level Partitioning in Teradata}, 
%in: CIKM'12, pp. 2214--2218.
%  
%\bibitem{sinclair:ppi}
%P.~Sinclair, {Using PPIs to Improve Performance},
%  \url{http://www.teradata.com/tdmo/v08n03/pdf/AR5731.pdf}
%  (accessed August 27, 2015).
%
%\bibitem{klindt09mlppi}
%J.~Klindt, {Single-level and Multilevel Partitioned Primary \linebreak Indexes},
%  \url{http://www.teradata.com/white-papers/Single-level-and-}\linebreak\url{Multilevel-Partitioned-Primary-Indexes-eb1889/}
%  (accessed July 27 2015).
%
%\bibitem{oneil:ssb}
%{P.~O’Neil, B.~O'Neil, and X.~Chen}, {The Star Schema Benchmark (SSB)},
%  \url{http://www.cs.umb.edu/~poneil/StarSchemaB.PDF} 
%  (accessed February 28, 2017).
%
%\bibitem{agrawal04:integrating}
%{S.~Agrawal, V.~Narasayya, and B.~Yang}, {Integrating Vertical and Horizontal Partitioning into Automated Physical Database Design}, 
%in: \hbox{SIGMOD} '04, pp. 359--370.
%
%\bibitem{Lightstone04:db2auto}
%{S.~Lightstone and B.~Bhattacharjee}, {Automated Design of
%  Multi-dimensional Clustering Tables for Relational Databases}, in:
%  VLDB '04, pp. 1170--1181.
%
%\bibitem{nehme2011automated}
%{R.~Nehme and N.~Bruno}, {Automated Partitioning Design in Parallel
%  Database Systems}, in: SIGMOD '11, pp. 1137--1148.
%
%\bibitem{chaudhuri1998autoadmin}
%{S.~Chaudhuri and V.~Narasayya}, {AutoAdmin `What-If' Index Analysis Utility}, ACM SIGMOD Rec. 27~(2) (1998) 367--378.
%
%\bibitem{finkelstein1988physical}
%{S.~Finkelstein, M. Schkolnick, and P. Tiberio}, {Physical Database Design for Relational Databases}, ACM Trans. on Datab. Syst. (TODS), 13~(1) (1988) 91--128.
%
%\bibitem{labio1997physical}
%{W.~Labio, D.~Quass, and B.~Adelberg}, {Physical Database Design for Data Warehouses}, in: ICDE '97, pp. 277--288.
%
%\bibitem{rozen1991framework}
%{S.~Rozen and D.~Shasha}, {A Framework for Automating Physical Database Design}, in: VLDB '91, pp. 401--411.
%
%\bibitem{Zilio1998}
%{C.~Zilio}, {Physical Database Design Decision Algorithms and Concurrent
%  \hbox{Reorganization} for Parallel Database Systems}, Ph.D. thesis,
%  Dept. of CS, Univ. of Toronto (1998).
%
%\bibitem{sheet2009oracle}
%{Oracle}, {Oracle White Paper---Partitioning with \hbox{Oracle} Database 11g
%  \hbox{Release 2}},
%  \url{http://www.oracle.com/technetwork/database/options/partitioning/twp-partitioning-11gr2-2009-09-130569.pdf}
%  (accessed June 23, 2014).
%
%\bibitem{saa2007oracle}
%{Oracle}, {SQL Access Advisor},
%  \url{http://docs.oracle.com/cd/B19306_01/server.102/b14211/advisor.htm}
%  (accessed March 30, 2012).
%
%\bibitem{rao2002automating}
%{Jun Rao, Chun Zhang, Nimrod Megiddo, and Guy Lohman}, {Automating
%  Physical Database Design in A Parallel Database}, in: SIGMOD '02, pp. 558--569.
%
%\bibitem{agrawal2000automated}
%{S.~Agrawal et. al.}, {Automated Selection of Materialized Views and Indexes in SQL Databases}, in:
%  {VLDB '00}, pp. 496--505.
%
%\bibitem{Agrawal04:sqlserver}
%{S.~Agrawal et. al.}, {Database Tuning Advisor for Microsoft
%  SQL Server 2005: Demo}, in: \hbox{SIGMOD} '05, pp. 930--932.
%
%\bibitem{dash2011cophy}
%{D.~Dash}, {CoPhy: A
%  Scalable, Portable, and Interactive Index Advisor for Large Workloads},
%  in: PVLDB, 4~(6) (2011) 362--372.
%
%\bibitem{Kimura11}
%{H.~Kimura et. al.}, {Compression Aware
%  \hbox{Physical} Database Design}, in: PVLDB, 4~(10) (2011) 657--668.
%
%\bibitem{itwsqlserver}
%{Microsoft SQL Server 2000}, {Index Tuning Wizard SQL Server 2000},
%  \url{http://technet.microsoft.com/en-us/library/cc966541.aspx} 
%  (accessed on March 21, 2014).
%
%\bibitem{valentin2000db2}
%{G.~Valentin}, {DB2 Advisor: An Optimizer Smart Enough to Recommend Its Own
%  Indexes}, in: {ICDE '00}, pp. 101--110.
%
%\bibitem{zilio2004recommending}
%{D.~Zilio et. al.},
%  {Recommending Materialized Views and Indexes with the IBM DB2 Design
%  Advisor}, in: ICAC '04, pp. 180--188.
%
%\bibitem{Zilio04:db2design}
%{D.~Zilio et. al.}, {DB2 Design Advisor:
%  Integrated Automatic Physical Database}, in: VLDB '04, pp. 1087--1097.
%
%\bibitem{tatarowicz2012lookup}
%{A.~Tatarowicz et. al.}, {Lookup Tables: Fine-grained Partitioning for Distributed Databases},
%  in: ICDE'12, pp. 102--113.
%
%\bibitem{idreoskm07}
%{S.~Idreos et. al.}, {Database Cracking}, in: CIDR'07, pp. 68--78.
%}
%
%\end{thebibliography}

\bibliographystyle{splncs03}
\bibliography{paper}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% For the final version of the paper: %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Author information
%\vspace{4ex}\noindent
%\textbf{Author One} is\dots
%
%\bigskip\noindent
%\textbf{Author Two} is\dots
%
%\bigskip\noindent
%\textbf{Author Three} is\dots

%% Reception and acceptance information
%\bigskip
%\paragraph{Received: Month DD, 20YY; Accepted: Month DD, 20YY.}

\end{document}