### 2017-03-10 pt_characterization/repeatability_test/paper.tex
Chracterizing draft

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{A Study of Experiment Repeatability}

\author{
Young-Kyoon Suh\\
}
\maketitle

\section{Description}
This document presents histograms rendered based on empirical data 
of a program under test, called {\em INC}. 
Our experiment increases a task length of {\em INC} 
from one second to 4096 seconds by a factor of two.
We already ran this experiment last month (February 2017), 
but we reran the experiment this month (March 2017) 
to check if experiment repeatability holds true.

\section{Experiment Notes}
Table~\ref{tab:exp_notes} provides a short description of our experimental runs, 
on which the following histograms are based.

%% done
\begin{table}[h]
\begin{center}
\begin{tabular}{|p{2cm}|p{3cm}|p{6cm}|p{4cm}|} \hline
Machine & Task Length (sec) & Description & Experiment Period\\ \hline
{\tt sodb9} &  INC1$\sim$INC512 & Each run with 1000 samples & 2017-03-02 $\sim$ 2017-03-07\\ \hline
%{\tt sodb10} & INC2048 & A run of 300 samples & 2017-02-12 $\sim$ 2017-02-20\\ \hline
%{\tt sodb12} & INC4096 & A run of 300 samples & 2017-02-12 $\sim$ 2017-02-27\\ \hline
%{\tt sodb12} & INC8192 & Two runs, each with 40 samples & 2017-01-16 $\sim$ 2017-01-20 / 2017-01-25 $\sim$ 2017-01-29\\ \hline
%{\tt sodb12} & INC16384 & Two runs, each with 40 samples & 2017-01-05 $\sim$ 2017-01-13 / 2017-01-29 $\sim$ 2017-02-06\\ \hline 
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Notes on experiment runs used for histograms\label{tab:exp_notes}}
\end{table}

Now we show histograms of elapsed time (ET) and process time (PT) of INC.

\newpage

\input{sodb9/res.tex}

\end{document}